### 2023-05-07 paper/paper.tex
Updates from Overleaf

% $Id$
\documentclass[twocolumn,prd,nofootinbib]{revtex4}

\newcommand\ForInternalReference[1]{}
\newcommand\SkipForEarlyCirculation[1]{}
\newcommand\AddedResponse[1]{{\color{blue} {#1}}}
%\newcommand\SkipForEarlyCirculation[1]{#1}
\newcommand\SkipPP[1]{}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}
\usepackage{color}
\usepackage{xspace}
\usepackage{url}
\usepackage{amsmath}
%\usepackage{adjustbox}
\usepackage{float}
\usepackage{multirow}
%
%
\usepackage{times}
%
%
%
\newcommand\optional[1]{}

%
\newcommand\E[1]{\left\langle #1\right\rangle}
\newcommand\qmstate[1]{\left|#1\right \rangle}
\newcommand\qmstateKet[1]{\left\langle#1\right|}
\newcommand\qmstateproduct[2]{\left\langle#1|#2\right\rangle}
\newcommand\qmoperatorelement[3]{\left\langle#1\left|#2\right|#3\right\rangle}
\newcommand\qmoperator[1]{{\bf #1}}
%
\newcommand\Y[1]{{{}_{#1}Y}}

\newcommand\lnL{ \ln {\cal L}}
\newcommand\lnLmarg{ \ln{\cal L}_{\rm marg}{}}
\newcommand\unit[1]{{\rm #1}}

\newcommand\rapidPEOrig{rapid\_pe1}
\newcommand\ILE{ILE}
\newcommand\editremark[1]{{\color{red} #1}}
%
%
%
\usepackage{color}
\definecolor{amber}{rgb}{1.0, 0.75, 0.0}
\definecolor{orange}{rgb}{1.0, 0.5, 0.0}
\definecolor{amaranth}{rgb}{0.9, 0.17, 0.31}
\def\fixme#1{\textcolor{red}{#1}}
\newcommand{\Richard}[1]{ {\color{blue}{#1}}}
\newcommand{\ros}[1]{ {\color{blue}{#1}}}
%

%

%
%
%
%
\graphicspath{{./figures/}}
\newcommand{\mc}{{\cal M}}
\newcommand{\Ms}{M_{\odot}}
\newcommand\LambdaTilde{\widetilde{\Lambda}}
\newcommand\DeltaLambdaTilde{\delta \widetilde{\Lambda}}
%
\def\ltsima{$\; \buildrel < \over \sim \;$}
\def\simlt{\lower.5ex\hbox{\ltsima}}
\def\gtsima{$\; \buildrel > \over \sim \;$}
\def\simgt{\lower.5ex\hbox{\gtsima}}

\newcommand\prx{Phys.~Rev.~X}
\def\aj{Astronomical Journal}                 %
\def\apj{Astrophysical Journal}                %
\def\apjl{Astrophysical Journal}             %
\def\pasj{PASJ}
\def\apjs{ApJS}              %
\def\mnras{MNRAS}            %
\def\prd{Phys.~Rev.~D}       %
\def\prl{Phys.~Rev.~Lett}    %
\def\cqg{Class.~Quant.~Grav.~}%
\def\araa{ARA\&A}             %
\def\nat{Nature}              %
\def\aap{A\&A}                %
\def\aapr{A\&A~Rev.~}    %
\def\pasp{PASP}    %
\def\sovast{Soviet Ast.}
%
%

\newcommand{\IMRPD}{\textsc{IMRPhenomD}\xspace}
\newcommand{\IMRPDT}{\textsc{IMRPhenomD\_NRTidal}\xspace}
\newcommand{\IMRP}{\textsc{IMRPhenomPv2}\xspace}
\newcommand{\SEOBP}{\textsc{SEOBNRv3}\xspace}
\newcommand{\SEOBA}{\textsc{SEOBNRv4}\xspace}
\newcommand{\SEOBAROM}{\textsc{SEOBNRv4\_ROM}\xspace}
\newcommand{\NRSur}{NRSur7dq2\xspace}
\newcommand{\TEOB}{SEOBNRv4T\xspace}
\newcommand{\Resum}{TEOBResumS\xspace}
\newcommand\RIFT{RIFT}
\newcommand{\Taylor}{TaylorF2\xspace}
\newcommand\PaperDetection{\underline{LVC-detect}\cite{DiscoveryPaper}}
\newcommand\PaperPE{\underline{LVC-PE}\cite{PEPaper}}
\newcommand\PaperTestGR{\underline{LVC-TestGR}\cite{TestingGRPaper}}
\newcommand\PaperPENRMethods{\underline{PE+NR-Methods}\cite{gwastro-PENR-Methods-Lange}}
\newcommand\PaperAstro{\underline{LVC-Astro}\cite{AstroPaper}}
\newcommand\PaperBurst{\underline{LVC-Burst}\cite{BurstPaper}}
\newcommand\PaperRates{\underline{LVC-Rates}\cite{RatesPaper}}
\newcommand\PaperStochastic{\underline{LVC-Stochastic}}
\newcommand\PaperSEOBNRvthree{\underline{LVC-SEOBNRv3}\cite{SEOBv3Paper}}

\def\RIT{Center for Computational Relativity and Gravitation, Rochester Institute of Technology, Rochester, New York 14623, USA}

\begin{document}

\title{Population Inference of Eccentric Black Holes Binaries}
\author{M. Zeeshan}
\affiliation{\RIT}
\author{R. O'Shaughnessy}
\affiliation{\RIT}
\begin{abstract}
Yet to work. 
\end{abstract}
\maketitle

\section{Introduction}
The Advanced LIGO  \cite{2015CQGra..32g4001T}  and Virgo \cite{gw-detectors-Virgo-original-preferred}  ground-based gravitational wave (GW) detectors have
identified several coalescing compact binaries
\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns}.  
%
The properties of the sources responsible have been inferred via Bayesian comparison of  each source with candidate
gravitational wave signals
\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns,LIGO-O2-Catalog,gwastro-PE-AlternativeArchitectures,gwastro-PENR-RIFT,gw-astro-PE-lalinference-v1}.
Many more are expected as observatories reach design sensitivity \cite{2016LRR....19....1A}.  
Both to handle the rapid pace of discovery and particularly to enable coordinated multimessenger followup, GW
observatories should be able to reconstruct the evidence for a signal being present in the data along with the full source parameters of coalescing binaries as fast as possible
\cite{LIGO-2016-WhitePaper,LIGO-Program-2018}.
However, particularly when using the best-available waveform models, these calculations can be very costly.  For binary
neutron stars, detailed inferences even using simplified waveforms take weeks.  Even for massive binary  black holes,
these calculations can take  months for costly time-domain effective-one-body models which incorporate the effects of precession.

% BORROWED FROM ROM: edit
Several strategies have been
developed to reduce the 
%overall computational cost and turnaround time of 
computational cost of parameter estimation \cite{gwastro-pe-Brandon-STF2,gwastro-PE-AlternativeArchitectures,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,2016PhRvD..94d4031S,2017CQGra..34k5006V,2018arXiv180608792Z}.  
% The computational cost of parameter estimation can be reduced 
Approaches that have appeared in the literature include
%by 
generating the approximate solutions more quickly \cite{gwastro-mergers-PE-ReducedOrder-2013,2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C,gwastro-mergers-IMRPhenomP,gwastro-SpinTaylorF2-2013}; 
%by 
interpolating some combination
of the waveform or likelihood \cite{gwastro-approx-ROMNR-Blackman2015,2013PhRvD..87d4008C,2013PhRvD..87l2002S,2014CQGra..31s5010P,2014PhRvD..90d4074S,gwastro-PE-AlternativeArchitectures,cole2014likelihood,2012MNRAS.421..169G};  or 
%by 
adopting a sparse representation to reduce the computational cost of data
handling \cite{antil2013two,gwastro-mergers-PE-ReducedOrder-2013,2016PhRvD..94d4031S,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,gwastro-PE-AlternativeArchitectures,2010arXiv1007.4820C,2018arXiv180608792Z}.   % 2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C
\AddedResponse{Not published but also important are code optimization projects that improve    infrastructure,
  such as better parallelization for \texttt{lalinference} \cite{gw-astro-PE-lalinference-v1,lalsuite}.}
 Some methods, however, achieve rapid turnaround through simplifying approximations.  
%
The RIFT/rapidPE strategy described in \cite{gwastro-PE-AlternativeArchitectures,2017CQGra..34n4002O,gwastro-PENR-RIFT,NRPaper} eschews these simplifications, performing
embarrassingly-parallel inferences even for costly models.   However, the method as previously implemented still had a
significant net  computational cost and noticeable startup time, limiting its diverse potential applications.
%
Previously, B. Miller  developed a custom-coded  implementation of the
relevant likelihood on graphics processing units (GPUs), suggesting substantial performance improvements were possible
\cite{Thesis-Brandon-2016,GPU-NU-work}.  
\AddedResponse{The pycbc search code\cite{2016CQGra..33u5004U} and its pycbc-inference extension
  \cite{2019PASP..131b4503B} can also make use of several GPU
  hardware-accelerated operations.
}
In this paper,  we describe a variant of the RIFT approach which flexibly translates one of its
algorithms to operate on GPUs, and adjusts its workflow to exploit the speed improvements
this re-implementation affords.

This paper is organized as follows.
In Section \ref{sec:methods}, we review the underlying marginalized likelihood calculations used by RIFT and  their updated GPU
implementation.  
In Section \ref{sec:demo}, we quantify the improved performance of our  GPU-accelerated code,   while assessing
operating settings which facilitate increased performance.
In Section \ref{sec:end-to-end}, we describe the performance of the end-to-end pipeline on the synthetic nonspinning signals used in
Section \ref{sec:demo}.  
In Section \ref{sec:conclude}, we summarize our results and discuss their potential applications to future GW source and
population inference.

\section{Methods}
\label{sec:methods}

I need to start writing from here.

\subsection{Power law Model}

% \begin{equation}
% \label{eq:lnL}
% \ln {\cal L}(\bm{\lambda} ,\theta )=-\frac{1}{2}\sum\limits_{k}\langle h_{k}(\bm{\lambda} ,\theta )-d_{k} |h_{k}(\bm{\lambda} ,\theta )-d_{k}\rangle _{k}-\langle d_{k}|d_{k}\rangle _{k},
% \end{equation}

% For the remaining dimensions, to evaluate the likelihood in regions of high importance, we use an adaptive Monte Carlo as described in
% \cite{gwastro-PE-AlternativeArchitectures}.

% \begin{align} \label{eq:strain_mode}
% h(t,\vartheta,\phi;\bm{\lambda}) = 
% \sum_{\ell=2}^{\infty} \sum_{m=-\ell}^{\ell} \frac{D_{\rm ref}}{D} h^{\ell m}(t;\bm{\lambda}) \Y{-2}_{\ell m} \left(\vartheta, \phi \right) \, ,
% \end{align}

% \begin{widetext}
% \begin{align}
% \ln {\cal L}(\bm{\lambda}, \theta) 
% &= (D_{\rm ref}/D) \text{Re} \sum_k \sum_{\ell m}(F_k \Y{-2}_{\ell m})^* Q_{k,lm}(\bm{\lambda},t_k)\nonumber \\
% &   -\frac{(D_{\rm ref}/D)^2}{4}\sum_k \sum_{\ell m \ell' m'}
% \left[
% {
% |F_k|^2 [\Y{-2}_{\ell m}]^*\Y{-2}_{\ell'm'} U_{k,\ell m,\ell' m'}(\bm{\lambda})
% }
% % \right. \nonumber \\ & \left.
%  {
% +  \text{Re} \left( F_k^2 \Y{-2}_{\ell m} \Y{-2}_{\ell'm'} V_{k,\ell m,\ell'm'} \right)
% }
% \right]
% \label{eq:def:lnL:Decomposed}
% \end{align}
% \end{widetext}

% \begin{eqnarray}
% {\cal L}_{\rm margT} \equiv  \int {\cal L} \frac{dt}{T}
% \label{eq:lnL:tmarg}
% \end{eqnarray}

\subsection{Tradeoff between Monte Carlo integration and accuracy}

Each marginalized likelihood evaluation has a relative uncertainty of order  $1/\sqrt{n_{\rm eff}}$, where the number of
effective samples $n_{\rm eff}=\epsilon N_{\rm it}$ increases linearly with the total number $N_{\rm it}$ of Monte Carlo
evaluations performed by \ILE{}.  Therefore, to increase (decrease) our accuracy for each likelihood evaluation by a factor of $A$
will require a factor $A^2$ more (fewer) iterations.   
%
We adopt a fixed threshold on $n_{\rm eff}$.

We do benefit slightly by re-using the adaptive Monte Carlo integrator for each extrinsic point $\bm{\lambda}$.  Since the
integrator has already identified the likely range of sky locations and distances on the first iteration, each
subsequent evaluation of the marginalized likelihood can converge marginally more quickly.

\section{Analysis of marginalized likelihood evaluation cost}
\label{sec:demo}

We illustrate the code using two synthetic signals: a binary black hole with masses $m_{1,2}=35,30 M_\odot$ and a binary
neutron star with masses $1.4 M_\odot$ and $1.35 M_\odot$.
 We perform this analysis on heterogeneous LVC collaboration computing resources \AddedResponse{described in more detail
   in the Appendix}, to assess our variable
performance across architectures.  Notably, we investigate the following GPU options: (a) GTX 1050 Ti at LIGO-CIT,
available in large numbers; (b) V100, available on selected high-performance machines; (c) and
GTX 1050 at LIGO-LHO.    Unless otherwise noted in the text, we will discuss code configurations using CPU-only and GPU (b) using
$4096\unit{Hz}$ sampling and only the  $\ell=2, m=\pm 2$ modes.  In this section, we  conservatively report computational cost for
parameters $\bm{\lambda}$ which are close to or within the posterior.  Because of their consistency with the data, 
the marginalized likelihood calculations converge the most slowly, as they have the best-determined extrinsic
parameters.

\subsection{Synthetic source generation and analysis settings}

We generate  synthetic data for a two-detector LIGO configuration, assuming operation at  initial (BBH) or advanced (BNS) design  sensitivity.
Both signals are generated with SEOBNRv4 effective-one-body waveform approximation \cite{2017PhRvD..95d4028B}, with a zero-noise data realization.   
To qualitatively reproduce the noise power spectrum and amplitude of typical binary black hole observations in O1 and
O2, the binary black hole signal has source distance $200 \unit{Mpc}$, chosen so SNR $\simeq 14$.  
Similarly, to qualitatively reproduce the analysis of GW170817, our synthetic binary neutron star is placed $100
\unit{Mpc}$ away, so the signal amplitude is SNR $\simeq 31$.

 We perform a multi-stage iterative RIFT analysis of these signal with the
time-domain SEOBNRv4 (BBH) or TaylorT4 \cite{gw-astro-PN-Comparison-AlessandraSathya2009}( BNS) approximation, under the simplifying assumption that both objects are point particles with zero
spin.   For both sources, we adopt the fiducial distance prior $p(D_L) \propto D_L^2$,   selecting  a maximum distance
roughly five times larger than the known source distance: 1Gpc for the binary black hole and 500 Mpc for the neutron
star.  We adopt a uniform prior on the component (redshifted) masses $m_{i,z}$ (and, when appropriate, spins
$\chi_{i,z}$).
Unless otherwise noted, all mass quantities described in this work are redshifted masses $m_{k,z}=(1+z)m_k$.
For the binary black hole we generate the signal with $f_{\rm min}=8 \unit{Hz}$ and analyze the signal with  $f_{\rm
  min}=10 \unit{Hz}$; for the binary neutron star, we generate the signal with $f_{\rm min}=20\unit{Hz}$ and anayze it
with $f_{\rm min}=23\unit{Hz}$.

\subsection{Massive compact binary black holes}
Previously, each instance of ILE examined one intrinsic point $\bm{\lambda}$.  The overall cost of this ILE evaluation
involved two parts: a startup cost, a setup cost, and a Monte Carlo integration cost.  The  startup cost $\tau_{start}$
is associated with code setup followed by reading, data conditioning, and Fourier-transforming the
data.  The setup cost $\tau_{setup}$ arises from waveform generation and  the inner products $U,V,Q$.  Finally, the Monte Carlo
cost $\tau_{mc}=N_{ad} \tau_{ad}+N_{it}\tau_{eval}$ increases with the number of Monte Carlo iterations $N_{it}$ with
and the number of times $N_{ad}$ the sampling prior used in adaptive Monte Carlo is regenerated from the most recent 
$n_{\rm chunk}$ data points, in proportion to their
cost.   In the standard configuration, the adaptive sampling prior is regenerated every $n_{\rm chunk}$ iterations (i.e,
 $N_{ad} \simeq N_{it}/n_{chunk}$), regenerating a sampling prior in two sky location parameters and distance.  
The choice of one $\bm{\lambda}$ for each instance of ILE was due to the substantial time  $\tau_{MC}$, which vastly
dominated the overall computational cost.   For example, for a typical analysis of a short signal --  a typical binary
black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$ --  this version has
cost elements as shown in Table \ref{tab:CostBreakdown}, based on the assumption that the MC terminates after $N_{\rm it}\simeq 2\times 10^6$ iterations using
$n_{\rm chunk}\simeq 10^4$.  
The marginalized likelihood described above  converges incredibly rapidly to a small value if the candidate model is
inconsistent with a signal (or the absence thereof) in the data, with only $N_{it} \simeq O(10^4)$  evaluations needed.

\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% ~/parse_report.sh profile_nogpu_pcdev13.log | more
CPU & 16384 & $\pm 2 $ & 20 & 2.4 &&540 & 20 &  690  \\ 
       & 4096 & $\pm 2 $ &   20  &&&& 20 \\ \hline
% ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
% /parse_report.sh ./profile_nogpu_HM_lowres_pcdev13.log 
%    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
%   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
%    
CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & 20 &  1060  \\ 
       & 4096 & $ \pm 2, \pm 1 $ &   20 &&&& 20  \\ \hline
% LIGO-WA, or a compute node at CIT with a grpahics catd: 
%  - old compute node: one marginalized likelihood in 2 minutes (!), 2x8
%  - CIT jobs, on compute nodes
%        20 points at 16kHz take about 90 min
%        20 points at kHz take anywhere from
%        condor_userlog iteration_0_ile/logs/*.log
GPU (a) & 16384 & $\pm 2 $  & 20  & & && & 270 \\
            & 4096 &$\pm 2 $  &  20 &  & & & & 45 \\ \hline
%% % LIGo CIT, pcdev13, 16 kHz : basically indistinguishable from cost at 4kHz
%%       1    0.027    0.027  560.572  560.572 integrate_likelihood_extrinsic_batchmode:8(<module>)
%%        20    0.082    0.004  552.004   27.600 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
%%        20   76.306    3.815  503.342   25.167 mcsampler.py:327(integrate)
%%      2345   57.019    0.024  241.951    0.103 mcsampler.py:198(draw)
%%     25785    0.135    0.000   99.535    0.004 shape_base.py:286(hstack)
%%     26561   99.102    0.004   99.102    0.004 {numpy.concatenate}
%%     21981    0.933    0.000   87.819    0.004 function_base.py:2063(__call__)
%%     21981   34.750    0.002   86.887    0.004 function_base.py:2154(_vectorize_call)
%%      2345    1.391    0.001   86.327    0.037 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
%%      2345    0.007    0.000   46.648    0.020 __init__.py:630(asnumpy)
%%      2345   46.557    0.020   46.640    0.020 {method 'get' of 'cupy.core.core.ndarray' objects}
%%        20    0.179    0.009   43.145    2.157 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
%%        20    0.174    0.009   37.498    1.875 lalsimutils.py:2582(hlmoft_SEOB_dict)
%%        20    0.099    0.005   37.324    1.866 lalsimutils.py:2743(complex_hoft)
%%        20   37.220    1.861   37.220    1.861 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
%%    444365    0.761    0.000   35.303    0.000 polyint.py:62(__call__)
%%      2345    5.260    0.002   33.956    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
%%       407    0.005    0.000   28.757    0.071 odepack.py:28(odeint)
%%       407    1.327    0.003   28.740    0.071 {scipy.integrate._odepack.odeint}
%%    444365    2.644    0.000   26.423    0.000 interpolate.py:657(_evaluate)
%% 117270874   22.569    0.000   22.569    0.000 mcsampler.py:685(uniform_samp)
% LIGO-CIT, pcdev13, 4kHz
%  - 20 iterations, about 450  to 500 sec to eval (i.e., about 25 sec per lnL)
%  -  precompute: can time manually
%   -  Overlap stages: PrecomputeLIkelihood + waveform, about 4
%  - Histogram time: about (12+30) seconds overall (histogram + ODEint)
%  - Actual likelihood evaluation: about 30 sec out of 20 eval, or about 1.5 sec
   %%   1    0.061    0.061  506.265  506.265 integrate_likelihood_extrinsic_batchmode:8(<module>)
   %%     20    0.194    0.010  498.581   24.929 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
   %%     20   80.397    4.020  464.263   23.213 mcsampler.py:327(integrate)
   %%   2140   54.358    0.025  235.480    0.110 mcsampler.py:198(draw)
   %%  23530    0.134    0.000  101.280    0.004 shape_base.py:286(hstack)
   %%  24306  100.861    0.004  100.861    0.004 {numpy.concatenate}
   %%  20136    0.896    0.000   84.698    0.004 function_base.py:2063(__call__)
   %%  20136   33.608    0.002   83.802    0.004 function_base.py:2154(_vectorize_call)
   %%   2140    1.403    0.001   45.791    0.021 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
   %% 461726    0.790    0.000   36.137    0.000 polyint.py:62(__call__)
   %%   2140    5.496    0.003   30.829    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
   %%    407    0.006    0.000   29.662    0.073 odepack.py:28(odeint)
   %%    407    1.360    0.003   29.644    0.073 {scipy.integrate._odepack.odeint}
   %%     20    0.052    0.003   28.785    1.439 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
   %% 461726    2.809    0.000   26.983    0.000 interpolate.py:657(_evaluate)
   %%     20    0.048    0.002   24.454    1.223 lalsimutils.py:2582(hlmoft_SEOB_dict)
   %%     20    0.030    0.002   24.406    1.220 lalsimutils.py:2743(complex_hoft)
   %%     20   24.373    1.219   24.373    1.219 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_welford | more
% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_16kHz | more
%   'residual' time: take mcsampler, subtract likleihood, divide by number of calls and number of evals per chunk
GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 20 &28 & 15\\
       & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 & 20  & 25\\ \hline
% ./parse_report.sh 20190130-profile_gpu_HM_16kHz_pcdev13.log | more
% ./parse_report.sh 20190130-profile_gpu_HM_lowres_pcdev13.log  | more
% ~/parse_report.sh profile_gpu_HM_lowres_pcdev13.log | more
%   - extra eval cost: (mcsampler - likelihood)/20/n_eval_likelihood/10^4  
GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & 20  & 38  \\
       & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  & 20 & 35 & \\ \hline
%
% LHO or LLO worker nodes
%  ./parse_report.sh 20190130-profile_gpu_LHOworker_16kHz.log | more
% ./parse_report.sh 20190130-profile_gpu_LHOworker_lowres.log | more
%   * worker iteration cose: (3027 - 687)/20/3851/10^4 * 10^6 // N
GPU (c) & 16384 & $\pm 2 $  & 20  &6  & & 18&  58&160 &  \\
            & 4096 &$\pm 2 $  &  20 & 3.7 & & 11  & 58  & 140 & $\simeq 50$ \\
\end{tabular}
% Compute node at LIGO-WA
\caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
  marginalized likelihood on default
  hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
  $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.
%\editremark{Comment: will update last number based on bulk profiling of a whole run; other numbers are from deep dive into individual run}
}
\end{table*}

With the new low-cost ${\cal L}_{\rm margT}$ evaluations, however, the startup and setup costs  $\tau_{start},\tau_{setup}$ now can be
comparable or in excess of the total time used to evaluate the marginalized likelihood ${\cal L}_{marg}$.   In fact, the total time per Monte Carlo
evaluation $\tau_{it}=\tau_{manage}+\tau_{eval}$  spent in general-purpose overhead $\tau_{manage}$ will be much larger than the time $\tau_{eval}$ spent
carrying out scientific calculations by evaluating the likelihood.  For these reasons, we reorganize the workflow, so each
instance of ILE loops over
$N_{eval}$ different choices for $\bm{\lambda}$. 
%
Additionally, particularly for massive binary black holes, we investigate two additional performance improvements.  First and
foremost, we lower the sampling rate, thus lowering the startup cost $\tau_{start}$ and particularly setup cost
$\tau_{setup}$.  Previously and out of an abundance of caution, \ILE{} employed a sampling rate $1/\Delta t = 16384
\unit{Hz}$ for all calculations, but terminated all inner products in Eq. (\ref{eq:overlap}) at roughly $f_{\rm
  max}=2048\unit{Hz}$ or less.   Reducing the sampling rate by a factor $s$ will \AddedResponse{reduce} the  cost of all operations with
timeseries -- they are shorter.  Depending on the relative cost of overhead versus array operations,  $\tau_{setup}$
\emph{and} the cost per evaluation $\tau_{eval}$ decrease modestly because of this effect.  
Also following Pankow et al, to insure safe inner products over short data sets in the presence of very narrow spectral lines, we operated on a
significantly-padded data buffer and modified the noise power spectrum by truncating the inverse power spectrum to a
finite response time.  For binary black holes, the degree of padding was significantly
in excess of the signal duration.  
Instead and following lalinference \cite{gw-astro-PE-lalinference-v1}, we adopt a much shorter inverse spectrum truncation length, a much shorter padding buffer, and
a Tukey window applied to the data to remove discontinuities at the start and end of each segment.   
Reducing the duration of data analyzed by a factor $s'$ will reduce the cost $\tau_{setup}$ by a factor $s'$.  
%
Combining these factors together, the overall computational cost  $T_{ILE}/N_{eval}$  of each marginalized likelihood
evaluation is
\begin{align}
T_{ILE}/N_{\rm eval} &= \frac{\tau_{start}}{N_{\rm eval}} 
 + 
 [ \tau_{setup} + N_{ad} \tau_{ad} + N_{it}\tau_{it}
 ] 
\end{align}
The first rows in table Table  \ref{tab:CostBreakdown} shows our estimated breakdown of these elements of the computational cost, for a typical binary black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$.
Because of the extremely high cost of adaptive overhead, we only adapt in two parameters, corresponding to sky location; 
and only adapt until a marginalied likelihood evaluation  returns significant $\ln {\cal L}_{\rm marg}$.  As a result, $N_{ad}$
(the number of adaptive stages \emph{per likelihood evaluation}) scales as $N_{it}/n_{chunk}/N_{\rm eval}$ and does not
increase with the overall number of samples, so the overall evaluation time scales as
\begin{align}
T_{ILE,mod}/N_{\rm eval} &= \frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \nonumber \\ &
 + 
 [ \tau_{setup}  + N_{it}\tau_{it}
 ]  \\
% (20 \[Tau]s1 + \[Tau]ad 1 2*10^6/10^4)/   20 + (\[Tau]start*1 + 2*10^6 \[Tau]it1 *10^(-6)) // Expand
& \simeq (\tau_{\rm start}/20\unit{s}) +  10 (\tau_{ad}/\mu s) 
 \nonumber \\ &
+ 2  [(\tau_{it,liike}/\mu s) + \tau_{it,rest}/\mu s]
\end{align}
where in the latter expression we substitute the usual values of $N_{it} \simeq 2\times 10^6$,  $n_{\rm chunk}\simeq
10^4$, and $N_{\rm eval}\simeq 20$ conservtively adopted
for problems involving many degrees of freedom $d$ (between 6 to 8).  We target $N_{\rm eval} \simeq 20$ so that startup and other one-time costs  are not a significant contributor to the overall
runtime.   In this configuration, the run time per marginalized likelihood is around $T_{ILE}/N_{\rm eval} \simeq 25
\unit{s}$ for a binary black hole.    Of this $25 \unit{s}$, only roughly
$1.5\unit{s}$ corresponds to evaluating the likelihood, implying the cost $\tau_{it} \simeq 1.5/N_{\rm it} \simeq
7.5\times 10^{-7}$.
Likewise, because overhead dominates over array manipulations, an analysis with  $16\unit{kHz}$ timeseries requires only
marginally more time than the corresponding $4\unit{kHz}$ analysis.
%
Conversely, the overhead of  performing the Monte Carlo (including both $\tau_{ad}$ and $\tau_{it,rest}$) is a quite substantial contribution to the overall runtime for the best-available hardware.

For low-latency analyses, we need to assess the overall wallclock time needed to perform an analysis, often with a
restricted number $N_{GPU}$ of available GPU-enabled machines.   To  complete $N_{\rm net}$ likelihood evaluations
with these resources requires
\begin{align}
T_{net,mod} &= \frac{N_{\rm net}}{N_{\rm GPU} N_{\rm eval}} T_{ILE,mod} \nonumber \\
& =
  \frac{N_{net}}{N_{GPU}}\left[
\frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
 \right. \nonumber \\
& \left. + 
 [ \tau_{setup}  + N_{it}\tau_{it}
 ] 
 \right]
\end{align}
Typically we target $N_{\rm eval}  \lesssim 3\times 10^4, N_{\rm GPU} \simeq 100$, or roughly $300\times$ the cost per
individual marginalized likelihood evaluation.  For the binary black hole configuration described above, marginalized
likelihood evaluations will complete in about $100\; \unit{minutes}$ on a cluster of (b), or more realistically 200
minutes on a cluster of (a).  This number can be reduced to
tens of minutes with a modestly larger GPU pool, or by  marginally more conservative convergence thresholds
 $n_{\rm   eff}$ or $N_{it}$.  
This discussion ignores the latency introduced by the  gaussian process interpolation stage at the end.  We will revisit
accelerated GP interpolation in subsequent work.

The \ILE{} likelihood  generates waveform dynamics and $h_{lm}(t)$ once per evaluation $\bm{\lambda}$.   Waveform generation can contribute significantly to the
time needed to evaluate  $\ln {\cal L}_{\rm marg}$ for extremely costly
waveform models which require $\tau_{wf} $ a significant fraction of $  \tau_{setup}$.   That said, the tests described
above used relatively costly waveform generation with $h_{lm}(t)$ allocation requiring $\tau_{wf} \simeq 1.2 \unit{s}$
at 4kHz and $1.9\unit{s}$ at 16 kHz.

The \ILE{} likelihood $\ln {\cal L}_{\rm margT}$ involves sums over modes, with the most expensive likelihood and setup operations (involving $Q$) growing linearly with
the number of modes used.  This increased cost is most apparent in Table \ref{tab:CostBreakdown} when comparing the
columns corresponding  $\tau_{it,like}$ for GPU (b) between the $\pm 2$ and $(\pm 2,\pm 1)$ rows.  The cost of
evaluating $\ln {\cal L}_{\rm margT}$   with 4 modes is roughly twice that of runs with two modes, as expected.

In a heterogeneous computing environment, code performance varies substantially with available resources, as seen by
contrasting corresponding results for (a,b,c) in Table \ref{tab:CostBreakdown}.      For
the slowest hardware,  the overall runtime will be only a factor few smaller than the corresponding CPU-only runtime.

\subsection{Binary neutron stars}
\begin{table*}
\begin{tabular}{lrr|ccccc|rr}
Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
% ./parse_report.sh 20190130-profile_bns_nogpu_16kHz_pcdev13.log | more
%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
%   (11261 - 10974)/20/100. = 0.14 sec for 16 kHz
CPU & 16384 & $\pm 2,\pm 1 $ & 35 & 26 & 0.14  & 680  & 25 & 590     \\ 
       & 4096 & $\pm 2, \pm 1$ & 35 & &  &   &&  25 &  \\ \hline
GPU (a) & 16384 & $\pm 2, \pm 1$   & 35 &  &&&  &    \\
       & 4096 & $ \pm 2, \pm 1 $        &  35  & &&&& 60--450  \\ \hline
% ./parse_report.sh 20190205-profile_bns_gpu_lowres_pcdev13_iteration3.log | more
%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
%   (226.489 - 17)/20/100  : 0.1 sec for 4kHz
%   (165 - 29.  )/20/100   : looks odd for 16 kHz
GPU (b) & 16384 & $\pm 2, \pm 1$   & 35 & 26 & 0.07 &  2.4 & &35 & 16 \\
       & 4096 & $ \pm 2, \pm 1 $        &  35  & 11.5 & 0.1 &  2.4  & &24 &  \\ \hline
% Runtime: Based on *production* data for the LHO result, with ADVANCED ITERATIONS to insure quality
%   demos/test_workflow_batch_bns_gpu_lowlatency_LHO_v0 : about 2.5 hours for 20 points
% Profiling: actual individual run,
%    - it, rest: based on mcsampler time - likelihood, divide by 20, divide by 10^4
GPU (c) & 16384 & $\pm 2, \pm 1$   & 35 & 71 &&20 & 38 &  105   \\
       & 4096 & $ \pm 2, \pm 1 $        &  35  & 28 && 12 && 60  \\ \hline
\end{tabular}
\caption{\label{tab:CostBreakdown:BNS}\textbf{Profiling performance: Binary neutron stars}: Evaluation costs for the
  marginalized likelihood on default
  hardware,  analyzing $T=8\unit{s}$ of data with a binary neutron stars
  $m_1=1.4 M_\odot,M_2=1.35 M_\odot$, with convergence threshold $n_{\rm eff} > 50$.
}
\end{table*}

Table \ref{tab:CostBreakdown:BNS} shows the corresponding performance breakdown for our fiducial binary neutron star,
which has signal SNR=32.  In this analysis, we have adopted the TaylorT4 model with $\ell=2$ and all $m=\pm 2,\pm 1$ modes as our
fiducial analysis template.   This source has significantly longer duration and higher amplitude, both of which contribute to slower convergence and longer runtime. 
%% Despite using more modes and longer signals, and despite an overall likelihood evaluation
%% cost $\tau_{it}$ consistent BBH analyses,  likelihood evaluations for BNS are often
%% \emph{as fast or faster} than the corresponding likelihood evaluations for BBH with otherwise comparable settings.

Low-mass compact binaries like binary neutron stars have much longer inspiral times from the same starting frequency of
$20\unit{Hz}$: roughly $160 \unit{s}$.   Nominally, the manipulation of a corresponding power-of-two duration data ($256\unit{s}$) involves
$32\times$ more costly Fourier transforms and time integrations than the BBH analysis described above.  
%  a = np.zeros(16384*256,dtype=np.complex64); a.nbytes/1024/1024
[An array of $256\unit{s}$ of data at $16\unit{kHz}$ corresponds to  32 Mb.]  Comparing $\tau_{setup}$ appearing in Tables
 \ref{tab:CostBreakdown} and  \ref{tab:CostBreakdown:BNS}, we indeed find our BNS analysis requires  a factor of order   $32\times$
 longer to set up all necessary inner products.  Unlike the BBH analysis, the inner product evaluation costs in
 $\tau_{setup}$ now dominates the
 overall evaluation time $T_{ILE}/N_{\rm eval}$.

Improved performance arises in part from re-using our adaptive integrator, which (after the first marginalized
likelihood evaluation) can exploit tight localization afforded by the many BNS cycles available in band and the high
amplitude of our fiducial BNS signal.   In some cases, the Monte Carlo integral converges to our target accuracy in of order $10\times $ fewer steps than for BBH.  
Additionally, particularly for first-stage RIFT grids, the improved performance arises from our choice of initial grid: many points fit poorly and identified as such well before the fiducial $2\times 10^6$ Monte Carlo iterations.  
% This latter benefit can occur more frequently as we move to higher dimensions.
Both sources of improved performance relative to the BBH analysis are independent of choice of hardware.

For the closed-form post-Newtonian approximation used in the study above, waveform costs are a modest contribution to
overall runtime, contributing only 4.1 sec to $\tau_{setup}$ at 16 kHz.
For other approximations not available in closed form, the waveform generation cost for binary neutron star inspirals
can be much more substantial, particularly as $f_{\rm min}$ decreases below the 20\unit{Hz} used in this study.

While GPU cards have finite memory, the modest size of our underlying data arrays and intemediate data products on-board ensures that we are unlikely to
saturate this bound, even for 16 kHz data, unless we investigate significantly lower starting frequencies or
employ vastly more angular modes $h_{lm}$.
%\editremark{some issue: memory usage is about 1Gb for 16 kHz,  close to $4Gb$ limit on commodity cards now.  Will be hard to
%  increase mode set substantially for BNS on the cheapest cards  }

\section{Performance and validation demonstrations with full pipeline}
\label{sec:end-to-end}

To better illustrate code performance in realistic settings, we also describe end-to-end analyses with
the original and modified code.     In a full analysis, many of the initially-proposed  evaluation points $\bm{\lambda}$ are
either not or marginally consistent with the data.  As a result, most marginalized likelihood evaluations in the first
iteration proceed extremely rapidly.  As the high-cost and low-cost evaluations are generally well-mixed between
different instances, the overall time to complete the full first iteration is typically much lower than subsequent
iterations.   With a well-positioned and sufficiently large initial grid,  few follow-on iterations are needed.  
%   Moreover, because the code progresses slowly only on candidate points which are most consistent with the data,

%* more variable, because many evaluations early on \emph{are not} consistent, so only a small fraction of the evaluations
%needed are consistent with the observations ... convergence can be faster as the signal amplitude decreases

\subsection{Binary black hole analysis}
\label{sec:sub:BBHFull}

\begin{figure*}
% BASED ON
%   CIT: test_workflow_batch_gpu_lowlatency
% COMMAND
% ../plot_last_iterations_with.sh --parameter m1 --parameter m2  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':'
% ../plot_last_iterations_with.sh --parameter mc --parameter delta_mc   --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':'
%
\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_m2.png}
\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_cum.png}
% python plot_mean_variance.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency_meanVar.dat  --output bbh_zerospin_lnL_meanVar
\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_meanVar.png}
% python plot_convergence.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency.dat
\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_converge.png}
\caption{\label{fig:BBH:MultiIterate}\textbf{Convergence of BBH analysis: Zero spin}: Results for marginal posterior distributions
  of our fiducial synthetic binary black hole.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
  value of the underlying marginalized likelihood evaluations.   
%The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
\emph{Left panel } Posterior distribution
  over  $\mc$ and
  $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
\emph{Bottom left}: Mean and variance of  \AddedResponse{the array $\ln{\cal L}_{\rm marg}(\bm{\lambda}_j)$  for
$j=1,2,\ldots N_{\rm eval}$ indexing all candidate sets of intrinsic parameters $\bm{\lambda}_j$ performed in that iteration},  showing that after the
first iteration the
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
\emph{Bottom right panel}: The estimated evidence $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
error budget, Monte Carlo error is not shown.
}
\end{figure*}

Modest-amplitude short-duration binary black holes empirically constitute the most frequent detection candidates
for current ground-based GW observatories \cite{LIGO-O2-Catalog}.    Because of their brevity and hence broad posterior,
the RIFT code's interpolation-based method converges  rapidly.  Combined with the low cost of each iteration,
these sources require an exceptionally low committment of resources and can be performed in extremely low latency, as
desired.
To demonstrate this, we use the GPU-accelerated code with $N_{\rm eval}=20$, analyzing data with $f_{\rm
  sample}=4096\unit{Hz}$.    We use 100 points in the first iteration, in a very coarse mass  grid (i.e., spacing
comparable to typical astrophysical mass scales), followed by 20 points in each subsequent iteration.  Figure \ref{fig:BBH:MultiIterate} shows our results.   The
final posterior is already well-explored with the initial gridpoints, and converged by the second iteration.  
Conversely, on average all iterations required roughly 45 seconds per $\bm{\lambda}$ to evaluate their grid on GPU (a) hardware.  This
configuration therefore converged within the 30 minutes needed for the first two iterations.    We can achieve  smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.
%
While this low-dimensional problem does not capture all  fitting and automation challenges associated with high-dimensional
 fully precessing binaries, it does capture the low cost and rapid response possible with RIFT.

\subsection{Binary neutron star analysis : Assuming  zero spin}
\label{sec:sub:BNSFull}

\begin{figure*}
% COMMAND
% plot_posterior_corner.py --posterior-file posterior_samples-4.dat --posterior-file posterior_samples-3.dat --posterior-file posterior_samples-2.dat --posterior-file posterior_samples-1.dat --parameter mc --parameter delta_mc --mc-range  '[1.19655,1.19682]' --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc_bns.xml.gz  --composite-file all.net --quantiles None --posterior-label 4 --posterior-label 3 --posterior-label 2 --posterior-label 1 --use-legend
\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_corner_mc_delta_mc.png}
\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_mc_cum.png}
% python plot_mean_variance.py --convergence-file 20190201-bns-zerospin-nobatch_meanVar.dat  --output bns_zerospin_lnL_meanVar
\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_lnL_meanVar.png}
% python plot_convergence.py --convergence-file 20190201-bns-zerospin-nobatch.dat --output bns_zerospin_lnL_converge
\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_lnL_converge.png}
\caption{\label{fig:BNS:MultiIterate}\textbf{Convergence of BNS analysis: Zero spin}: Results for marginal posterior distributions
  of our fiducial synthetic neutron star.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
  value of the underlying marginalized likelihood evaluations.  
% The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
\emph{Left panel } Posterior distribution
  over  $\mc$ and
  $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
first iteration the
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
error budget early on, Monte Carlo error is not shown.
}
\end{figure*}

%% \begin{figure*}
%% \caption{\label{fig:BNS:Comparison}\textbf{Comparison of GPU, CPU analysis}: \editremark{worthwhile}
%% }
%% \end{figure*}

Significant-amplitude  binary neutron star mergers are the most important scenario for rapid parameter inference, as low
latency can enable multimessenger followup  \cite{LIGO-O2-Catalog}.    
% 
In our second test we compare two workflows, one with the original embarrassingly-parallel RIFT code using $N_{\rm
  eval}=1$ and one with the
``batched'' GPU-accelerated code with $N_{\rm eval}=20$, analyzing data with $f_{\rm sample}=4096\unit{Hz}$.    
% Code configuration
In this example, we use $100$ initial points spread over the two mass dimensions $\mc_z,\delta$, adding $20$ evaluations per
iteration.   We
choose this simple low-dimension, small-size, and slowly-converging configuration to facilitate visualization of the grid, posterior, and convergence.     The initial coarse grid covers a region $\mc_z\in[1.1962,1.1970] M_\odot$ and $\delta \in [0,0.25]$,  insuring the posterior was smaller than our initial
coarse grid spacing.     The top  panels of Figure  \ref{fig:BNS:MultiIterate} show posterior distributions derived
from the first several RIFT iterations, while the bottom panels show convergence diagnostics.
%% Figure \ref{fig:BNS:Comparison} compares results from the GPU and CPU codes on \editremark{identical test points: the
%%   initial grid}.\footnote{This comparison of $\ln {\cal L}_{\rm marg}$ is limited by Monte Carlo integration error and
%%   is provided  as a pedagogically convenient sanity check.  When comparing 
%%   $\ln {\cal L}_{\rm margT}$ results on identical data and parameters, the GPU and CPU implementations agree to
%%   significantly higher precision.}  The two results agree.
Because our analysis uses a post-Newtonian model to interpret a signal generated with SEOBNRv4, the peak posterior
density is slightly  offset
 from the synthetic signal's parameters

This specific workflow configuration  reduces overall core usage by maximizing GPU use per iteration: after the first
iteration, only one GPU is active.   
Specifically,  with the updated RIFT workflow used here, with one instance analyzing each 20 evaluations,  we use five
core+GPU pairs in the
first iteration, followed by one core+GPU for remaining iterations.
By contrast, with the original CPU-only RIFT code analyzing each $\bm{\lambda}$ in parallel, this process requires roughly 20 core-minutes
per $\bm{\lambda}$, using 100 cores in the first iteration and 20 cores in each
subsequent evaluation.   
Note that  because   $N_{\rm eval}=20$ for the GPU is larger than the (hardware-and $f_{\rm sample}$ dependent)  speedup factor between the CPU and
GPU implementation, the overall wallclock time needed for a end-to-end
analysis analysis is larger for the GPU workflow.
We can achieve comparable or smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.

\subsection{Binary neutron star analysis : Assuming  nonprecessing spin}
\label{sec:sub:BNS_spin}

Binary neutron star models with more parameters like spin and tides require correspondingly larger numbers of points in the
  initial grid and per iteration.  Fortunately, the number of observationally significant and accessible dimensions is often
  substantially less than the  prior  dimensionality.  In practice,we use roughly $20\times$ more points per
  iteration for precessing massive BBH systems ($d=8$) or for spinning binary neutron stars with tides ($d=6$).    As a
  result, we can still achieve relatively rapid turnaround on a high-dimensional binary neutron star analysis even in
  resource-constrained environments.

As a concrete demonstration of a realistic modest-latency analysis using the GPU-accelerated code, we reanalyze our fiducial binary
neutron star signal, accounting for the possibility of nonzero (aligned) neutron star spins.  Our initial grid consists
of 5000 points, spread \AddedResponse{approximately} uniformly across a 4-dimensional
hypercube $\mc_z\in[1.1962,1.1970] M_\odot$, $\delta \in[0,0.25]$ and $\chi_{i,z}\in[-0.05,0.05]$.  Subsequent
iterations use 500 random draws from the estimated posterior samples produced from the previous iteration.  
Because  our fiducial post-Newtonian model (TaylorT4) as implemented in \texttt{lalsuite}\cite{lalsuite} does not include spin, we
employ both the \texttt{SEOBNRv4\_ROM} and \texttt{SpinTaylorT4} waveform approximations for nonprecessing binaries,
omitting higher-order modes.   For SEOBNRv4\_ROM, we use 48 GPU (c) enabled nodes; for SpinTaylorT4, we use 100 GPU (a) nodes.
%
Figure \ref{fig:BNS:Spin} shows our results.
%

Due to our self-imposed resource constraints on $N_{\rm eval}=20$ and the number of GPUs, the first iteration is
resource-limited and requires
of order $(5000/20/100)\simeq
2.5$ to $(5000/20/48)\simeq
 5$ times as long as the first iteration of the zero-spin BNS analysis described above.  Each marginalized likelihood
 evaluation in the first iteration (as well as subsequent iterations) requires roughly 60 seconds for both SpinTaylorT4 and SEOBNRv4ROM, on average.   Subsequent
marginalized likelihood iterations are not resource-limited and complete in roughly 35 minutes, depending on hardware.
As previously, we can achieve comparable or smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.

\begin{figure*}
% COMMAND
% plot_posterior_corner.py  --posterior-file posterior_samples-5.dat  --posterior-file posterior_samples-4.dat   --posterior-file posterior_samples-3.dat --posterior-file posterior_samples-2.dat --posterior-file posterior_samples-1.dat --parameter mc --parameter delta_mc  --parameter chi_eff --mc-range  '[1.19655,1.197]' --chi-max 0.025 --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc_bns.xml.gz  --composite-file all.net --quantiles None  --posterior-label 5 --posterior-label 4 --posterior-label 3 --posterior-label 2 --posterior-label 1 --use-legend --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle ':'  --posterior-color 'k' --posterior-color 'r' --posterior-color 'g' --posterior-color 'b' --posterior-color 'k' --lnL-cut 50
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_corner_mc_delta_xi.png}
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_mc_cum.png}
% python plot_mean_variance.py --convergence-file 20190203-bns-withspin-batch_gpu_lowlatency_noHM_v4ROM_meanVar.dat  --output bns_withspin_lnL_meanVar
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_meanVar.png}
% python plot_convergence.py --convergence-file 20190203-bns-withspin-batch_gpu_lowlatency_noHM_v4ROM.dat  --output bns_withspin_lnL_converge
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_converge.png}
\caption{\label{fig:BNS:Spin}\textbf{Demonstration of low-latency spinning analysis}:
Results for marginal posterior distributions
  of our fiducial synthetic neutron star, assuming the binary has nonprecessing spins.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
  value of the underlying marginalized likelihood evaluations.  Only evaluations with marginalized log-likelihoods within $30$ of the maximum are shown, to increase contrast. The solid curves show an analysis with SEOBNRv4\_ROM.
%  the dotted curves show an analysis with SpinTaylorT4.  \editremark{currenly only v4ROM shown} 
 \emph{Left panel } Posterior distribution
  over  $\mc$,
  $\delta=(m_1-m_2)/M$, and $\chi_{\rm eff}$.    \emph{Right panel}: Marginal 1d CDFs of $\delta$, showing convergence.
\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
first iteration the
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
error budget early on, Monte Carlo error is not shown.
}
\end{figure*}

\SkipForEarlyCirculation{
\subsection{Precessing binary black hole analysis}
\label{sec:sub:BBH_v3}

In this section, we reanalyze our fiducial (zero spin) binary black hole signal without any restrictions on spin
magnitude or alignment, using a precessing effective--one-body approximation (SEOBNRv3 \editremark{cite}).   
We start with an initial  \emph{three-dimensional} grid of 5000 points in $\mc \in [20,35] M_\odot$, $\eta
\in[0.2,0.25]$ and $\chi_{\rm eff} \in [-0.5,0.5]$.  We very conservatively perform 10 iterations, adding 500 points
candidate evaluation points each iteration, using a volumetric spin prior on both dimensionless spins.

We break our  pipeline into two stages.  In the first five
iterations, the likelihood is assumed to depend only on  $\mc,\eta,\chi_{\rm eff}$.   By construction, during these
stages our algorithm proposes random choices for the perpendicular spins $\chi_{i,\perp}$  and the remaining
antisymmetric degree of freedom for the two candidate spins.    As demonstrated in Figure \ref{fig:BBH:Spin}, well
before the first five iterations complete, we've thoroughly explored the $\mc,\eta,\chi_{\rm eff}$ distribution and
converged to a (biased) posterior.

In the second stage, we perform a gaussian process fit which includes  all binary parameters ($\mc,\eta,\chi_{\rm
  eff},\chi_-,\chi_{i,x},\chi_{i,y}$).  As demonstrated in Figure \ref{fig:BBH:Spin}, we quickly converge to a posterior
distribution.

Throughout this calculation, each individual marginalized likelihood required roughly 1 minute to evaluate.  With 100
available GPU+cores, this analysis could require as little as 1.7 hours to evaluate the marginalized likelihood for all
10 iterations.  In practice, by using $N_{\rm eval}=20$ and therefore restricting to 25 cores for the iterative stages, we insured each iteration after the first required no less than
20 minutes to complete.  As a result of this self-imposed restriction, the wallclock time necessary to complete all 10 iterations was several hours.
That said, in this particular case substantially fewer iterations were required to reach convergence.

\begin{figure*}
% COMMAND
% ../plot_last_iterations_with.sh --parameter m1 --parameter m2  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':'
% ../plot_last_iterations_with.sh --parameter mc --parameter delta_mc --parameter chi_eff  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':'
%  ../plot_last_iterations_with.sh --parameter a1z --parameter a2z --parameter chi1_perp --parameter chi2_perp  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-color 'g' --posterior-color 'magenta'

\includegraphics[width=0.45\textwidth]{figures/bns_withspin_mc_cum.png}
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_meanVar.png}
\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_converge.png}
\caption{\label{fig:BBH:Spin}\textbf{Demonstration precessing binary black hole analysis}:
Results for marginal posterior distributions
  of our fiducial binary black hole, assuming the binary could have precessing spins.  Solid contours show credible intervals; solid one-dimensional distributions
  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
  value of the underlying marginalized likelihood evaluations.  Only evaluations with marginalized log-likelihoods within $30$ of the maximum are shown, to increase contrast. The solid curves show an analysis with SEOBNRv3.
%  the dotted curves show an analysis with SpinTaylorT4.  \editremark{currenly only v4ROM shown} 
 \emph{Left panel } Posterior distribution
  over  $\mc$,
  $\delta=(m_1-m_2)/M$, and $\chi_{\rm eff}$.    \emph{Right panel}: Marginal 1d CDFs of $\delta$, showing convergence.
\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
first iteration the
candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
error budget early on, Monte Carlo error is not shown.
}
\end{figure*}
}

\SkipForEarlyCirculation{
\section{ Results (optional?)}

*   Rerun all O2 BBHs with SEOBNRv3?

* PP plot demo for zero-spin binaries, with profiling of full run?

* reminder: HM for BNS break degeneracy, useful for source classification.  Show an example.

\editremark{bonus mentions}: eccentricity

Bonus code: lnLcut tapering, possibly neff tapering

Puffball in DAG every few iterations, to insure stability around edges and not overcover the core

}

\section{Conclusions}
\label{sec:conclude}
We have demonstrated that the marginalized likelihood $\ln {\cal L}_{\rm marg}(\bm{\lambda})$ appearing in the RIFT/rapidPE parameter
inference calculation can be evaluated at fixed $\bm{\lambda}$ in tens of seconds on average for
both binary black holes and binary neutron stars.  This performance improvement could enable very low latency source
parameter inference for compact binaries, which can be of use for targeting multimessenger followup observations via sky
localization and  precise source characterization.     This prospect is particularly interesting because RIFT can
often achieve this performance using computationally costly models for binary merger with rich physics like higher modes
or eccentricity, as the waveform generation cost does not usually limit code performance.

\AddedResponse{In addition to producing results rapidly, RIFT results can be produced with a noticably smaller overall resource footprint than
  loosely similar LI analyses, even without tuning to optimize  RIFT pipeline settings.  As a concrete and non-optimized
  example,  for all five of the iterations of the spinning binary neutron star parameter inference with \texttt{SEOBNRv4\_ROM} from
  $20\unit{Hz}$ described in this work, our  RIFT analysis expended  roughly 14 core-days.  By contrast, a TaylorF2 analysis of GW170817
  with LI in MCMC mode starting from $23\unit{Hz}$ required 228 core-days.  % C02-1_TF2LS
  For precessing binary black holes using SEOBNRv3, the improvement is equally substantial: 10 core-days for a
  10-iteration investigation of a synthetic GW150914-like source with RIFT, versus 291 core-days for a LI analysis of
  GW170729.  
  We defer detailed relative benchmarking using comparably-converged parameter inference to future work.
}

The overall code performance and thus latency can be further decreased substantially, notably by converting the Monte
Carlo random number generation and inner products to GPU-based operations.  In such a configuration, the marginalized
likelihood code would perform almost all calculations (except waveform generation) on a GPU, with minimal communication off
the board.  This configuration should further reduce the average time needed to compute $\ln {\cal L}_{\rm
  marg}(\bm{\lambda})$ for both binary black holes (which are Monte Carlo limited) and binary neutron stars (which are
inner-product limited).   We anticipate a further factor of roughly 10 reduction in overall evaluation time can be
achieved soon.  At that level of performance, a single 8-GPU machine with contemporary hardware could perform parameter inference less than 10
minutes.    Since binary compact objects intersect our past light cone only once every roughly 15 minutes, accounting
for all past history, such a configuration would be able to address low-latency parameter inference for the duration of
2nd-generation ground-based observing.
% based on Neval ~ 10^4
Alternatively, if larger resource pools are available in low latency, both the original and now GPU-accelerated RIFT  can
perform extremely rapid parameter inference if large iterations are performed completely in parallel (i.e., $N_{\rm
  eval} \simeq 1$).
%

By allowing  models with higher-order modes to be used in low latency, our code can exploit the tighter constraints which higher modes can enable on  the properties of low-mass
binaries with suitable amplitudes and orientations.  These tighter constraints could
better  inform low-latency source classification and hence multimessenger followup observations of compact binary mergers.

Beyond low-latency multimessenger astronomy, rapid parameter inference enables new applications.  For example, every
parameter inference provide \emph{evidence} for a signal being present in the data; with rapid parameter inference, this
evidence could be used as (the last stage in a hierarchical pipeline for) a detection statistic
%\editremark{citations:old idea,see Thrane/Rory papers recently} 
\cite{2018PhRvX...8b1019S}.    
This approach can identify individual events and
even a population.
Alternatively and in many ways equivalently, one can identify a population of GW sources without assuming any one is
real, by  applying parameter inference to more candidate events and self-consistently separating foreground and
background
\cite{2015PhRvD..91b3005F,2019MNRAS.tmp..230G}.

When suitable surrogate models are available, the overall code performance and thus latency could be yet again further
reduced by eliminating the iteration and fitting stages entirely, performing one Monte Carlo at once
\cite{gwastro-PE-AlternativeArchitecturesROM}.  This approach exploits a linear representation of  $h_{lm}(t)$  via
basis functions, to enable rapid likelihood evaluation as a function of both extrinsic parameters and  $\bm{\lambda}$.
Though not necessarily or compactly available for all surrogate models, particularly for the small basis sizes necessary
to fit onboard GPUs, this approach could enable exceedingly low latency at small computational cost.  This alternative
architecture would be exceptionally well-suited to the alternative applications of low-latency PE described above.

The use of \texttt{cupy} enables our code to be highly portable across architectures and heterogeneous GPU environments,
while transitioning smoothly between GPU and CPU mode.     The techniques we used here will be transported to other
Bayesian inference modeling codes used to interpret GW observations \cite{gwastro-PopulationReconstruct-Parametric-Wysocki2018,gwastro-PopulationReconstruct-Code-PopModels}.
% Will swap version of code

\AddedResponse{In this work, we have focused exclusively on profiling a simplified pipeline  to produce posterior distributions
  for detector-frame intrinsic parameters.   The  code  also produces
reliable extrinsic parameter distributions \cite{gwastro-PE-AlternativeArchitectures}.  With some postprocessing,
this pipeline  provides joint posterior distributions for all intrinsic and extrinsic parameter distributions together;
examples of these distributions have been published elsewhere \cite{LIGO-O2-Catalog}.  
Presently, rather than harvest extrinsic information from every iteration, we harvest joint intrinsic and
extrinsic information with a single final iteration, which we will implement shortly in our production pipeline.
}

\begin{acknowledgements}
We thank our anonymous referee for helpful feedback.
DW thanks the RIT COS and CCRG for support, and  NSF-1707965.
JL is supported by NSF PHY-1707965.  
RO'S is supported by NSF PHY-1707965 and PHY-1607520.
Y-LLF was supported in part by BNL LDRD projects No.\,17-029 and No.\,19-002, 
and New York State Urban Development Corporation, d/b/a Empire State Development, under contract No.\,AA289.
We thank the Computational Science Initiative at BNL for hosting GPU Hackathon 2018, during which a part of this work was conducted. We also thank Jeff Layton at NVIDIA for bringing CuPy to our attention.
\end{acknowledgements}

\appendix
\section{Properties of resources used}
% https://ldas-gridmon.ligo.caltech.edu/condor/
%
% > CIT:
%% > 560 nodes with single GTX 1050 Ti
%% > 32 nodes with dual K10 PCI cards (shows up as 4 GPU devices) and dual GTX 750 Ti
%% > 16 nodes with dual GTX 750 Ti (and dual Xeon Phi co-processor)
%% > 
%% > LHO and LLO each:
%% > 40 nodes with GTX 1050 Ti
%% > single DGX-1 system with 8 tightly coupled V100

\AddedResponse{
LIGO-CIT worker nodes, denoted by (a)  in the text, are principally S6 nodes with 2-CPU $\times 8$ core Opteron
2.3 GHz machines with 16 Gb of RAM, with GTX 1050 Ti cards with 4 Gb of RAM.  For LIGO-CIT, profiling reports reflect
performance averaged over the whole cluster and hence lacks the detailed reporting produced for the other configurations.
% 
LIGO-LHO worker nodes with GPUs, denoted by (c) in the text, are a heterogeneous configuration mostly consisting of
(a).  Unlike profiling at LIGO-CIT, the profiling for LIGO-LHO reflects controlled tests on a single node.
% nvidia-smi
The V100 machine (ldas-pcdev13, denoted by (b) in the text) is  a 24-core ES-2650 v4 machine with 4 GPUs, only one of
which is active in our tests: Tesla V100 with 16 Gb of RAM.
All non-GPU profiling was  also performed on this machine.  
}

%\bibstyle{unsrt}
\bibliography{paperexport,LIGO-publications}
\end{document}

### 2023-05-07 run-10-20-1000.sh
run.sh : Adding time command in bash script

@@ -1,0 +1,1 @@
+time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_20_1000.hdf5 --n-walkers 20 --n-samples 1000

### 2023-05-07 run-10-50-5000.sh

@@ -1,1 +1,0 @@
-python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_50_5000.hdf5 --n-walkers 50 --n-samples 5000

### 2023-05-09 run-10-20-1000.sh
Adding time in Bash script and corner plot in Plots.ipynb

@@ -1,1 +1,1 @@
-time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_20_1000.hdf5 --n-walkers 20 --n-samples 1000
+time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_120_20000.hdf5 --n-walkers 120 --n-samples 20000

### 2023-05-09 paper/paper.tex
Method section update

@@
 
-\title{Population Inference of Eccentric Black Holes Binaries}
+\title{Population Inference of Eccentric Binary Black Holes}
 \author{M. Zeeshan}
@@
 %
-The properties of the sources responsible have been inferred via Bayesian comparison of  each source with candidate
-gravitational wave signals
-\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns,LIGO-O2-Catalog,gwastro-PE-AlternativeArchitectures,gwastro-PENR-RIFT,gw-astro-PE-lalinference-v1}.
-Many more are expected as observatories reach design sensitivity \cite{2016LRR....19....1A}.  
-Both to handle the rapid pace of discovery and particularly to enable coordinated multimessenger followup, GW
-observatories should be able to reconstruct the evidence for a signal being present in the data along with the full source parameters of coalescing binaries as fast as possible
-\cite{LIGO-2016-WhitePaper,LIGO-Program-2018}.
-However, particularly when using the best-available waveform models, these calculations can be very costly.  For binary
-neutron stars, detailed inferences even using simplified waveforms take weeks.  Even for massive binary  black holes,
-these calculations can take  months for costly time-domain effective-one-body models which incorporate the effects of precession.
-
-% BORROWED FROM ROM: edit
-Several strategies have been
-developed to reduce the 
-%overall computational cost and turnaround time of 
-computational cost of parameter estimation \cite{gwastro-pe-Brandon-STF2,gwastro-PE-AlternativeArchitectures,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,2016PhRvD..94d4031S,2017CQGra..34k5006V,2018arXiv180608792Z}.
-% The computational cost of parameter estimation can be reduced 
-Approaches that have appeared in the literature include
-%by 
-generating the approximate solutions more quickly \cite{gwastro-mergers-PE-ReducedOrder-2013,2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C,gwastro-mergers-IMRPhenomP,gwastro-SpinTaylorF2-2013};
-%by 
-interpolating some combination
-of the waveform or likelihood \cite{gwastro-approx-ROMNR-Blackman2015,2013PhRvD..87d4008C,2013PhRvD..87l2002S,2014CQGra..31s5010P,2014PhRvD..90d4074S,gwastro-PE-AlternativeArchitectures,cole2014likelihood,2012MNRAS.421..169G};  or
-%by 
-adopting a sparse representation to reduce the computational cost of data
-handling \cite{antil2013two,gwastro-mergers-PE-ReducedOrder-2013,2016PhRvD..94d4031S,gw-astro-ReducedOrderQuadraturePE-TiglioEtAl2014,gwastro-PE-AlternativeArchitectures,2010arXiv1007.4820C,2018arXiv180608792Z}.
-% 2014PhRvX...4c1006F,2013PhRvD..87l2002S,2013PhRvD..87d4008C
-\AddedResponse{Not published but also important are code optimization projects that improve    infrastructure,
-  such as better parallelization for \texttt{lalinference} \cite{gw-astro-PE-lalinference-v1,lalsuite}.}
- Some methods, however, achieve rapid turnaround through simplifying approximations.  
-%
-The RIFT/rapidPE strategy described in \cite{gwastro-PE-AlternativeArchitectures,2017CQGra..34n4002O,gwastro-PENR-RIFT,NRPaper} eschews these simplifications, performing
-embarrassingly-parallel inferences even for costly models.   However, the method as previously implemented still had a
-significant net  computational cost and noticeable startup time, limiting its diverse potential applications.
-%
-Previously, B. Miller  developed a custom-coded  implementation of the
-relevant likelihood on graphics processing units (GPUs), suggesting substantial performance improvements were possible
-\cite{Thesis-Brandon-2016,GPU-NU-work}.  
-\AddedResponse{The pycbc search code\cite{2016CQGra..33u5004U} and its pycbc-inference extension
-  \cite{2019PASP..131b4503B} can also make use of several GPU
-  hardware-accelerated operations.
-}
-In this paper,  we describe a variant of the RIFT approach which flexibly translates one of its
-algorithms to operate on GPUs, and adjusts its workflow to exploit the speed improvements
-this re-implementation affords.
-
-This paper is organized as follows.
-In Section \ref{sec:methods}, we review the underlying marginalized likelihood calculations used by RIFT and  their updated GPU
-implementation.  
-In Section \ref{sec:demo}, we quantify the improved performance of our  GPU-accelerated code,   while assessing
-operating settings which facilitate increased performance.
-In Section \ref{sec:end-to-end}, we describe the performance of the end-to-end pipeline on the synthetic nonspinning signals used in
-Section \ref{sec:demo}.  
-In Section \ref{sec:conclude}, we summarize our results and discuss their potential applications to future GW source and
-population inference.
 
@@
 
-I need to start writing from here.
+The Binary Black Holes (BBHs) can be described by three intrinsic and seven extrinsic parameters.
+The intrinsic parameters: such as mass ($m_i$), spin $( \chi_i)$, and eccentricity $(e)$, are subject to the orbital evolution of the binary.
+The extrinsic parameters: orientation (orbital phase, polarization, and inclination), sky location (right ascension, declination), luminosity distance, and coalescence time depend on the observer.
+
+{\color{red}Most of the eccentric binaries are driven by low mass events \cite{?}}.
+Therefore,  in this study, we  will consider modest masses $(1M_\odot-40M_\odot)$, non-spinning $(\chi_i = 0)$, and eccentric $(0-1)$ binaries.
+In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2$.
 
@@
 
-% \begin{equation}
-% \label{eq:lnL}
-% \ln {\cal L}(\bm{\lambda} ,\theta )=-\frac{1}{2}\sum\limits_{k}\langle h_{k}(\bm{\lambda} ,\theta )-d_{k} |h_{k}(\bm{\lambda} ,\theta )-d_{k}\rangle _{k}-\langle d_{k}|d_{k}\rangle _{k},
-% \end{equation}
-
-% For the remaining dimensions, to evaluate the likelihood in regions of high importance, we use an adaptive Monte Carlo as described in
-% \cite{gwastro-PE-AlternativeArchitectures}.
+\begin{align}
+\label{eq:plaw}
+p(m_1,m_2,e) = &C(\alpha,m_{min},m_{max},M_{max},e)  
+\nonumber \\ & \frac{m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
+\nonumber \\ &
+\sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
+\end{align}
 
@@
 
-\subsection{Tradeoff between Monte Carlo integration and accuracy}
+% \begin{table*}
+% \begin{tabular}{lrr|ccccc|rr}
+% Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
+% $\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
+%   &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
+% % ~/parse_report.sh profile_nogpu_pcdev13.log | more
+% CPU & 16384 & $\pm 2 $ & 20 & 2.4 &&540 & 20 &  690  \\ 
+%        & 4096 & $\pm 2 $ &   20  &&&& 20 \\ \hline
+% % ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
+% % /parse_report.sh ./profile_nogpu_HM_lowres_pcdev13.log 
+% %    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
+% %   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
+% %    
+% CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & 20 &  1060  \\ 
+%        & 4096 & $ \pm 2, \pm 1 $ &   20 &&&& 20  \\ \hline
 
-Each marginalized likelihood evaluation has a relative uncertainty of order  $1/\sqrt{n_{\rm eff}}$, where the number of
-effective samples $n_{\rm eff}=\epsilon N_{\rm it}$ increases linearly with the total number $N_{\rm it}$ of Monte Carlo
-evaluations performed by \ILE{}.  Therefore, to increase (decrease) our accuracy for each likelihood evaluation by a factor of $A$
-will require a factor $A^2$ more (fewer) iterations.   
-%
-We adopt a fixed threshold on $n_{\rm eff}$.
+%GPU (a) & 16384 & $\pm 2 $  & 20  & & && & 270 \\
+%            & 4096 &$\pm 2 $  &  20 &  & & & & 45 \\ \hline
 
-We do benefit slightly by re-using the adaptive Monte Carlo integrator for each extrinsic point $\bm{\lambda}$.  Since the
-integrator has already identified the likely range of sky locations and distances on the first iteration, each
-subsequent evaluation of the marginalized likelihood can converge marginally more quickly.
+%GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 20 &28 & 15\\
+%        & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 & 20  & 25\\ \hline
 
-\section{Analysis of marginalized likelihood evaluation cost}
-\label{sec:demo}
+% GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & 20  & 38  \\
+%        & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  & 20 & 35 & \\ \hline
 
-We illustrate the code using two synthetic signals: a binary black hole with masses $m_{1,2}=35,30 M_\odot$ and a binary
-neutron star with masses $1.4 M_\odot$ and $1.35 M_\odot$.
- We perform this analysis on heterogeneous LVC collaboration computing resources \AddedResponse{described in more detail
-   in the Appendix}, to assess our variable
-performance across architectures.  Notably, we investigate the following GPU options: (a) GTX 1050 Ti at LIGO-CIT,
-available in large numbers; (b) V100, available on selected high-performance machines; (c) and
-GTX 1050 at LIGO-LHO.    Unless otherwise noted in the text, we will discuss code configurations using CPU-only and GPU (b) using
-$4096\unit{Hz}$ sampling and only the  $\ell=2, m=\pm 2$ modes.  In this section, we  conservatively report computational cost for
-parameters $\bm{\lambda}$ which are close to or within the posterior.  Because of their consistency with the data, 
-the marginalized likelihood calculations converge the most slowly, as they have the best-determined extrinsic
-parameters.
+% GPU (c) & 16384 & $\pm 2 $  & 20  &6  & & 18&  58&160 &  \\
+%             & 4096 &$\pm 2 $  &  20 & 3.7 & & 11  & 58  & 140 & $\simeq 50$ \\
+% \end{tabular}
+% Compute node at LIGO-WA
+% \caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
+%   marginalized likelihood on default
+%   hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
+%   $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.
+% }
+%\end{table*}
 
-\subsection{Synthetic source generation and analysis settings}
+% \subsection{Binary black hole analysis}
+% \label{sec:sub:BBHFull}
 
-We generate  synthetic data for a two-detector LIGO configuration, assuming operation at  initial (BBH) or advanced (BNS) design  sensitivity.
-Both signals are generated with SEOBNRv4 effective-one-body waveform approximation \cite{2017PhRvD..95d4028B}, with a zero-noise data realization.   
-To qualitatively reproduce the noise power spectrum and amplitude of typical binary black hole observations in O1 and
-O2, the binary black hole signal has source distance $200 \unit{Mpc}$, chosen so SNR $\simeq 14$.  
-Similarly, to qualitatively reproduce the analysis of GW170817, our synthetic binary neutron star is placed $100
-\unit{Mpc}$ away, so the signal amplitude is SNR $\simeq 31$.
+% \begin{figure*}
 
- We perform a multi-stage iterative RIFT analysis of these signal with the
-time-domain SEOBNRv4 (BBH) or TaylorT4 \cite{gw-astro-PN-Comparison-AlessandraSathya2009}( BNS) approximation, under the simplifying assumption that both objects are point particles with zero
-spin.   For both sources, we adopt the fiducial distance prior $p(D_L) \propto D_L^2$,   selecting  a maximum distance
-roughly five times larger than the known source distance: 1Gpc for the binary black hole and 500 Mpc for the neutron
-star.  We adopt a uniform prior on the component (redshifted) masses $m_{i,z}$ (and, when appropriate, spins
-$\chi_{i,z}$).
-Unless otherwise noted, all mass quantities described in this work are redshifted masses $m_{k,z}=(1+z)m_k$.
-For the binary black hole we generate the signal with $f_{\rm min}=8 \unit{Hz}$ and analyze the signal with  $f_{\rm
-  min}=10 \unit{Hz}$; for the binary neutron star, we generate the signal with $f_{\rm min}=20\unit{Hz}$ and anayze it
-with $f_{\rm min}=23\unit{Hz}$.
-
-\subsection{Massive compact binary black holes}
-Previously, each instance of ILE examined one intrinsic point $\bm{\lambda}$.  The overall cost of this ILE evaluation
-involved two parts: a startup cost, a setup cost, and a Monte Carlo integration cost.  The  startup cost $\tau_{start}$
-is associated with code setup followed by reading, data conditioning, and Fourier-transforming the
-data.  The setup cost $\tau_{setup}$ arises from waveform generation and  the inner products $U,V,Q$.  Finally, the Monte Carlo
-cost $\tau_{mc}=N_{ad} \tau_{ad}+N_{it}\tau_{eval}$ increases with the number of Monte Carlo iterations $N_{it}$ with
-and the number of times $N_{ad}$ the sampling prior used in adaptive Monte Carlo is regenerated from the most recent 
-$n_{\rm chunk}$ data points, in proportion to their
-cost.   In the standard configuration, the adaptive sampling prior is regenerated every $n_{\rm chunk}$ iterations (i.e,
- $N_{ad} \simeq N_{it}/n_{chunk}$), regenerating a sampling prior in two sky location parameters and distance.  
-The choice of one $\bm{\lambda}$ for each instance of ILE was due to the substantial time  $\tau_{MC}$, which vastly
-dominated the overall computational cost.   For example, for a typical analysis of a short signal --  a typical binary
-black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$ --  this version has
-cost elements as shown in Table \ref{tab:CostBreakdown}, based on the assumption that the MC terminates after $N_{\rm it}\simeq 2\times 10^6$ iterations using
-$n_{\rm chunk}\simeq 10^4$.  
-The marginalized likelihood described above  converges incredibly rapidly to a small value if the candidate model is
-inconsistent with a signal (or the absence thereof) in the data, with only $N_{it} \simeq O(10^4)$  evaluations needed.
-
-\begin{table*}
-\begin{tabular}{lrr|ccccc|rr}
-Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
-$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
-  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
-% ~/parse_report.sh profile_nogpu_pcdev13.log | more
-CPU & 16384 & $\pm 2 $ & 20 & 2.4 &&540 & 20 &  690  \\ 
-       & 4096 & $\pm 2 $ &   20  &&&& 20 \\ \hline
-% ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
-% /parse_report.sh ./profile_nogpu_HM_lowres_pcdev13.log 
-%    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
-%   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
-%    
-CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & 20 &  1060  \\ 
-       & 4096 & $ \pm 2, \pm 1 $ &   20 &&&& 20  \\ \hline
-% LIGO-WA, or a compute node at CIT with a grpahics catd: 
-%  - old compute node: one marginalized likelihood in 2 minutes (!), 2x8
-%  - CIT jobs, on compute nodes
-%        20 points at 16kHz take about 90 min
-%        20 points at kHz take anywhere from
-%        condor_userlog iteration_0_ile/logs/*.log
-GPU (a) & 16384 & $\pm 2 $  & 20  & & && & 270 \\
-            & 4096 &$\pm 2 $  &  20 &  & & & & 45 \\ \hline
-%% % LIGo CIT, pcdev13, 16 kHz : basically indistinguishable from cost at 4kHz
-%%       1    0.027    0.027  560.572  560.572 integrate_likelihood_extrinsic_batchmode:8(<module>)
-%%        20    0.082    0.004  552.004   27.600 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
-%%        20   76.306    3.815  503.342   25.167 mcsampler.py:327(integrate)
-%%      2345   57.019    0.024  241.951    0.103 mcsampler.py:198(draw)
-%%     25785    0.135    0.000   99.535    0.004 shape_base.py:286(hstack)
-%%     26561   99.102    0.004   99.102    0.004 {numpy.concatenate}
-%%     21981    0.933    0.000   87.819    0.004 function_base.py:2063(__call__)
-%%     21981   34.750    0.002   86.887    0.004 function_base.py:2154(_vectorize_call)
-%%      2345    1.391    0.001   86.327    0.037 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
-%%      2345    0.007    0.000   46.648    0.020 __init__.py:630(asnumpy)
-%%      2345   46.557    0.020   46.640    0.020 {method 'get' of 'cupy.core.core.ndarray' objects}
-%%        20    0.179    0.009   43.145    2.157 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
-%%        20    0.174    0.009   37.498    1.875 lalsimutils.py:2582(hlmoft_SEOB_dict)
-%%        20    0.099    0.005   37.324    1.866 lalsimutils.py:2743(complex_hoft)
-%%        20   37.220    1.861   37.220    1.861 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
-%%    444365    0.761    0.000   35.303    0.000 polyint.py:62(__call__)
-%%      2345    5.260    0.002   33.956    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
-%%       407    0.005    0.000   28.757    0.071 odepack.py:28(odeint)
-%%       407    1.327    0.003   28.740    0.071 {scipy.integrate._odepack.odeint}
-%%    444365    2.644    0.000   26.423    0.000 interpolate.py:657(_evaluate)
-%% 117270874   22.569    0.000   22.569    0.000 mcsampler.py:685(uniform_samp)
-% LIGO-CIT, pcdev13, 4kHz
-%  - 20 iterations, about 450  to 500 sec to eval (i.e., about 25 sec per lnL)
-%  -  precompute: can time manually
-%   -  Overlap stages: PrecomputeLIkelihood + waveform, about 4
-%  - Histogram time: about (12+30) seconds overall (histogram + ODEint)
-%  - Actual likelihood evaluation: about 30 sec out of 20 eval, or about 1.5 sec
-   %%   1    0.061    0.061  506.265  506.265 integrate_likelihood_extrinsic_batchmode:8(<module>)
-   %%     20    0.194    0.010  498.581   24.929 integrate_likelihood_extrinsic_batchmode:690(analyze_event)
-   %%     20   80.397    4.020  464.263   23.213 mcsampler.py:327(integrate)
-   %%   2140   54.358    0.025  235.480    0.110 mcsampler.py:198(draw)
-   %%  23530    0.134    0.000  101.280    0.004 shape_base.py:286(hstack)
-   %%  24306  100.861    0.004  100.861    0.004 {numpy.concatenate}
-   %%  20136    0.896    0.000   84.698    0.004 function_base.py:2063(__call__)
-   %%  20136   33.608    0.002   83.802    0.004 function_base.py:2154(_vectorize_call)
-   %%   2140    1.403    0.001   45.791    0.021 integrate_likelihood_extrinsic_batchmode:810(likelihood_function)
-   %% 461726    0.790    0.000   36.137    0.000 polyint.py:62(__call__)
-   %%   2140    5.496    0.003   30.829    0.014 factored_likelihood.py:1676(DiscreteFactoredLogLikelihoodViaArrayVectorNoLoop)
-   %%    407    0.006    0.000   29.662    0.073 odepack.py:28(odeint)
-   %%    407    1.360    0.003   29.644    0.073 {scipy.integrate._odepack.odeint}
-   %%     20    0.052    0.003   28.785    1.439 factored_likelihood.py:92(PrecomputeLikelihoodTerms)
-   %% 461726    2.809    0.000   26.983    0.000 interpolate.py:657(_evaluate)
-   %%     20    0.048    0.002   24.454    1.223 lalsimutils.py:2582(hlmoft_SEOB_dict)
-   %%     20    0.030    0.002   24.406    1.220 lalsimutils.py:2743(complex_hoft)
-   %%     20   24.373    1.219   24.373    1.219 {lalsimulation._lalsimulation.SimInspiralChooseTDWaveform}
-% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_welford | more
-% ./parse_report.sh 20190130-profile_gpu_lowres_pcdev13.log_16kHz | more
-%   'residual' time: take mcsampler, subtract likleihood, divide by number of calls and number of evals per chunk
-GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 20 &28 & 15\\
-       & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 & 20  & 25\\ \hline
-% ./parse_report.sh 20190130-profile_gpu_HM_16kHz_pcdev13.log | more
-% ./parse_report.sh 20190130-profile_gpu_HM_lowres_pcdev13.log  | more
-% ~/parse_report.sh profile_gpu_HM_lowres_pcdev13.log | more
-%   - extra eval cost: (mcsampler - likelihood)/20/n_eval_likelihood/10^4  
-GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & 20  & 38  \\
-       & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  & 20 & 35 & \\ \hline
-%
-% LHO or LLO worker nodes
-%  ./parse_report.sh 20190130-profile_gpu_LHOworker_16kHz.log | more
-% ./parse_report.sh 20190130-profile_gpu_LHOworker_lowres.log | more
-%   * worker iteration cose: (3027 - 687)/20/3851/10^4 * 10^6 // N
-GPU (c) & 16384 & $\pm 2 $  & 20  &6  & & 18&  58&160 &  \\
-            & 4096 &$\pm 2 $  &  20 & 3.7 & & 11  & 58  & 140 & $\simeq 50$ \\
-\end{tabular}
-% Compute node at LIGO-WA
-\caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
-  marginalized likelihood on default
-  hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
-  $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.
-%\editremark{Comment: will update last number based on bulk profiling of a whole run; other numbers are from deep dive into individual run}
-}
-\end{table*}
-
-With the new low-cost ${\cal L}_{\rm margT}$ evaluations, however, the startup and setup costs  $\tau_{start},\tau_{setup}$ now can be
-comparable or in excess of the total time used to evaluate the marginalized likelihood ${\cal L}_{marg}$.   In fact, the total time per Monte Carlo
-evaluation $\tau_{it}=\tau_{manage}+\tau_{eval}$  spent in general-purpose overhead $\tau_{manage}$ will be much larger than the time $\tau_{eval}$ spent
-carrying out scientific calculations by evaluating the likelihood.  For these reasons, we reorganize the workflow, so each
-instance of ILE loops over
-$N_{eval}$ different choices for $\bm{\lambda}$. 
-%
-Additionally, particularly for massive binary black holes, we investigate two additional performance improvements.  First and
-foremost, we lower the sampling rate, thus lowering the startup cost $\tau_{start}$ and particularly setup cost
-$\tau_{setup}$.  Previously and out of an abundance of caution, \ILE{} employed a sampling rate $1/\Delta t = 16384
-\unit{Hz}$ for all calculations, but terminated all inner products in Eq. (\ref{eq:overlap}) at roughly $f_{\rm
-  max}=2048\unit{Hz}$ or less.   Reducing the sampling rate by a factor $s$ will \AddedResponse{reduce} the  cost of all operations with
-timeseries -- they are shorter.  Depending on the relative cost of overhead versus array operations,  $\tau_{setup}$
-\emph{and} the cost per evaluation $\tau_{eval}$ decrease modestly because of this effect.  
-Also following Pankow et al, to insure safe inner products over short data sets in the presence of very narrow spectral lines, we operated on a
-significantly-padded data buffer and modified the noise power spectrum by truncating the inverse power spectrum to a
-finite response time.  For binary black holes, the degree of padding was significantly
-in excess of the signal duration.  
-Instead and following lalinference \cite{gw-astro-PE-lalinference-v1}, we adopt a much shorter inverse spectrum truncation length, a much shorter padding buffer, and
-a Tukey window applied to the data to remove discontinuities at the start and end of each segment.   
-Reducing the duration of data analyzed by a factor $s'$ will reduce the cost $\tau_{setup}$ by a factor $s'$.  
-%
-Combining these factors together, the overall computational cost  $T_{ILE}/N_{eval}$  of each marginalized likelihood
-evaluation is
-\begin{align}
-T_{ILE}/N_{\rm eval} &= \frac{\tau_{start}}{N_{\rm eval}} 
- + 
- [ \tau_{setup} + N_{ad} \tau_{ad} + N_{it}\tau_{it}
- ] 
-\end{align}
-The first rows in table Table  \ref{tab:CostBreakdown} shows our estimated breakdown of these elements of the computational cost, for a typical binary black hole signal with $f_{\rm min}=20 \unit{Hz}$ and $m_1\simeq m_2\simeq 30 M_\odot$.
-Because of the extremely high cost of adaptive overhead, we only adapt in two parameters, corresponding to sky location; 
-and only adapt until a marginalied likelihood evaluation  returns significant $\ln {\cal L}_{\rm marg}$.  As a result, $N_{ad}$
-(the number of adaptive stages \emph{per likelihood evaluation}) scales as $N_{it}/n_{chunk}/N_{\rm eval}$ and does not
-increase with the overall number of samples, so the overall evaluation time scales as
-\begin{align}
-T_{ILE,mod}/N_{\rm eval} &= \frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
- \nonumber \\ &
- + 
- [ \tau_{setup}  + N_{it}\tau_{it}
- ]  \\
-% (20 \[Tau]s1 + \[Tau]ad 1 2*10^6/10^4)/   20 + (\[Tau]start*1 + 2*10^6 \[Tau]it1 *10^(-6)) // Expand
-& \simeq (\tau_{\rm start}/20\unit{s}) +  10 (\tau_{ad}/\mu s) 
- \nonumber \\ &
-+ 2  [(\tau_{it,liike}/\mu s) + \tau_{it,rest}/\mu s]
-\end{align}
-where in the latter expression we substitute the usual values of $N_{it} \simeq 2\times 10^6$,  $n_{\rm chunk}\simeq
-10^4$, and $N_{\rm eval}\simeq 20$ conservtively adopted
-for problems involving many degrees of freedom $d$ (between 6 to 8).  We target $N_{\rm eval} \simeq 20$ so that startup and other one-time costs  are not a significant contributor to the overall
-runtime.   In this configuration, the run time per marginalized likelihood is around $T_{ILE}/N_{\rm eval} \simeq 25
-\unit{s}$ for a binary black hole.    Of this $25 \unit{s}$, only roughly
-$1.5\unit{s}$ corresponds to evaluating the likelihood, implying the cost $\tau_{it} \simeq 1.5/N_{\rm it} \simeq
-7.5\times 10^{-7}$.
-Likewise, because overhead dominates over array manipulations, an analysis with  $16\unit{kHz}$ timeseries requires only
-marginally more time than the corresponding $4\unit{kHz}$ analysis.
-%
-Conversely, the overhead of  performing the Monte Carlo (including both $\tau_{ad}$ and $\tau_{it,rest}$) is a quite substantial contribution to the overall runtime for the best-available hardware.
-
-For low-latency analyses, we need to assess the overall wallclock time needed to perform an analysis, often with a
-restricted number $N_{GPU}$ of available GPU-enabled machines.   To  complete $N_{\rm net}$ likelihood evaluations
-with these resources requires
-\begin{align}
-T_{net,mod} &= \frac{N_{\rm net}}{N_{\rm GPU} N_{\rm eval}} T_{ILE,mod} \nonumber \\
-& =
-  \frac{N_{net}}{N_{GPU}}\left[
-\frac{\tau_{start} + \tau_{ad}N_{it}/n_{\rm chunk}}{N_{\rm eval}} 
- \right. \nonumber \\
-& \left. + 
- [ \tau_{setup}  + N_{it}\tau_{it}
- ] 
- \right]
-\end{align}
-Typically we target $N_{\rm eval}  \lesssim 3\times 10^4, N_{\rm GPU} \simeq 100$, or roughly $300\times$ the cost per
-individual marginalized likelihood evaluation.  For the binary black hole configuration described above, marginalized
-likelihood evaluations will complete in about $100\; \unit{minutes}$ on a cluster of (b), or more realistically 200
-minutes on a cluster of (a).  This number can be reduced to
-tens of minutes with a modestly larger GPU pool, or by  marginally more conservative convergence thresholds
- $n_{\rm   eff}$ or $N_{it}$.  
-This discussion ignores the latency introduced by the  gaussian process interpolation stage at the end.  We will revisit
-accelerated GP interpolation in subsequent work.
-
-The \ILE{} likelihood  generates waveform dynamics and $h_{lm}(t)$ once per evaluation $\bm{\lambda}$.   Waveform generation can contribute significantly to the
-time needed to evaluate  $\ln {\cal L}_{\rm marg}$ for extremely costly
-waveform models which require $\tau_{wf} $ a significant fraction of $  \tau_{setup}$.   That said, the tests described
-above used relatively costly waveform generation with $h_{lm}(t)$ allocation requiring $\tau_{wf} \simeq 1.2 \unit{s}$
-at 4kHz and $1.9\unit{s}$ at 16 kHz.
-
-The \ILE{} likelihood $\ln {\cal L}_{\rm margT}$ involves sums over modes, with the most expensive likelihood and setup operations (involving $Q$) growing linearly with
-the number of modes used.  This increased cost is most apparent in Table \ref{tab:CostBreakdown} when comparing the
-columns corresponding  $\tau_{it,like}$ for GPU (b) between the $\pm 2$ and $(\pm 2,\pm 1)$ rows.  The cost of
-evaluating $\ln {\cal L}_{\rm margT}$   with 4 modes is roughly twice that of runs with two modes, as expected.
-
-In a heterogeneous computing environment, code performance varies substantially with available resources, as seen by
-contrasting corresponding results for (a,b,c) in Table \ref{tab:CostBreakdown}.      For
-the slowest hardware,  the overall runtime will be only a factor few smaller than the corresponding CPU-only runtime.
-
-\subsection{Binary neutron stars}
-\begin{table*}
-\begin{tabular}{lrr|ccccc|rr}
-Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
-$\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
-  &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
-% ./parse_report.sh 20190130-profile_bns_nogpu_16kHz_pcdev13.log | more
-%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
-%   (11261 - 10974)/20/100. = 0.14 sec for 16 kHz
-CPU & 16384 & $\pm 2,\pm 1 $ & 35 & 26 & 0.14  & 680  & 25 & 590     \\ 
-       & 4096 & $\pm 2, \pm 1$ & 35 & &  &   &&  25 &  \\ \hline
-GPU (a) & 16384 & $\pm 2, \pm 1$   & 35 &  &&&  &    \\
-       & 4096 & $ \pm 2, \pm 1 $        &  35  & &&&& 60--450  \\ \hline
-% ./parse_report.sh 20190205-profile_bns_gpu_lowres_pcdev13_iteration3.log | more
-%   - resampling time: take (mcsampler - likelihood), divide by 20=Neval, divide by number of adaptive steps (100)
-%   (226.489 - 17)/20/100  : 0.1 sec for 4kHz
-%   (165 - 29.  )/20/100   : looks odd for 16 kHz
-GPU (b) & 16384 & $\pm 2, \pm 1$   & 35 & 26 & 0.07 &  2.4 & &35 & 16 \\
-       & 4096 & $ \pm 2, \pm 1 $        &  35  & 11.5 & 0.1 &  2.4  & &24 &  \\ \hline
-% Runtime: Based on *production* data for the LHO result, with ADVANCED ITERATIONS to insure quality
-%   demos/test_workflow_batch_bns_gpu_lowlatency_LHO_v0 : about 2.5 hours for 20 points
-% Profiling: actual individual run,
-%    - it, rest: based on mcsampler time - likelihood, divide by 20, divide by 10^4
-GPU (c) & 16384 & $\pm 2, \pm 1$   & 35 & 71 &&20 & 38 &  105   \\
-       & 4096 & $ \pm 2, \pm 1 $        &  35  & 28 && 12 && 60  \\ \hline
-\end{tabular}
-\caption{\label{tab:CostBreakdown:BNS}\textbf{Profiling performance: Binary neutron stars}: Evaluation costs for the
-  marginalized likelihood on default
-  hardware,  analyzing $T=8\unit{s}$ of data with a binary neutron stars
-  $m_1=1.4 M_\odot,M_2=1.35 M_\odot$, with convergence threshold $n_{\rm eff} > 50$.
-}
-\end{table*}
-
-Table \ref{tab:CostBreakdown:BNS} shows the corresponding performance breakdown for our fiducial binary neutron star,
-which has signal SNR=32.  In this analysis, we have adopted the TaylorT4 model with $\ell=2$ and all $m=\pm 2,\pm 1$ modes as our
-fiducial analysis template.   This source has significantly longer duration and higher amplitude, both of which contribute to slower convergence and longer runtime. 
-%% Despite using more modes and longer signals, and despite an overall likelihood evaluation
-%% cost $\tau_{it}$ consistent BBH analyses,  likelihood evaluations for BNS are often
-%% \emph{as fast or faster} than the corresponding likelihood evaluations for BBH with otherwise comparable settings.
-
-Low-mass compact binaries like binary neutron stars have much longer inspiral times from the same starting frequency of
-$20\unit{Hz}$: roughly $160 \unit{s}$.   Nominally, the manipulation of a corresponding power-of-two duration data ($256\unit{s}$) involves
-$32\times$ more costly Fourier transforms and time integrations than the BBH analysis described above.  
-%  a = np.zeros(16384*256,dtype=np.complex64); a.nbytes/1024/1024
-[An array of $256\unit{s}$ of data at $16\unit{kHz}$ corresponds to  32 Mb.]  Comparing $\tau_{setup}$ appearing in Tables
- \ref{tab:CostBreakdown} and  \ref{tab:CostBreakdown:BNS}, we indeed find our BNS analysis requires  a factor of order   $32\times$
- longer to set up all necessary inner products.  Unlike the BBH analysis, the inner product evaluation costs in
- $\tau_{setup}$ now dominates the
- overall evaluation time $T_{ILE}/N_{\rm eval}$.
-
-Improved performance arises in part from re-using our adaptive integrator, which (after the first marginalized
-likelihood evaluation) can exploit tight localization afforded by the many BNS cycles available in band and the high
-amplitude of our fiducial BNS signal.   In some cases, the Monte Carlo integral converges to our target accuracy in of order $10\times $ fewer steps than for BBH.  
-Additionally, particularly for first-stage RIFT grids, the improved performance arises from our choice of initial grid: many points fit poorly and identified as such well before the fiducial $2\times 10^6$ Monte Carlo iterations.
-% This latter benefit can occur more frequently as we move to higher dimensions.
-Both sources of improved performance relative to the BBH analysis are independent of choice of hardware.
-
-For the closed-form post-Newtonian approximation used in the study above, waveform costs are a modest contribution to
-overall runtime, contributing only 4.1 sec to $\tau_{setup}$ at 16 kHz.
-For other approximations not available in closed form, the waveform generation cost for binary neutron star inspirals
-can be much more substantial, particularly as $f_{\rm min}$ decreases below the 20\unit{Hz} used in this study.
-
-While GPU cards have finite memory, the modest size of our underlying data arrays and intemediate data products on-board ensures that we are unlikely to
-saturate this bound, even for 16 kHz data, unless we investigate significantly lower starting frequencies or
-employ vastly more angular modes $h_{lm}$.
-%\editremark{some issue: memory usage is about 1Gb for 16 kHz,  close to $4Gb$ limit on commodity cards now.  Will be hard to
-%  increase mode set substantially for BNS on the cheapest cards  }
-
-\section{Performance and validation demonstrations with full pipeline}
-\label{sec:end-to-end}
-
-To better illustrate code performance in realistic settings, we also describe end-to-end analyses with
-the original and modified code.     In a full analysis, many of the initially-proposed  evaluation points $\bm{\lambda}$ are
-either not or marginally consistent with the data.  As a result, most marginalized likelihood evaluations in the first
-iteration proceed extremely rapidly.  As the high-cost and low-cost evaluations are generally well-mixed between
-different instances, the overall time to complete the full first iteration is typically much lower than subsequent
-iterations.   With a well-positioned and sufficiently large initial grid,  few follow-on iterations are needed.  
-%   Moreover, because the code progresses slowly only on candidate points which are most consistent with the data,
-
-%* more variable, because many evaluations early on \emph{are not} consistent, so only a small fraction of the evaluations
-%needed are consistent with the observations ... convergence can be faster as the signal amplitude decreases
-
-\subsection{Binary black hole analysis}
-\label{sec:sub:BBHFull}
-
-\begin{figure*}
-% BASED ON
-%   CIT: test_workflow_batch_gpu_lowlatency
-% COMMAND
-% ../plot_last_iterations_with.sh --parameter m1 --parameter m2  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':'
-% ../plot_last_iterations_with.sh --parameter mc --parameter delta_mc   --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':'
-%
-\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_m2.png}
-\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_cum.png}
-% python plot_mean_variance.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency_meanVar.dat  --output bbh_zerospin_lnL_meanVar
-\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_meanVar.png}
-% python plot_convergence.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency.dat
-\includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_converge.png}
-\caption{\label{fig:BBH:MultiIterate}\textbf{Convergence of BBH analysis: Zero spin}: Results for marginal posterior distributions
-  of our fiducial synthetic binary black hole.  Solid contours show credible intervals; solid one-dimensional distributions
-  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
-  value of the underlying marginalized likelihood evaluations.   
-%The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
-\emph{Left panel } Posterior distribution
-  over  $\mc$ and
-  $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
-\emph{Bottom left}: Mean and variance of  \AddedResponse{the array $\ln{\cal L}_{\rm marg}(\bm{\lambda}_j)$  for
-$j=1,2,\ldots N_{\rm eval}$ indexing all candidate sets of intrinsic parameters $\bm{\lambda}_j$ performed in that iteration},  showing that after the
-first iteration the
-candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
-\emph{Bottom right panel}: The estimated evidence $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
-error budget, Monte Carlo error is not shown.
-}
-\end{figure*}
-
-Modest-amplitude short-duration binary black holes empirically constitute the most frequent detection candidates
-for current ground-based GW observatories \cite{LIGO-O2-Catalog}.    Because of their brevity and hence broad posterior,
-the RIFT code's interpolation-based method converges  rapidly.  Combined with the low cost of each iteration,
-these sources require an exceptionally low committment of resources and can be performed in extremely low latency, as
-desired.
-To demonstrate this, we use the GPU-accelerated code with $N_{\rm eval}=20$, analyzing data with $f_{\rm
-  sample}=4096\unit{Hz}$.    We use 100 points in the first iteration, in a very coarse mass  grid (i.e., spacing
-comparable to typical astrophysical mass scales), followed by 20 points in each subsequent iteration.  Figure \ref{fig:BBH:MultiIterate} shows our results.   The
-final posterior is already well-explored with the initial gridpoints, and converged by the second iteration.  
-Conversely, on average all iterations required roughly 45 seconds per $\bm{\lambda}$ to evaluate their grid on GPU (a) hardware.  This
-configuration therefore converged within the 30 minutes needed for the first two iterations.
-We can achieve  smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.
-%
-While this low-dimensional problem does not capture all  fitting and automation challenges associated with high-dimensional
- fully precessing binaries, it does capture the low cost and rapid response possible with RIFT.
-
-\subsection{Binary neutron star analysis : Assuming  zero spin}
-\label{sec:sub:BNSFull}
-
-\begin{figure*}
-% COMMAND
-% plot_posterior_corner.py --posterior-file posterior_samples-4.dat --posterior-file posterior_samples-3.dat --posterior-file posterior_samples-2.dat --posterior-file posterior_samples-1.dat --parameter mc --parameter delta_mc --mc-range  '[1.19655,1.19682]' --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc_bns.xml.gz  --composite-file all.net --quantiles None --posterior-label 4 --posterior-label 3 --posterior-label 2 --posterior-label 1 --use-legend
-\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_corner_mc_delta_mc.png}
-\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_mc_cum.png}
-% python plot_mean_variance.py --convergence-file 20190201-bns-zerospin-nobatch_meanVar.dat  --output bns_zerospin_lnL_meanVar
-\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_lnL_meanVar.png}
-% python plot_convergence.py --convergence-file 20190201-bns-zerospin-nobatch.dat --output bns_zerospin_lnL_converge
-\includegraphics[width=0.45\textwidth]{figures/bns_zerospin_lnL_converge.png}
-\caption{\label{fig:BNS:MultiIterate}\textbf{Convergence of BNS analysis: Zero spin}: Results for marginal posterior distributions
-  of our fiducial synthetic neutron star.  Solid contours show credible intervals; solid one-dimensional distributions
-  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
-  value of the underlying marginalized likelihood evaluations.  
-% The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
-\emph{Left panel } Posterior distribution
-  over  $\mc$ and
-  $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
-\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
-first iteration the
-candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
-\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
-error budget early on, Monte Carlo error is not shown.
-}
-\end{figure*}
-
-%% \begin{figure*}
-%% \caption{\label{fig:BNS:Comparison}\textbf{Comparison of GPU, CPU analysis}: \editremark{worthwhile}
-%% }
-%% \end{figure*}
-
-Significant-amplitude  binary neutron star mergers are the most important scenario for rapid parameter inference, as low
-latency can enable multimessenger followup  \cite{LIGO-O2-Catalog}.    
-% 
-In our second test we compare two workflows, one with the original embarrassingly-parallel RIFT code using $N_{\rm
-  eval}=1$ and one with the
-``batched'' GPU-accelerated code with $N_{\rm eval}=20$, analyzing data with $f_{\rm sample}=4096\unit{Hz}$.    
-% Code configuration
-In this example, we use $100$ initial points spread over the two mass dimensions $\mc_z,\delta$, adding $20$ evaluations per
-iteration.   We
-choose this simple low-dimension, small-size, and slowly-converging configuration to facilitate visualization of the grid, posterior, and convergence.
-The initial coarse grid covers a region $\mc_z\in[1.1962,1.1970] M_\odot$ and $\delta \in [0,0.25]$,  insuring the posterior was smaller than our initial
-coarse grid spacing.     The top  panels of Figure  \ref{fig:BNS:MultiIterate} show posterior distributions derived
-from the first several RIFT iterations, while the bottom panels show convergence diagnostics.
-%% Figure \ref{fig:BNS:Comparison} compares results from the GPU and CPU codes on \editremark{identical test points: the
-%%   initial grid}.\footnote{This comparison of $\ln {\cal L}_{\rm marg}$ is limited by Monte Carlo integration error and
-%%   is provided  as a pedagogically convenient sanity check.  When comparing 
-%%   $\ln {\cal L}_{\rm margT}$ results on identical data and parameters, the GPU and CPU implementations agree to
-%%   significantly higher precision.}  The two results agree.
-Because our analysis uses a post-Newtonian model to interpret a signal generated with SEOBNRv4, the peak posterior
-density is slightly  offset
- from the synthetic signal's parameters
-
-This specific workflow configuration  reduces overall core usage by maximizing GPU use per iteration: after the first
-iteration, only one GPU is active.   
-Specifically,  with the updated RIFT workflow used here, with one instance analyzing each 20 evaluations,  we use five
-core+GPU pairs in the
-first iteration, followed by one core+GPU for remaining iterations.
-By contrast, with the original CPU-only RIFT code analyzing each $\bm{\lambda}$ in parallel, this process requires roughly 20 core-minutes
-per $\bm{\lambda}$, using 100 cores in the first iteration and 20 cores in each
-subsequent evaluation.   
-Note that  because   $N_{\rm eval}=20$ for the GPU is larger than the (hardware-and $f_{\rm sample}$ dependent)  speedup factor between the CPU and
-GPU implementation, the overall wallclock time needed for a end-to-end
-analysis analysis is larger for the GPU workflow.
-We can achieve comparable or smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.
-
-\subsection{Binary neutron star analysis : Assuming  nonprecessing spin}
-\label{sec:sub:BNS_spin}
-
-Binary neutron star models with more parameters like spin and tides require correspondingly larger numbers of points in the
-  initial grid and per iteration.  Fortunately, the number of observationally significant and accessible dimensions is often
-  substantially less than the  prior  dimensionality.  In practice,we use roughly $20\times$ more points per
-  iteration for precessing massive BBH systems ($d=8$) or for spinning binary neutron stars with tides ($d=6$).    As a
-  result, we can still achieve relatively rapid turnaround on a high-dimensional binary neutron star analysis even in
-  resource-constrained environments.
-
-As a concrete demonstration of a realistic modest-latency analysis using the GPU-accelerated code, we reanalyze our fiducial binary
-neutron star signal, accounting for the possibility of nonzero (aligned) neutron star spins.  Our initial grid consists
-of 5000 points, spread \AddedResponse{approximately} uniformly across a 4-dimensional
-hypercube $\mc_z\in[1.1962,1.1970] M_\odot$, $\delta \in[0,0.25]$ and $\chi_{i,z}\in[-0.05,0.05]$.  Subsequent
-iterations use 500 random draws from the estimated posterior samples produced from the previous iteration.  
-Because  our fiducial post-Newtonian model (TaylorT4) as implemented in \texttt{lalsuite}\cite{lalsuite} does not include spin, we
-employ both the \texttt{SEOBNRv4\_ROM} and \texttt{SpinTaylorT4} waveform approximations for nonprecessing binaries,
-omitting higher-order modes.   For SEOBNRv4\_ROM, we use 48 GPU (c) enabled nodes; for SpinTaylorT4, we use 100 GPU (a) nodes.
-%
-Figure \ref{fig:BNS:Spin} shows our results.
-%
-
-Due to our self-imposed resource constraints on $N_{\rm eval}=20$ and the number of GPUs, the first iteration is
-resource-limited and requires
-of order $(5000/20/100)\simeq
-2.5$ to $(5000/20/48)\simeq
- 5$ times as long as the first iteration of the zero-spin BNS analysis described above.  Each marginalized likelihood
- evaluation in the first iteration (as well as subsequent iterations) requires roughly 60 seconds for both SpinTaylorT4 and SEOBNRv4ROM, on average.   Subsequent
-marginalized likelihood iterations are not resource-limited and complete in roughly 35 minutes, depending on hardware.
-As previously, we can achieve comparable or smaller turnaround time for an otherwise identical analysis by adjusting $N_{\rm eval}$ appropriate to the available hardware.
-
-\begin{figure*}
-% COMMAND
-% plot_posterior_corner.py  --posterior-file posterior_samples-5.dat  --posterior-file posterior_samples-4.dat   --posterior-file posterior_samples-3.dat --posterior-file posterior_samples-2.dat --posterior-file posterior_samples-1.dat --parameter mc --parameter delta_mc  --parameter chi_eff --mc-range  '[1.19655,1.197]' --chi-max 0.025 --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc_bns.xml.gz  --composite-file all.net --quantiles None  --posterior-label 5 --posterior-label 4 --posterior-label 3 --posterior-label 2 --posterior-label 1 --use-legend --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-linestyle ':'  --posterior-color 'k' --posterior-color 'r' --posterior-color 'g' --posterior-color 'b' --posterior-color 'k' --lnL-cut 50
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_corner_mc_delta_xi.png}
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_mc_cum.png}
-% python plot_mean_variance.py --convergence-file 20190203-bns-withspin-batch_gpu_lowlatency_noHM_v4ROM_meanVar.dat  --output bns_withspin_lnL_meanVar
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_meanVar.png}
-% python plot_convergence.py --convergence-file 20190203-bns-withspin-batch_gpu_lowlatency_noHM_v4ROM.dat  --output bns_withspin_lnL_converge
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_converge.png}
-\caption{\label{fig:BNS:Spin}\textbf{Demonstration of low-latency spinning analysis}:
-Results for marginal posterior distributions
-  of our fiducial synthetic neutron star, assuming the binary has nonprecessing spins.  Solid contours show credible intervals; solid one-dimensional distributions
-  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
-  value of the underlying marginalized likelihood evaluations.
-Only evaluations with marginalized log-likelihoods within $30$ of the maximum are shown, to increase contrast.
-The solid curves show an analysis with SEOBNRv4\_ROM.
-%  the dotted curves show an analysis with SpinTaylorT4.  \editremark{currenly only v4ROM shown} 
- \emph{Left panel } Posterior distribution
-  over  $\mc$,
-  $\delta=(m_1-m_2)/M$, and $\chi_{\rm eff}$.    \emph{Right panel}: Marginal 1d CDFs of $\delta$, showing convergence.
-\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
-first iteration the
-candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
-\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
-error budget early on, Monte Carlo error is not shown.
-}
-\end{figure*}
-
-\SkipForEarlyCirculation{
-\subsection{Precessing binary black hole analysis}
-\label{sec:sub:BBH_v3}
-
-In this section, we reanalyze our fiducial (zero spin) binary black hole signal without any restrictions on spin
-magnitude or alignment, using a precessing effective--one-body approximation (SEOBNRv3 \editremark{cite}).   
-We start with an initial  \emph{three-dimensional} grid of 5000 points in $\mc \in [20,35] M_\odot$, $\eta
-\in[0.2,0.25]$ and $\chi_{\rm eff} \in [-0.5,0.5]$.  We very conservatively perform 10 iterations, adding 500 points
-candidate evaluation points each iteration, using a volumetric spin prior on both dimensionless spins.
-
-We break our  pipeline into two stages.  In the first five
-iterations, the likelihood is assumed to depend only on  $\mc,\eta,\chi_{\rm eff}$.   By construction, during these
-stages our algorithm proposes random choices for the perpendicular spins $\chi_{i,\perp}$  and the remaining
-antisymmetric degree of freedom for the two candidate spins.    As demonstrated in Figure \ref{fig:BBH:Spin}, well
-before the first five iterations complete, we've thoroughly explored the $\mc,\eta,\chi_{\rm eff}$ distribution and
-converged to a (biased) posterior.
-
-In the second stage, we perform a gaussian process fit which includes  all binary parameters ($\mc,\eta,\chi_{\rm
-  eff},\chi_-,\chi_{i,x},\chi_{i,y}$).  As demonstrated in Figure \ref{fig:BBH:Spin}, we quickly converge to a posterior
-distribution.
-
-Throughout this calculation, each individual marginalized likelihood required roughly 1 minute to evaluate.  With 100
-available GPU+cores, this analysis could require as little as 1.7 hours to evaluate the marginalized likelihood for all
-10 iterations.  In practice, by using $N_{\rm eval}=20$ and therefore restricting to 25 cores for the iterative stages, we insured each iteration after the first required no less than
-20 minutes to complete.  As a result of this self-imposed restriction, the wallclock time necessary to complete all 10 iterations was several hours.
-That said, in this particular case substantially fewer iterations were required to reach convergence.
-
-\begin{figure*}
-% COMMAND
-% ../plot_last_iterations_with.sh --parameter m1 --parameter m2  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':'
-% ../plot_last_iterations_with.sh --parameter mc --parameter delta_mc --parameter chi_eff  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':'
-%  ../plot_last_iterations_with.sh --parameter a1z --parameter a2z --parameter chi1_perp --parameter chi2_perp  --plot-1d-extra --ci-list  '[0.9]' --truth-file /home/oshaughn/unixhome/Projects/LIGO-ILE-Applications/ILE-GPU/ILE-GPU-Paper/demos/mdc.xml.gz  --composite-file all.net --quantiles None  --use-legend --posterior-color 'b' --posterior-color 'k' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle ':' --posterior-linestyle '-' --posterior-linestyle '-' --posterior-color 'g' --posterior-color 'magenta'
-
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_mc_cum.png}
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_meanVar.png}
-\includegraphics[width=0.45\textwidth]{figures/bns_withspin_lnL_converge.png}
-\caption{\label{fig:BBH:Spin}\textbf{Demonstration precessing binary black hole analysis}:
-Results for marginal posterior distributions
-  of our fiducial binary black hole, assuming the binary could have precessing spins.  Solid contours show credible intervals; solid one-dimensional distributions
-  show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
-  value of the underlying marginalized likelihood evaluations.
-Only evaluations with marginalized log-likelihoods within $30$ of the maximum are shown, to increase contrast.
-The solid curves show an analysis with SEOBNRv3.
-%  the dotted curves show an analysis with SpinTaylorT4.  \editremark{currenly only v4ROM shown} 
- \emph{Left panel } Posterior distribution
-  over  $\mc$,
-  $\delta=(m_1-m_2)/M$, and $\chi_{\rm eff}$.    \emph{Right panel}: Marginal 1d CDFs of $\delta$, showing convergence.
-\emph{Bottom left}: Mean and variance of  $\ln{\cal L}_{\rm marg}$ on the evaluation points,  showing that after the
-first iteration the
-candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
-\emph{Bottom right panel}: $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
-error budget early on, Monte Carlo error is not shown.
-}
-\end{figure*}
-}
-
-\SkipForEarlyCirculation{
-\section{ Results (optional?)}
-
-*   Rerun all O2 BBHs with SEOBNRv3?
-
-* PP plot demo for zero-spin binaries, with profiling of full run?
-
-* reminder: HM for BNS break degeneracy, useful for source classification.  Show an example.
-
-\editremark{bonus mentions}: eccentricity
-
-Bonus code: lnLcut tapering, possibly neff tapering
-
-Puffball in DAG every few iterations, to insure stability around edges and not overcover the core
-
-}
+% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_m2.png}
+% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_cum.png}
+% % python plot_mean_variance.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency_meanVar.dat  --output bbh_zerospin_lnL_meanVar
+% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_meanVar.png}
+% % python plot_convergence.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency.dat
+% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_converge.png}
+% \caption{\label{fig:BBH:MultiIterate}\textbf{Convergence of BBH analysis: Zero spin}: Results for marginal posterior distributions
+%   of our fiducial synthetic binary black hole.  Solid contours show credible intervals; solid one-dimensional distributions
+%   show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
+%   value of the underlying marginalized likelihood evaluations.   
+% %The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
+% \emph{Left panel } Posterior distribution
+%   over  $\mc$ and
+%   $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
+% \emph{Bottom left}: Mean and variance of  \AddedResponse{the array $\ln{\cal L}_{\rm marg}(\bm{\lambda}_j)$  for
+% $j=1,2,\ldots N_{\rm eval}$ indexing all candidate sets of intrinsic parameters $\bm{\lambda}_j$ performed in that iteration},  showing that after the
+% first iteration the
+% candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
+% \emph{Bottom right panel}: The estimated evidence $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
+% error budget, Monte Carlo error is not shown.
+% }
+% \end{figure*}
 
@@
 \label{sec:conclude}
-We have demonstrated that the marginalized likelihood $\ln {\cal L}_{\rm marg}(\bm{\lambda})$ appearing in the RIFT/rapidPE parameter
-inference calculation can be evaluated at fixed $\bm{\lambda}$ in tens of seconds on average for
-both binary black holes and binary neutron stars.  This performance improvement could enable very low latency source
-parameter inference for compact binaries, which can be of use for targeting multimessenger followup observations via sky
-localization and  precise source characterization.     This prospect is particularly interesting because RIFT can
-often achieve this performance using computationally costly models for binary merger with rich physics like higher modes
-or eccentricity, as the waveform generation cost does not usually limit code performance.
-
-\AddedResponse{In addition to producing results rapidly, RIFT results can be produced with a noticably smaller overall resource footprint than
-  loosely similar LI analyses, even without tuning to optimize  RIFT pipeline settings.  As a concrete and non-optimized
-  example,  for all five of the iterations of the spinning binary neutron star parameter inference with \texttt{SEOBNRv4\_ROM} from
-  $20\unit{Hz}$ described in this work, our  RIFT analysis expended  roughly 14 core-days.  By contrast, a TaylorF2 analysis of GW170817
-  with LI in MCMC mode starting from $23\unit{Hz}$ required 228 core-days.  % C02-1_TF2LS
-  For precessing binary black holes using SEOBNRv3, the improvement is equally substantial: 10 core-days for a
-  10-iteration investigation of a synthetic GW150914-like source with RIFT, versus 291 core-days for a LI analysis of
-  GW170729.  
-  We defer detailed relative benchmarking using comparably-converged parameter inference to future work.
-}
-
-The overall code performance and thus latency can be further decreased substantially, notably by converting the Monte
-Carlo random number generation and inner products to GPU-based operations.  In such a configuration, the marginalized
-likelihood code would perform almost all calculations (except waveform generation) on a GPU, with minimal communication off
-the board.  This configuration should further reduce the average time needed to compute $\ln {\cal L}_{\rm
-  marg}(\bm{\lambda})$ for both binary black holes (which are Monte Carlo limited) and binary neutron stars (which are
-inner-product limited).   We anticipate a further factor of roughly 10 reduction in overall evaluation time can be
-achieved soon.  At that level of performance, a single 8-GPU machine with contemporary hardware could perform parameter inference less than 10
-minutes.    Since binary compact objects intersect our past light cone only once every roughly 15 minutes, accounting
-for all past history, such a configuration would be able to address low-latency parameter inference for the duration of
-2nd-generation ground-based observing.
-% based on Neval ~ 10^4
-Alternatively, if larger resource pools are available in low latency, both the original and now GPU-accelerated RIFT  can
-perform extremely rapid parameter inference if large iterations are performed completely in parallel (i.e., $N_{\rm
-  eval} \simeq 1$).
-%
-
-By allowing  models with higher-order modes to be used in low latency, our code can exploit the tighter constraints which higher modes can enable on  the properties of low-mass
-binaries with suitable amplitudes and orientations.  These tighter constraints could
-better  inform low-latency source classification and hence multimessenger followup observations of compact binary mergers.
-
-Beyond low-latency multimessenger astronomy, rapid parameter inference enables new applications.  For example, every
-parameter inference provide \emph{evidence} for a signal being present in the data; with rapid parameter inference, this
-evidence could be used as (the last stage in a hierarchical pipeline for) a detection statistic
-%\editremark{citations:old idea,see Thrane/Rory papers recently} 
-\cite{2018PhRvX...8b1019S}.    
-This approach can identify individual events and
-even a population.
-Alternatively and in many ways equivalently, one can identify a population of GW sources without assuming any one is
-real, by  applying parameter inference to more candidate events and self-consistently separating foreground and
-background
-\cite{2015PhRvD..91b3005F,2019MNRAS.tmp..230G}.
-
-When suitable surrogate models are available, the overall code performance and thus latency could be yet again further
-reduced by eliminating the iteration and fitting stages entirely, performing one Monte Carlo at once
-\cite{gwastro-PE-AlternativeArchitecturesROM}.  This approach exploits a linear representation of  $h_{lm}(t)$  via
-basis functions, to enable rapid likelihood evaluation as a function of both extrinsic parameters and  $\bm{\lambda}$.
-Though not necessarily or compactly available for all surrogate models, particularly for the small basis sizes necessary
-to fit onboard GPUs, this approach could enable exceedingly low latency at small computational cost.  This alternative
-architecture would be exceptionally well-suited to the alternative applications of low-latency PE described above.
-
-The use of \texttt{cupy} enables our code to be highly portable across architectures and heterogeneous GPU environments,
-while transitioning smoothly between GPU and CPU mode.     The techniques we used here will be transported to other
-Bayesian inference modeling codes used to interpret GW observations \cite{gwastro-PopulationReconstruct-Parametric-Wysocki2018,gwastro-PopulationReconstruct-Code-PopModels}.
-% Will swap version of code
-
-\AddedResponse{In this work, we have focused exclusively on profiling a simplified pipeline  to produce posterior distributions
-  for detector-frame intrinsic parameters.   The  code  also produces
-reliable extrinsic parameter distributions \cite{gwastro-PE-AlternativeArchitectures}.  With some postprocessing,
-this pipeline  provides joint posterior distributions for all intrinsic and extrinsic parameter distributions together;
-examples of these distributions have been published elsewhere \cite{LIGO-O2-Catalog}.  
-Presently, rather than harvest extrinsic information from every iteration, we harvest joint intrinsic and
-extrinsic information with a single final iteration, which we will implement shortly in our production pipeline.
-}
+In Progress,
 
 \begin{acknowledgements}
-We thank our anonymous referee for helpful feedback.
-DW thanks the RIT COS and CCRG for support, and  NSF-1707965.
-JL is supported by NSF PHY-1707965.  
-RO'S is supported by NSF PHY-1707965 and PHY-1607520.
-Y-LLF was supported in part by BNL LDRD projects No.\,17-029 and No.\,19-002, 
-and New York State Urban Development Corporation, d/b/a Empire State Development, under contract No.\,AA289.
-We thank the Computational Science Initiative at BNL for hosting GPU Hackathon 2018, during which a part of this work was conducted.
-We also thank Jeff Layton at NVIDIA for bringing CuPy to our attention.
+We thank our anonymous referee for the helpful feedback.
+
 \end{acknowledgements}
@@
 \appendix
-\section{Properties of resources used}
-% https://ldas-gridmon.ligo.caltech.edu/condor/
-%
-% > CIT:
-%% > 560 nodes with single GTX 1050 Ti
-%% > 32 nodes with dual K10 PCI cards (shows up as 4 GPU devices) and dual GTX 750 Ti
-%% > 16 nodes with dual GTX 750 Ti (and dual Xeon Phi co-processor)
-%% > 
-%% > LHO and LLO each:
-%% > 40 nodes with GTX 1050 Ti
-%% > single DGX-1 system with 8 tightly coupled V100
-
-\AddedResponse{
-LIGO-CIT worker nodes, denoted by (a)  in the text, are principally S6 nodes with 2-CPU $\times 8$ core Opteron
-2.3 GHz machines with 16 Gb of RAM, with GTX 1050 Ti cards with 4 Gb of RAM.  For LIGO-CIT, profiling reports reflect
-performance averaged over the whole cluster and hence lacks the detailed reporting produced for the other configurations.
-% 
-LIGO-LHO worker nodes with GPUs, denoted by (c) in the text, are a heterogeneous configuration mostly consisting of
-(a).  Unlike profiling at LIGO-CIT, the profiling for LIGO-LHO reflects controlled tests on a single node.
-% nvidia-smi
-The V100 machine (ldas-pcdev13, denoted by (b) in the text) is  a 24-core ES-2650 v4 machine with 4 GPUs, only one of
-which is active in our tests: Tesla V100 with 16 Gb of RAM.
-All non-GPU profiling was  also performed on this machine.  
-}
+In Progress,
 

### 2023-05-09 paper/paper.tex
Added HBM and inference subsection in method section

@@
 
+\subsection{Hierarchical Bayesian Modeling}
+\begin{itemize}
+    \item We have N discrete detection.
+    \item each detection has data $d_1,d_2,d_3,...,d_N$
+    \item each data point $d_N$ has properties denoted by $\lambda_1,\lambda_2,...,\lambda_i$
+    \item  $ \lambda_i$ are parameters such as mass, spin, eccentricity, and location.
+    \item Each parameter has its uncertainty.
+    \item This uncertainty is described by the probability of the data given the parameter value. We call it the likelihood function of one single event. $\mathcal{L}(\lambda) = p(d|\lambda)$
+    \item the waveform gives the parameter value.
+    \item waveform is a mathematical function.
+    \item Once you have a likelihood function, you can use prior( usually it's uniform: which gives equal probability to each event) to find the posterior.
+$p(\lambda|d) \propto p(d|\lambda) p(\lambda)$
+
+\end{itemize}
+
+\subsection{Population Inference}
+
+\begin{itemize}
+    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
+    \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
+    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
+    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
+    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
+    $p(\Lambda|d_1,d_2,d_3,...,d_N)= \frac{p(\Lambda)*p(d_1,d_2,d_3,...,d_N|\Lambda)}{p(d_1,d_2,d_3,...,d_N)}$
+    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
+    \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
+    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
+    $\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)$
+    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$.
+$l_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
+    \item now we can find the posterior by Baye's theorem. \\
+    $p(\mathcal{R,\Lambda})\propto p(\mathcal{R},\Lambda) * \mathcal{L}(\mathcal{R},\Lambda)$
+    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.
+
+\end{itemize}
+
 \subsection{Power law Model}

### 2023-05-09 paper/paper.tex
Adding Power law subsection and VT Subsection

@@
     \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
-    $p(\Lambda|d_1,d_2,d_3,...,d_N)= \frac{p(\Lambda)*p(d_1,d_2,d_3,...,d_N|\Lambda)}{p(d_1,d_2,d_3,...,d_N)}$
+\begin{equation}
+\label{eq:Bayes}    
+p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)}
+\end{equation}
+
     \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
@@
     {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
-    $\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)$
+\begin{equation}
+\label{eq: likelihood}
+\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)    
+\end{equation}
+
     \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$.
@@
     \item now we can find the posterior by Baye's theorem. \\
-    $p(\mathcal{R,\Lambda})\propto p(\mathcal{R},\Lambda) * \mathcal{L}(\mathcal{R},\Lambda)$
+    $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$
     \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.
@@
 \end{itemize}
+\subsection{VT Estimation}
+\begin{itemize}
+    \item Estimation of VT (sensitivity of the LVK) only depend on the mass.
+For the sake of simplicity {\color{red}(physics reason)}, we consider only non-spinning, non-eccentric and nonprecessing BH Binaries.
+    \item Sources are detected if SNR is greater than 8.
+    \item V is characteristic volume in which source can be detected by LVK. It is orientation averaged sensitive 3-volume.
+\begin{equation}
+\label{eq:volume}
+V(\lambda) = \int P((<D(z))/D_n(\lambda))\frac{dV_c}{dz}\frac{dz}{1+z}
+\end{equation}    
+    \item where $D(z)$ is the luminosity distance for redshift z, $V_c$ is the comoving volume.
+
+    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this senstivity for the time $T$, the average number of detections would be as follows: \\
+\begin{equation}
+\label{eq:mu}
+  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda  
+\end{equation}
+    \item where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic paramter $\lambda$.
+Keep in mind that $\lambda$ is equal to all the 15 parameters of a binary.
+\end{itemize}
 
 \subsection{Power law Model}
+The following is the power law model. We added the Gaussian eccentricity in it.
+
+\begin{align}
+\label{eq:plawg}
+p(m_1,m_2,e) = &C(\alpha,k_m,m_{min},m_{max},M_{max},e)  
+\nonumber \\ & \frac{(\frac{m_2}{m_1})^{k_m}m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
+\nonumber \\ &
+\sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
+\end{align}
+We use $k_m=0$, because GW networks are much more senstive to more massive BHs with $M\geq 200 M_\odot$
 
@@
 \end{align}
+
+The probability density would be non zero only for the two following conditions.
+
+\begin{itemize}
+    \item $m_{min}\leq m_2 \leq m_1 \leq m_{max}$
+    \item $m_1+m_2 \leq M_{max}$
+\end{itemize}
+Where $M_{max}$ depends on the detector (LVK)
 

### 2023-05-23 original_pop_code/run_syn_scale_60_8000.sh
adding files

@@ -1,1 +1,1 @@
-time python original_powerlaw_mcmc.py syn_events/event_2.dat syn_events/event_3.dat syn_events/event_6.dat syn_events/event_7.dat syn_events/event_8.dat syn_events/event_10.dat syn_events/event_12.dat syn_events/event_13.dat syn_events/event_14.dat syn_events/event_15.dat syn_events/event_16.dat syn_events/event_17.dat syn_events/event_19.dat syn_events/event_21.dat syn_events/event_22.dat syn_events/event_24.dat syn_events/event_25.dat syn_events/event_26.dat syn_events/event_27.dat syn_events/event_28.dat syn_events/event_29.dat syn_events/event_30.dat syn_events/event_31.dat syn_events/event_32.dat syn_events/event_33.dat syn_events/event_34.dat syn_events/event_35.dat syn_events/event_36.dat syn_events/event_37.dat syn_events/event_38.dat syn_events/event_39.dat syn_events/event_40.dat syn_events/event_41.dat syn_events/event_42.dat syn_events/event_43.dat syn_events/event_44.dat syn_events/event_45.dat syn_events/event_46.dat syn_events/event_48.dat syn_events/event_49.dat syn_events/event_50.dat syn_events/event_51.dat syn_events/event_52.dat syn_events/event_53.dat syn_events/event_54.dat syn_events/event_55.dat syn_events/event_56.dat syn_events/event_57.dat syn_events/event_58.dat syn_events/event_59.dat syn_events/event_60.dat syn_events/event_61.dat syn_events/event_62.dat syn_events/event_63.dat syn_events/event_64.dat syn_events/event_65.dat syn_events/event_66.dat syn_events/event_67.dat syn_events/event_68.dat syn_events/event_69.dat syn_events/event_70.dat syn_events/event_71.dat syn_events/event_72.dat syn_events/event_73.dat syn_events/event_75.dat syn_events/event_76.dat syn_events/event_77.dat syn_events/event_78.dat syn_events/event_79.dat syn_events/event_80.dat syn_events/event_82.dat syn_events/event_83.dat syn_events/event_84.dat syn_events/event_86.dat syn_events/event_87.dat syn_events/event_89.dat syn_events/event_90.dat syn_events/event_91.dat syn_events/event_92.dat syn_events/event_94.dat syn_events/event_95.dat syn_events/event_96.dat syn_events/event_97.dat syn_events/event_98.dat syn_e
[…]

### 2023-05-23 pop_genrator/run_genrator.sh
correcting the script for generalization

@@ -1,2 +1,2 @@
 rm combine_pop.txt ; rm -r txt_events ; rm -r dat_events ; rm -r __pycache__ ;
-python population_gen.py ; mkdir txt_events ; mkdir dat_events; mv eve* ./txt_events ; cd txt_events ; cp eve* /home/muhammad.zeeshan/ms-work/ms-project/pop_genrator/dat_events ; cd .. ; cd dat_events;  rename txt dat *.txt
+python population_gen.py ; mkdir txt_events ; mkdir dat_events; mv eve* ./txt_events ; cd txt_events ; cp eve* ../dat_events ; cd .. ; cd dat_events;  rename txt dat *.txt

### 2023-05-23 original_pop_code/run_syn_scale_60_8000.sh
Rearanging the files

@@ -1,1 +1,0 @@
-time python original_powerlaw_mcmc.py syn_events/event_2.dat syn_events/event_3.dat syn_events/event_6.dat syn_events/event_7.dat syn_events/event_8.dat syn_events/event_10.dat syn_events/event_12.dat syn_events/event_13.dat syn_events/event_14.dat syn_events/event_15.dat syn_events/event_16.dat syn_events/event_17.dat syn_events/event_19.dat syn_events/event_21.dat syn_events/event_22.dat syn_events/event_24.dat syn_events/event_25.dat syn_events/event_26.dat syn_events/event_27.dat syn_events/event_28.dat syn_events/event_29.dat syn_events/event_30.dat syn_events/event_31.dat syn_events/event_32.dat syn_events/event_33.dat syn_events/event_34.dat syn_events/event_35.dat syn_events/event_36.dat syn_events/event_37.dat syn_events/event_38.dat syn_events/event_39.dat syn_events/event_40.dat syn_events/event_41.dat syn_events/event_42.dat syn_events/event_43.dat syn_events/event_44.dat syn_events/event_45.dat syn_events/event_46.dat syn_events/event_48.dat syn_events/event_49.dat syn_events/event_50.dat syn_events/event_51.dat syn_events/event_52.dat syn_events/event_53.dat syn_events/event_54.dat syn_events/event_55.dat syn_events/event_56.dat syn_events/event_57.dat syn_events/event_58.dat syn_events/event_59.dat syn_events/event_60.dat syn_events/event_61.dat syn_events/event_62.dat syn_events/event_63.dat syn_events/event_64.dat syn_events/event_65.dat syn_events/event_66.dat syn_events/event_67.dat syn_events/event_68.dat syn_events/event_69.dat syn_events/event_70.dat syn_events/event_71.dat syn_events/event_72.dat syn_events/event_73.dat syn_events/event_75.dat syn_events/event_76.dat syn_events/event_77.dat syn_events/event_78.dat syn_events/event_79.dat syn_events/event_80.dat syn_events/event_82.dat syn_events/event_83.dat syn_events/event_84.dat syn_events/event_86.dat syn_events/event_87.dat syn_events/event_89.dat syn_events/event_90.dat syn_events/event_91.dat syn_events/event_92.dat syn_events/event_94.dat syn_events/event_95.dat syn_events/event_96.dat syn_events/event_97.dat syn_events/event_98.dat
[…]

### 2023-05-23 run-10-20-1000.sh

@@ -1,1 +1,0 @@
-time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_120_20000.hdf5 --n-walkers 120 --n-samples 20000

### 2023-05-23 original_pop_code/run_syn_org.sh
adding more files in script to run

@@ -1,0 +1,1 @@
+time python original_powerlaw_mcmc.py syn_events/event_1.dat syn_events/event_2.dat syn_events/event_3.dat syn_events/event_4.dat syn_events/event_5.dat syn_events/event_6.dat syn_events/event_7.dat syn_events/event_8.dat syn_events/event_9.dat syn_events/event_10.dat syn_events/event_11.dat syn_events/event_12.dat syn_events/event_13.dat syn_events/event_14.dat syn_events/event_15.dat syn_events/event_16.dat syn_events/event_17.dat syn_events/event_18.dat syn_events/event_19.dat syn_events/event_20.dat syn_events/event_21.dat syn_events/event_22.dat syn_events/event_23.dat syn_events/event_24.dat syn_events/event_25.dat syn_events/event_26.dat syn_events/event_27.dat syn_events/event_28.dat syn_events/event_29.dat syn_events/event_30.dat syn_events/event_31.dat syn_events/event_32.dat syn_events/event_33.dat syn_events/event_34.dat syn_events/event_35.dat syn_events/event_36.dat syn_events/event_37.dat syn_events/event_38.dat syn_events/event_39.dat syn_events/event_40.dat syn_events/event_41.dat syn_events/event_42.dat syn_events/event_43.dat syn_events/event_44.dat syn_events/event_45.dat syn_events/event_46.dat syn_events/event_47.dat syn_events/event_48.dat syn_events/event_49.dat syn_events/event_50.dat syn_events/event_51.dat syn_events/event_52.dat syn_events/event_53.dat syn_events/event_54.dat syn_events/event_55.dat syn_events/event_56.dat syn_events/event_57.dat syn_events/event_58.dat syn_events/event_59.dat syn_events/event_60.dat syn_events/event_61.dat syn_events/event_62.dat syn_events/event_63.dat syn_events/event_64.dat syn_events/event_65.dat syn_events/event_66.dat syn_events/event_67.dat syn_events/event_68.dat syn_events/event_69.dat syn_events/event_70.dat syn_events/event_71.dat syn_events/event_72.dat syn_events/event_73.dat syn_events/event_74.dat syn_events/event_75.dat syn_events/event_76.dat syn_events/event_77.dat syn_events/event_78.dat syn_events/event_79.dat syn_events/event_80.dat syn_events/event_81.dat syn_events/event_82.dat syn_events/event_83.dat syn_events/event_84.dat syn_e
[…]

### 2023-05-23 run.sh
adding script file after correction

@@ -1,0 +1,1 @@
+time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_120_20000.hdf5 --n-walkers 120 --n-samples 20000

### 2023-05-23 run_syn_new.sh

@@ -1,0 +1,1 @@
+time python new_powerlaw_mcmc.py syn_events/event_1.dat syn_events/event_2.dat syn_events/event_3.dat syn_events/event_4.dat syn_events/event_5.dat syn_events/event_6.dat syn_events/event_7.dat syn_events/event_8.dat syn_events/event_9.dat syn_events/event_10.dat syn_events/event_11.dat syn_events/event_12.dat syn_events/event_13.dat syn_events/event_14.dat syn_events/event_15.dat syn_events/event_16.dat syn_events/event_17.dat syn_events/event_18.dat syn_events/event_19.dat syn_events/event_20.dat syn_events/event_21.dat syn_events/event_22.dat syn_events/event_23.dat syn_events/event_24.dat syn_events/event_25.dat syn_events/event_26.dat syn_events/event_27.dat syn_events/event_28.dat syn_events/event_29.dat syn_events/event_30.dat syn_events/event_31.dat syn_events/event_32.dat syn_events/event_33.dat syn_events/event_34.dat syn_events/event_35.dat syn_events/event_36.dat syn_events/event_37.dat syn_events/event_38.dat syn_events/event_39.dat syn_events/event_40.dat syn_events/event_41.dat syn_events/event_42.dat syn_events/event_43.dat syn_events/event_44.dat syn_events/event_45.dat syn_events/event_46.dat syn_events/event_47.dat syn_events/event_48.dat syn_events/event_49.dat syn_events/event_50.dat syn_events/event_51.dat syn_events/event_52.dat syn_events/event_53.dat syn_events/event_54.dat syn_events/event_55.dat syn_events/event_56.dat syn_events/event_57.dat syn_events/event_58.dat syn_events/event_59.dat syn_events/event_60.dat syn_events/event_61.dat syn_events/event_62.dat syn_events/event_63.dat syn_events/event_64.dat syn_events/event_65.dat syn_events/event_66.dat syn_events/event_67.dat syn_events/event_68.dat syn_events/event_69.dat syn_events/event_70.dat syn_events/event_71.dat syn_events/event_72.dat syn_events/event_73.dat syn_events/event_74.dat syn_events/event_75.dat syn_events/event_76.dat syn_events/event_77.dat syn_events/event_78.dat syn_events/event_79.dat syn_events/event_80.dat syn_events/event_81.dat syn_events/event_82.dat syn_events/event_83.dat syn_events/event_84.dat syn_eve
[…]

### 2023-05-23 paper/paper.tex
Updates from Overleaf

@@
 p(m_1,m_2,e) = &C(\alpha,k_m,m_{min},m_{max},M_{max},e)  
-\nonumber \\ & \frac{(\frac{m_2}{m_1})^{k_m}m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
+\nonumber \\ & \frac{(m_2/m_1)^{k_m}m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
 \nonumber \\ &
@@
 Where $M_{max}$ depends on the detector (LVK)
+
+\section{Inference of Synthetic BBH Population}
+We have generated two distinct populations consistent of $100$ BBH by fixing the power law model parameter $\Lambda$ given in Eq. \ref{eq:plaw}. 
+The parameters $\alpha=-2$, $m_{min}=10$, $m_{max}=50$ are same for both populations.
+The only difference between two generated populations is the dispersion of $\sigma_e$.
+The range for the first populations is $0<\sigma_e<0.5$ and for the second one is $0<\sigma_e<0.38$.
+Furthermore, to make the comparison of the eccentric population with non-eccentric, we have created two more different populations by scaling already generated two populations by using the Eq.
+\ref{eq:scaling}.
+The Eq.
+\ref{eq:scaling} removes the eccentric component from the population, scales the masses and eventually we get non-eccentric population.
+\begin{align}
+\label{eq:scaling}
+M^{ecc} = \frac{M}{(1-\frac{157}{24}e^2)^{3/5}}
+\end{align}
+
+\subsection{First Population: $0<\sigma_e<0.5$}
+The following is the figure which shows the
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d05.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05scl.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05diff.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss05.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\subsection{Second Population: $0<\sigma_e<0.38$}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\end{figure}
 

### 2023-05-23 paper/paper.tex
Updates from Overleaf

@@
 
-    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this senstivity for the time $T$, the average number of detections would be as follows: \\
+    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this sensitivity for the time $T$, the average number of detection would be as follows: \\
 \begin{equation}
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d038.png}
 \caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038.png}
 \caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
@@
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038scl.png}
 \caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
 \end{figure}
-
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038diff.png}
 \caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
 \end{figure}
-
 \begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss038.png}
 \caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}

### 2023-05-25 paper/paper.tex
Converted Bullets into paragraph with few citations for the Method section A, B , C

@@
 
-\title{Population Inference of Eccentric Binary Black Holes}
+\title{Population Inference of Non-Spinning Eccentric Binary Black Holes}
 \author{M. Zeeshan}
@@
 \section{Introduction}
-The Advanced LIGO  \cite{2015CQGra..32g4001T}  and Virgo \cite{gw-detectors-Virgo-original-preferred}  ground-based gravitational wave (GW) detectors have
-identified several coalescing compact binaries
-\cite{DiscoveryPaper,LIGO-O1-BBH,2017PhRvL.118v1101A,LIGO-GW170814,LIGO-GW170608,LIGO-GW170817-bns}.  
-%
+Later this section
 
@@
 
-The Binary Black Holes (BBHs) can be described by three intrinsic and seven extrinsic parameters.
-The intrinsic parameters: such as mass ($m_i$), spin $( \chi_i)$, and eccentricity $(e)$, are subject to the orbital evolution of the binary.
-The extrinsic parameters: orientation (orbital phase, polarization, and inclination), sky location (right ascension, declination), luminosity distance, and coalescence time depend on the observer.
+The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters.
+The intrinsic parameters mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $(e)$  are subject to the orbital evolution of the binary.
+The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.
 
-{\color{red}Most of the eccentric binaries are driven by low mass events \cite{?}}.
-Therefore,  in this study, we  will consider modest masses $(1M_\odot-40M_\odot)$, non-spinning $(\chi_i = 0)$, and eccentric $(0-1)$ binaries.
-In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2$.
+Most of the eccentric binary black hole (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary have mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
+\cite{mass_ecc_limit_2018}.
+Therefore,  in this study, we  will consider binaries with non-spinning $(\chi_i = 0)$, eccentric $(0-0.5)$ and moderate mass range $(10M_\odot-50M_\odot)$.
+In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2=100M_\odot$.
 
-\subsection{Hierarchical Bayesian Modeling}
-\begin{itemize}
-    \item We have N discrete detection.
-    \item each detection has data $d_1,d_2,d_3,...,d_N$
-    \item each data point $d_N$ has properties denoted by $\lambda_1,\lambda_2,...,\lambda_i$
-    \item  $ \lambda_i$ are parameters such as mass, spin, eccentricity, and location.
-    \item Each parameter has its uncertainty.
-    \item This uncertainty is described by the probability of the data given the parameter value. We call it the likelihood function of one single event. $\mathcal{L}(\lambda) = p(d|\lambda)$
-    \item the waveform gives the parameter value.
-    \item waveform is a mathematical function.
-    \item Once you have a likelihood function, you can use prior( usually it's uniform: which gives equal probability to each event) to find the posterior.
+\subsection{Hierarchical Bayesian Modeling (HBM)}
+
+We discussed the various formation scenario of EBBHs in the previous section; each one has a distinct signature in the GW signals.
+Hence, it's trivial to observe the effect of each scenario or the mixed effect on the merging population.
+It leads us to make the population model, which considers each scenario and allows distinct merger rates for them.
+
+Once we have the population models, our next obstacle is constraining them with observational data.
+While constraining the model, we must consider the sensitivity of LIGO-VIRGO-KAGRA (LVK) and uncertainty in each detection.
+Therefore, the HBM is a good choice to constrain the population model.
+
+In the HBM, we have the N number of discrete detection, in our case, from LVK.
+Those detections provide merger data denoted as $d_1,d_2,d_3,...,d_N$ where each $d_i$ shows the BBH.
+Each $d_i$ or merger has mass, spin, and eccentricity properties.
+These properties, often called parameters, are denoted by $\lambda_1,\lambda_2,\lambda_3,...,\lambda_i$.
+Each parameter has its uncertainty, and we express it by the probability of the data given the parameter value.
+We also refer to it as the likelihood function $\mathcal{L}(\lambda)=p(d|\lambda)$ of one event/detection.
+The given value of $\lambda$ to calculate the likelihood is obtained from the mathematical model of the waveform.
+Once you have a likelihood function, you can use a prior (usually, a uniform) to find the posterior probability using the Bayes theorem.
+
+\begin{equation}
+\label{eq:Bayes_ind}    
+p(\lambda|d) \propto p(d|\lambda) p(\lambda)
+\end{equation}
+
+This posterior probability will constrain the properties of each binary, such as mass, spin, and eccentricity.
+We may get the inference of those parameters using Rapid parameter Inference on gravitational wave sources via Iterative Fitting (RIFT): an open source code for parameter estimation of the binary sources \cite{rift_2018}.
+
+%\begin{itemize}
+%    \item We have N discrete detection.
+%    \item each detection has data $d_1,d_2,d_3,...,d_N$
+%    \item each data point $d_N$ has properties denoted by $\lambda_1,\lambda_2,...,\lambda_i$
+%    \item  $ \lambda_i$ are parameters such as mass, spin, eccentricity, and location.
+%    \item Each parameter has its uncertainty.
+%    \item This uncertainty is described by the probability of the data given the parameter value. We call it the likelihood function of one single event. $\mathcal{L}(\lambda) = p(d|\lambda)$
+%    \item the waveform gives the parameter value.
+%    \item waveform is a mathematical function.
+%    \item Once you have a likelihood function, you can use prior( usually it's uniform: which gives equal probability to each event) to find the posterior.
 $p(\lambda|d) \propto p(d|\lambda) p(\lambda)$
 
-\end{itemize}
+%\end{itemize}
 
@@
 
-\begin{itemize}
-    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
-    \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
-    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
-    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
-    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
+Now having the parameters of individual sources, we are ready to make a population inference.
+Following the Bayesian framework for inference, our first task is to find the likelihood of the population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ equivalent to the probability of the individual sources given the population parameter $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, \sigma_e)$ written as follows.
+
+\begin{equation}
+\label{eq:likelihood_pop}    
+\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)
+\end{equation}
+
+where $\alpha$ is the power law index given in Eq.
+\ref{eq:plaw}, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum, and maximum masses of the binary components in the population respectively, and $\sigma_e$ is the orbital eccentricity distribution.
+
+The second step is to find the posterior probability, and we can do this by using the above likelihood in the Bayes theorem as given below.
+
 \begin{equation}
@@
 
-    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
-    \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
-    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
+where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or sometime known as evidence.
+
+To conduct our specific analysis for mass and eccentricity distribution, we will use the inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
+
 \begin{equation}
 \label{eq: likelihood}
-\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda l_n(\lambda) \mathcal{R} p(\lambda|\Lambda)    
+\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda \ell_n(\lambda) \mathcal{R} p(\lambda|\Lambda),
 \end{equation}
 
-    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$.
+Where $\mu(\mathcal{R},\Lambda)$ is the expected number of detection under the given population parametrization $\Lambda$ with the overall rate $\mathcal{R}$.
+$\ell_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
+Finally, we will get our posterior as $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$.
+
+The soundest approach is to find the analytical posterior.
+However, most of the time, it's difficult to find the analytical solution to such a complicated equation involving multidimensional integrals.
+Therefore, a practical approach is to use the computational methods.
+Our study will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} method to find the posterior distribution of population parameters.
+This method draws samples (set of parameters $\alpha,\mathcal{R},m_{min},m_{max},\sigma_e$ that make one model) from the targeted distribution; in our case, it's the power law model given in Eq.
+\ref{eq:plaw}, then compares it with the given data (collection of individual events) and store the best-fit sample.
+We may iterate this as we need and store multiple sample values untill they converge.
+This method is in a Python package called EMCEE \cite{emcee_paper}.
+
+%\begin{itemize}
+%    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
+ %   \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
+%    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
+%    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
+%    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
+%\begin{equation}
+%\label{eq:Bayes}    
+%p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)}
+%\end{equation}
+
+%    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
+
+%   \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
+%    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
+%\begin{equation}
+%\label{eq: likelihood}
+%\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-%\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda %\ell_n(\lambda) \mathcal{R} p(\lambda|\Lambda)    
+%\end{equation}
+
+%    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$.
 $l_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
-    \item now we can find the posterior by Baye's theorem. \\
-    $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$
-    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.
+ %   \item now we can find the posterior by Baye's theorem. \\
+ %   $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$
+%    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.
 
-\end{itemize}
-\subsection{VT Estimation}
-\begin{itemize}
-    \item Estimation of VT (sensitivity of the LVK) only depend on the mass.
-For the sake of simplicity {\color{red}(physics reason)}, we consider only non-spinning, non-eccentric and nonprecessing BH Binaries.
-    \item Sources are detected if SNR is greater than 8.
-    \item V is characteristic volume in which source can be detected by LVK. It is orientation averaged sensitive 3-volume.
+%\end{itemize}
+
+\subsection{Volume Time (VT) Estimation}
+
+To make our study realistic, including the sensitivity of the LVK instruments is significant.
+This sensitivity is defined by VT estimation.
+$V$ is the characteristic volume with units $Gpc^{-3}yr^{-1}$, which refers to the possible detection region in the sky for the LVK \cite{Volume_1993}, and $T$ is the time duration of making observations at this sensitivity.
+Existing LVK instruments' sensitivity depends only on the mass; therefore, we will get the same VT calculated previously \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries.
+Hence, we briefly explain the calculations, but one may read Daniel Wysocki's article \cite{Dan_2019} for a deeper study.
+
+The Eq. \ref{eq:volume} calculates the orientation averaged sensitive volume \cite{Abbott_2016,richard2010volume}
+
 \begin{equation}
@@
 \end{equation}    
-    \item where $D(z)$ is the luminosity distance for redshift z, $V_c$ is the comoving volume.
+where $D(z)$ is the luminosity distance for redshift $z$, $V_c$ is the comoving volume.
+Finally, to compute the average number of detection, we will use Eq.
+\ref{eq:mu} and keep in mind that the merger will be accepted only if the signal-to-noise ratio exceeds 8 \cite{SNR_2010}.
 
-    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this sensitivity for the time $T$, the average number of detection would be as follows: \\
 \begin{equation}
@@
 \end{equation}
-    \item where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic paramter $\lambda$.
+where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic parameter $\lambda$.
+Keep in mind that $\lambda$ is equal to all intrinsic and extrinsic parameters.
+
+%\begin{itemize}
+%    \item Estimation of VT (sensitivity of the LVK) only depend on the mass.
+For the sake of simplicity {\color{red}(physics reason)}, we consider only non-spinning, non-eccentric and nonprecessing BH Binaries.
+%    \item Sources are detected if SNR is greater than 8.
+%    \item V is characteristic volume in which source can be detected by LVK. It is orientation averaged sensitive 3-volume.
+%\begin{equation}
+%\label{eq:volume}
+%V(\lambda) = \int P((<D(z))/D_n(\lambda))\frac{dV_c}{dz}\frac{dz}{1+z}
+%\end{equation}    
+%    \item where $D(z)$ is the luminosity distance for redshift z, $V_c$ is the comoving volume.
+
+%    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this sensitivity for the time $T$, the average number of detection would be as follows: \\
+%\begin{equation}
+%\label{eq:mu}
+%  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda  
+%\end{equation}
+%    \item where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic paramter $\lambda$.
 Keep in mind that $\lambda$ is equal to all the 15 parameters of a binary.
-\end{itemize}
+%\end{itemize}
 

### 2023-05-25 paper/paper.tex
Adding power law section with citation and commenting the bullets

@@
 \subsection{Power law Model}
-The following is the power law model. We added the Gaussian eccentricity in it.
+There are various weak and pure phenomenological population models proposed in previous studies \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model,2018talbot_bbh_model}.
+However, our analysis used the pure truncated power law defined in \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model} and added the Gaussian eccentricity.
+This model computes the intrinsic probability of $m_1$, $m_2$, and $e$.
+We also assume that non-zero probability density only exists for $m_{min}<\leq m_2 \leq m_1 \leq m_{max}$ and for total mass $M_{max}=m_1+m_2 = 200 M_\odot$.
+The condition of the total mass is only because of the limitations of the detectors towards the higher masses \cite{2016PRXAbbot_BBH_model}.
+The generalized form of the truncated power-law model with parameters $m_{min}$, $m_{max}$, $k_m$, $\alpha$, $\sigma_e$ and random variable $m_1$, $m_2$, and $e$ has the functional form in Eq.
+\ref{eq:plawg} within provided mass limit.
 
@@
 \end{align}
-We use $k_m=0$, because GW networks are much more senstive to more massive BHs with $M\geq 200 M_\odot$
+
+The Eq.
+\ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha_m$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
+For our analysis, we defined the constant of integration as equal to $\int_A dm_1 dm_2 p(m_1,m_2) dm_1 dm_2 = 1$.
+Our detectors are very sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$; therefore, we will use $k_m=0$ throughout the studies.
+Therefore, we have our reduced form of the truncated power law in Eq.
+\ref{eq:plaw}
 
@@
 \label{eq:plaw}
-p(m_1,m_2,e) = &C(\alpha,m_{min},m_{max},M_{max},e)  
-\nonumber \\ & \frac{m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
-\nonumber \\ &
+p(m_1,m_2,e) =   \frac{m_1^{-\alpha}}{(m_1-m_{min})\sigma_e}  
 \sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
 \end{align}
-
-The probability density would be non zero only for the two following conditions.
-
-\begin{itemize}
-    \item $m_{min}\leq m_2 \leq m_1 \leq m_{max}$
-    \item $m_1+m_2 \leq M_{max}$
-\end{itemize}
-Where $M_{max}$ depends on the detector (LVK)
 

### 2023-05-26 paper/paper.tex
Adding Fiducial model with figures and few citation

@@
 \usepackage{multirow}
+\usepackage{amssymb}
 %
@@
 The Eq.
-\ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha_m$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
+\ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
 For our analysis, we defined the constant of integration as equal to $\int_A dm_1 dm_2 p(m_1,m_2) dm_1 dm_2 = 1$.
@@
 
-\section{Inference of Synthetic BBH Population}
-We have generated two distinct populations consistent of $100$ BBH by fixing the power law model parameter $\Lambda$ given in Eq. \ref{eq:plaw}. 
-The parameters $\alpha=-2$, $m_{min}=10$, $m_{max}=50$ are same for both populations.
-The only difference between two generated populations is the dispersion of $\sigma_e$.
-The range for the first populations is $0<\sigma_e<0.5$ and for the second one is $0<\sigma_e<0.38$.
-Furthermore, to make the comparison of the eccentric population with non-eccentric, we have created two more different populations by scaling already generated two populations by using the Eq.
-\ref{eq:scaling}.
+\subsection{Fiducial Model for Synthetic Population}
+
+We choose the fixed parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$ to generate the population of $N=100$ sources from the power model given in Eq.
+\ref{eq:plawg}.
+We have generated two different populations from the fiducial model to show the significance of the eccentricity on the population.
+The selection of all the parameters is the same for both populations instead of $\sigma_e$.
+This different eccentricity distribution in both models will show how the events can be missed by not including the eccentricity in our analysis.
+
+Furthermore, to compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we have created two more different populations by scaling the existing populations using the Eq.
+\ref{eq:scaling} defined in \cite{2021_scaling_paper}.
 The Eq.
-\ref{eq:scaling} removes the eccentric component from the population, scales the masses and eventually we get non-eccentric population.
+\ref{eq:scaling} removes the eccentric component from the population, scales the masses, and eventually, we get a non-eccentric population to compare with the existing eccentric population.
 \begin{align}
@@
 
-\subsection{First Population: $0<\sigma_e<0.5$}
-The following is the figure which shows the
+\subsubsection{EBBH and NEBBH Population with $0\leq \sigma_e \leq 0.5$}
+ The first population has the Gaussian distribution bounded between $0\leq \sigma_e \leq 0.5$ given in Fig. \ref{fig:pop3d05}.
 
-\begin{figure}
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/pop3d05.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs for $0\leq \sigma_e \leq 0.5$}}
 \end{figure}
 
-\begin{figure}
+We have the total $100$ events with a minimum and maximum primary mass of $17.73 M_\odot$ and $49.14 M_\odot$, respectively.
+Similarly, we have the secondary minimum and maximum masses $10.11 M_\odot$ and $46.68 M_\odot$ given in Fig \ref{fig:pop2d05}.
+
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/pop2d05.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:pop2d05}\textbf{Mass distribution of EBBHs for $0\leq \sigma_e \leq 0.5$}}
 \end{figure}
 
-\begin{figure}
+To get the non-eccentric population, we apply the Eq.
+\ref{eq:scaling} on the population given in Fig.
+\ref{fig:pop2d05} and get the new population with scaled masses.
+We get the new minimum and maximum values for the primary and secondary masses in Fig.
+\ref{fig:pop2dscal05}.
+
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/pop2d05scl.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05diff.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
-\end{figure}
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss05.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:pop2dscal05}\textbf{Scaled Mass distribution of NEBBHs for  $0\leq \sigma_e \leq 0.5$}}
 \end{figure}
 
-\subsection{Second Population: $0<\sigma_e<0.38$}
+The significant effect after the scaling is not only the mass shift, but one may notice in Fig.
+\ref{fig:pop2dscal05} that $14$ sources were missed out of $100$.
+This reflects that we may miss the various sources by ignoring the effect of eccentricity.
+The mass scaling effect is visible in Fig.
+\ref{fig:pop2d05diff}, and we also notice that sources start missing after $e \gtrsim 0.38$.
 
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d038.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\begin{figure}[H]
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05diff.png}
+\caption{\label{fig:pop2d05diff}\textbf{Total Mass of Binaries for EBBHs and NEBBHs for $0\leq \sigma_e \leq 0.5$:} The blue dots show the EBBHs, the orange shows scaled NEBBHs, and green color shows the increasing difference between those masses with eccentricity}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\begin{figure}[H]
+\includegraphics[width=0.45\textwidth]{paper/figures/massloss05.png}
+\caption{\label{fig:population05}\textbf{The relation between eccentricity and mass loss}}
 \end{figure}
 
-\begin{figure}
+\subsubsection{EBBH and NEBBH Population with $0\leq \sigma_e \leq 0.38$}
+ The second population has the Gaussian distribution bounded between $0\leq \sigma_e \leq 0.38$ given in Fig. \ref{fig:pop3d038}.
+
+\begin{figure}[H]
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d038.png}
+\caption{\label{fig:pop3d038}\textbf{3D-Synthetic Population of EBBHs for $0\leq \sigma_e \leq 0.38$}}
+\end{figure}
+
+We have the total $100$ events with a minimum and maximum primary mass of $15.12 M_\odot$ and $49.98 M_\odot$, respectively.
+Similarly, we have the secondary minimum and maximum masses $10.12 M_\odot$ and $46.01 M_\odot$ given in Fig \ref{fig:pop2d038}.
+
+\begin{figure}[H]
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038.png}
+\caption{\label{fig:pop2d038}\textbf{Mass distribution of EBBHs with $0\leq \sigma_e \leq 0.38$}}
+\end{figure}
+
+Similarly, to get the non-eccentric population, we apply the Eq.
+\ref{eq:scaling} on the population given in Fig.
+\ref{fig:pop2d038} and get the new population with scaled masses.
+We get the new minimum and maximum values for the primary and secondary masses in Fig.
+\ref{fig:pop2d038scl}.
+
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/pop2d038scl.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:pop2d038scl}\textbf{Scaled mass distribution of NEBBHs for $0\leq \sigma_e \leq 0.38$}}
 \end{figure}
-\begin{figure}
+
+Selecting a reduced range of $\sigma_e$ leads to avoiding the missing events.
+Although mass gets scaled, we did not see any missing events in Fig.
+\ref{fig:pop2d038scl}.
+The mass scaling effect is visible in Fig.
+\ref{fig:pop2d038diff} and reflects the importance of eccentricity.
+
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/pop2d038diff.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:pop2d038diff}\textbf{Total Mass of Binaries for EBBHs and NEBBHs for $0\leq \sigma_e \leq 0.38$:} The blue dots show the EBBHs, the orange shows scaled NEBBHs, and green color shows the increasing difference between those masses with eccentricity}
 \end{figure}
-\begin{figure}
+
+\begin{figure}[H]
 \includegraphics[width=0.45\textwidth]{paper/figures/massloss038.png}
-\caption{\label{fig:population05}\textbf{Synthetic Population of Eccentric BBHs}}
+\caption{\label{fig:population038}\textbf{The relation between eccentricity and mass loss}}
 \end{figure}
+
+\section{Inference of Synthetic BBH Population}
+
+Need to write it later.
 

### 2023-06-02 population-genrator/genrating.py
Weighting the Population with VTs

@@ -1,0 +1,36 @@
+'''
+We are creating this file to generate the synthetic population given the fixed paramters of the powerlaw model defined in the paper.
+The power law model which is defined as 'prob.joint' function in the syn_pop_prob.py file, we are importing this fucntion  here and generating the random events by putting the fix parameters of the population model. suh as N=number of events, alpha = power law factor, m_min, m_max, and M_max = m1+m2 : we are impsing uper bound 200 for the whole mass of the binary due to astrophysical reasons. 
+
+To get any desired non spinining population, you can just change the paramter values in the follwing fucntion. 
+'''
+import numpy as np
+import syn_pop_prob
+
+# syn_pop_prob.joint_rvs(N,alpha,m_min,m_max,M_max=m1+m2,sigma_e)
+
+array = syn_pop_prob.joint_rvs(10000,-2,10,50,100,0.2)
+
+
+col_names = ["m1_source", "m2_source", "ecc"]
+
+#saving array in a single file
+
+np.savetxt("population.dat", array, header="\t".join(col_names))
+
+# saving each row into a seprate file
+
+#for i in range(array.shape[0]):
+#    filename = f"row_{i+1}.txt"
+#    row = array[i,:]
+#    np.savetxt(filename, [row], delimiter="\t", fmt="%.3f",header="\t".join(col_names))
+
+# Saving each row of the array to a separate file with column names and copying the row 4000 times
+
+#for i in range(array.shape[0]):
+#    filename = f"event_{i+1}.txt"
+#    row = array[i,:]
+#    data = np.repeat([row], 4000, axis=0)
+#    np.savetxt(filename, data, delimiter="\t", header="\t".join(col_names))
+
+#print(array)

### 2023-06-02 population-genrator/population_gen.py

@@ -1,36 +1,0 @@
-'''
-We are creating this file to generate the synthetic population given the fixed paramters of the powerlaw model defined in the paper.
-The power law model which is defined as 'prob.joint' function in the syn_pop_prob.py file, we are importing this fucntion  here and generating the random events by putting the fix parameters of the population model. suh as N=number of events, alpha = power law factor, m_min, m_max, and M_max = m1+m2 : we are impsing uper bound 200 for the whole mass of the binary due to astrophysical reasons. 
-
-To get any desired non spinining population, you can just change the paramter values in the follwing fucntion. 
-'''
-import numpy as np
-import syn_pop_prob
-
-# syn_pop_prob.joint_rvs(N,alpha,m_min,m_max,M_max=m1+m2,sigma_e)
-
-array = syn_pop_prob.joint_rvs(100,-2,10,50,100,0.2)
-
-
-col_names = ["m1_source", "m2_source", "ecc"]
-
-#saving array in a single file
-
-np.savetxt("combine_pop.txt", array, header="\t".join(col_names))
-
-# saving each row into a seprate file
-
-#for i in range(array.shape[0]):
-#    filename = f"row_{i+1}.txt"
-#    row = array[i,:]
-#    np.savetxt(filename, [row], delimiter="\t", fmt="%.3f",header="\t".join(col_names))
-
-# Saving each row of the array to a separate file with column names and copying the row 4000 times
-
-for i in range(array.shape[0]):
-    filename = f"event_{i+1}.txt"
-    row = array[i,:]
-    data = np.repeat([row], 4000, axis=0)
-    np.savetxt(filename, data, delimiter="\t", fmt="%.3f", header="\t".join(col_names))
-
-#print(array)

### 2023-06-02 population-genrator/run_complete.sh

@@ -1,0 +1,13 @@
+# Script for population genration
+rm population.dat ;
+rm -r __pycache__ ;
+python genrating.py
+
+#script for weighting 
+rm -r __pycache__ ;
+rm -r weighted_events ;
+mkdir weighted_events;
+python weighting.py --dat population.dat --vt vt_m1_m2.hdf5 --n-out 100
+
+#script for scaling
+rm -r scaled_events; mkdir scaled_events ; python scaling.py

### 2023-06-02 population-genrator/run_scaling.sh

@@ -1,1 +1,0 @@
-rm -r scaled_events; mkdir scaled_events ; python scaling.py ; cd scaled_events ; rename txt dat *.txt

### 2023-06-02 population-genrator/scaling.py

@@ -2,7 +2,7 @@ import numpy as np
 import os
 
 # Directory containing the input files
-input_dir = './txt_events/'
+input_dir = './weighted_events/'
 
 # Directory to save the output files
 output_dir = './scaled_events/'
@@ -10,9 +10,9 @@ col_names = ["m1_source", "m2_source"]
 # Iterate over the text files
 for i in range(1, 101):
     # Generate the file path for the current iteration
-    file_path = os.path.join(input_dir, f'event_{i}.txt')
+    file_path = os.path.join(input_dir, f'event_{i}.dat')
 
-    # Load the data from the text file
+    # Load the data from the .dat file
     data = np.loadtxt(file_path)
 
     # Extract the columns
@@ -28,11 +28,11 @@ for i in range(1, 101):
     scaled_data = np.column_stack((scaled_col1, scaled_col2))
 
     # Generate the output file path for the current iteration
-    output_file_path = os.path.join(output_dir, f'event_{i}.txt')
+    output_file_path = os.path.join(output_dir, f'event_{i}.dat')
 
     # Save the scaled data to a new text file
     #np.savetxt(output_file_path, scaled_data, header='m1_source m2_source', comments='')
-    np.savetxt(output_file_path, scaled_data, delimiter="\t", fmt="%.3f", header="\t".join(col_names))
+    np.savetxt(output_file_path, scaled_data, delimiter="\t", header="\t".join(col_names))
 
 
 
@@ -50,7 +50,7 @@ count = 0
 
 # Iterate over the files in the folder
 for file_name in os.listdir(folder_path):
-    if file_name.endswith('.txt'):
+    if file_name.endswith('.dat'):
         file_path = os.path.join(folder_path, file_name)
 
         # Load the data from the text file
@@ -70,7 +70,7 @@ print(f"Total files with NaN or zero values: {count}")
 
 
 # Load the data from the text file and scaling the single txt file which have the complete population
-data01 = np.loadtxt('combine_pop.txt')
+data01 = np.loadtxt('weighted_population.dat')
 
 # Extract the columns
 col01 = data01[:, 0]
@@ -87,5 +87,5 @@ scaled_col02 = col02 * scale_factor01
 scaled_data01 = np.column_stack((scaled_col01, scaled_col02))
 
 # Save the scaled data to a new text file
[…]

### 2023-06-02 population-genrator/vt.py

@@ -1,0 +1,455 @@
+"""
+This module is a slightly modified version of a module written by Will Farr,
+available at
+<https://git.ligo.org/RatesAndPopulations/O2Populations/blob/fe81c5d064283c94e12a19567e020b9e9930efef/code/vt.py>
+
+The main changes made here are for data I/O, and the majority of the
+implementation is the original code written by Will Farr.
+"""
+
+
+from __future__ import print_function
+
+
+
+import astropy.cosmology as cosmo
+import astropy.units as u
+import multiprocessing as multi
+import numpy as np
+from numpy import (
+    array, linspace,
+    sin, cos,
+    square,
+    trapz,
+)
+
+#from pylab import *
+
+def draw_thetas(N):
+    """Draw `N` random angular factors for the SNR.
+
+    Theta is as defined in [Finn & Chernoff
+    (1993)](https://ui.adsabs.harvard.edu/#abs/1993PhRvD..47.2198F/abstract).
+    """
+
+    cos_thetas = np.random.uniform(low=-1, high=1, size=N)
+    cos_incs = np.random.uniform(low=-1, high=1, size=N)
+    phis = np.random.uniform(low=0, high=2*np.pi, size=N)
+    zetas = np.random.uniform(low=0, high=2*np.pi, size=N)
+
+    Fps = 0.5*cos(2*zetas)*(1 + square(cos_thetas))*cos(2*phis) - sin(2*zetas)*cos_thetas*sin(2*phis)
+    Fxs = 0.5*sin(2*zetas)*(1 + square(cos_thetas))*cos(2*phis) + cos(2*zetas)*cos_thetas*sin(2*phis)
+
+    return np.sqrt(0.25*square(Fps)*square(1 + square(cos_incs)) + square(Fxs)*square(cos_incs))
+
+
+
+def next_pow_two(x):
+    """Return the next (integer) power of two above `x`.
+
+    """
+
+    x2 = 1
+    while x2 < x:
+        x2 = x2 << 1
+    return x2
+
+def optimal_snr(m1_intrinsic, m2_intrinsic, z, psd_fn=None):
+    """Return the optimal SNR of a signal.
+
+    :param m1_intrinsic: The source-frame mass 1.
+
+    :param m2_intrinsic: The source-frame mass 2.
+
+    :param z: The redshift.
+
+    :param psd_fn: A function that returns the detector PSD at a given
+      frequency (default is early aLIGO high sensitivity, defined in
[…]

### 2023-06-02 population-genrator/weighting.py

@@ -1,0 +1,41 @@
+#! /usr/bin/env python
+import argparse
+import h5py
+import vt
+import numpy as np
+import os
+
+parser = argparse.ArgumentParser()
+parser.add_argument("--dat", type=str,default=None)
+parser.add_argument("--vt", type=str,default=None)
+parser.add_argument("--n-out", type=int,default=100)
+opts = parser.parse_args()
+
+col_names = ["m1_source", "m2_source", "ecc"]
+
+# load in VT filte
+with h5py.File(opts.vt, "r") as VTs:
+    raw_interpolator = vt.interpolate_hdf5(VTs)
+
+# load in data 
+dat = np.genfromtxt(opts.dat,names=True) # should work
+dat_mass = np.c_[dat['m1_source'], dat['m2_source'],dat['ecc']]
+
+# compute VTs, which are weights for choices
+weights = raw_interpolator(dat_mass[:,:2])
+weights *= 1./np.sum(weights) # normalizes
+
+indexes_all = np.arange(len(dat_mass))
+downselected = np.random.choice(indexes_all, p=weights,size=opts.n_out)
+
+dat_mass = dat_mass[downselected]
+
+np.savetxt("weighted_population.dat", dat_mass,delimiter="\t", header="\t".join(col_names))
+
+output_dir = './weighted_events/'
+for i in range(dat_mass.shape[0]):
+    row = dat_mass[i,:]
+    data = np.repeat([row], 4000, axis=0)
+    output_file_path = os.path.join(output_dir, f'event_{i+1}.dat')
+    np.savetxt(output_file_path, data, delimiter="\t",header="\t".join(col_names))
+

### 2023-06-05 population-genrator/run_complete.sh
Removing old plots

@@ -11,3 +11,5 @@ python weighting.py --dat population.dat --vt vt_m1_m2.hdf5 --n-out 100
 
 #script for scaling
 rm -r scaled_events; mkdir scaled_events ; python scaling.py
+
+#moving the events in the desired directory

### 2023-06-05 paper/paper.tex

@@
 \section{Introduction}
-Later this section
+Big Picture
+\begin{itemize}
+    \item GWs are important
+    \item ecccentricity is a signature that can be measued with GWs
+    \item it is a unique signature of strong events
+\end{itemize}
+Why Population, smaller picture
+\begin{itemize}
+    \item population inference of eccentricity is important.
+    \item some events may have unabmoguis eccentricty (refs to PE, richard will send them), but many will have only faint signatures or none at all.
+    \item critical for astrophysoics to acces distribution of eccentricites and any subpopulation
+\end{itemize}
+Previous work
+\begin{itemize}
+    \item What other people have done with eccentricty, 
+    \item cite any preious work realted to eccentricty with population.
+    \item mostly it's real events or high mass evenets
+    \item what sort of eccentricy we expect to appear
+\end{itemize}
+\begin{itemize}
+    \item Explain,what and how you are doing here.
+    \item assess how well eccentricity of population  can be inferreed using low mass BHS where eccentricity measurement should be much more precise.
+    \item for simplicity using a proof of concept population model to emphasize eccentric inference
+\end{itemize}
 
-\section{Methods}
-\label{sec:methods}
-
-The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters.
-The intrinsic parameters mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $(e)$  are subject to the orbital evolution of the binary.
-The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.
+We discussed the various formation scenario of EBBHs in the previous section; each one has a distinct signature in the GW signals.
+Hence, it's trivial to observe the effect of each scenario or the mixed effect on the merging population.
+It leads us to make the population model, which considers each scenario and allows distinct merger rates for them.
 
@@
 
+\section{Methods}
+\label{sec:methods}
+
+The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters.
+The intrinsic parameters mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary.
+The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.
+
 \subsection{Hierarchical Bayesian Modeling (HBM)}
-
-We discussed the various formation scenario of EBBHs in the previous section; each one has a distinct signature in the GW signals.
-Hence, it's trivial to observe the effect of each scenario or the mixed effect on the merging population.
-It leads us to make the population model, which considers each scenario and allows distinct merger rates for them.
 
@@
 Now having the parameters of individual sources, we are ready to make a population inference.
-Following the Bayesian framework for inference, our first task is to find the likelihood of the population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ equivalent to the probability of the individual sources given the population parameter $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, \sigma_e)$ written as follows.
+Following the Bayesian framework for inference, our first task is to find the likelihood of the population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ equivalent to the probability of the individual sources given the population parameter $\Lambda$ written as follows.
 
@@
 \end{equation}
-
-where $\alpha$ is the power law index given in Eq.
-\ref{eq:plaw}, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum, and maximum masses of the binary components in the population respectively, and $\sigma_e$ is the orbital eccentricity distribution.
 
@@
 
-The soundest approach is to find the analytical posterior.
-However, most of the time, it's difficult to find the analytical solution to such a complicated equation involving multidimensional integrals.
-Therefore, a practical approach is to use the computational methods.
-Our study will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} method to find the posterior distribution of population parameters.
-This method draws samples (set of parameters $\alpha,\mathcal{R},m_{min},m_{max},\sigma_e$ that make one model) from the targeted distribution; in our case, it's the power law model given in Eq.
-\ref{eq:plaw}, then compares it with the given data (collection of individual events) and store the best-fit sample.
+These calculations are analytically intractable and must be performed numerically.
+Specifically, we will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} method to find the posterior distribution of population parameters.
+This method draws samples from the targeted distribution for $\Lambda$; in our case, it's the power law model given in Eq.
+\ref{eq:plawg}, then compares it with the given data (collection of individual events) and store the best-fit sample.
 We may iterate this as we need and store multiple sample values untill they converge.
@@
 Existing LVK instruments' sensitivity depends only on the mass; therefore, we will get the same VT calculated previously \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries.
-Hence, we briefly explain the calculations, but one may read Daniel Wysocki's article \cite{Dan_2019} for a deeper study.
+Hence, we briefly explain the calculations, see \cite{Dan_2019} for details.
 
@@
 However, our analysis used the pure truncated power law defined in \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model} and added the Gaussian eccentricity.
-This model computes the intrinsic probability of $m_1$, $m_2$, and $e$.
-We also assume that non-zero probability density only exists for $m_{min}<\leq m_2 \leq m_1 \leq m_{max}$ and for total mass $M_{max}=m_1+m_2 = 200 M_\odot$.
+This model computes the intrinsic probability of $m_1$, $m_2$, and $\epsilon$.
+We also assume that non-zero probability density only exists for $m_{min}\leq m_2 \leq m_1 \leq m_{max}$ and for total mass $M_{max}=m_1+m_2 = 200 M_\odot$.
 The condition of the total mass is only because of the limitations of the detectors towards the higher masses \cite{2016PRXAbbot_BBH_model}.
-The generalized form of the truncated power-law model with parameters $m_{min}$, $m_{max}$, $k_m$, $\alpha$, $\sigma_e$ and random variable $m_1$, $m_2$, and $e$ has the functional form in Eq.
+The generalized form of the truncated power-law model with parameters $\Lambda \equiv  (\alpha, \mathcal{R}, k_m, m_{min}, m_{max}, \sigma_\epsilon, M_{max})$ and random variable $m_1$, $m_2$, and $\epsilon$ has the functional form in Eq.
 \ref{eq:plawg} within provided mass limit.
@@
 \label{eq:plawg}
-p(m_1,m_2,e) = &C(\alpha,k_m,m_{min},m_{max},M_{max},e)  
-\nonumber \\ & \frac{(m_2/m_1)^{k_m}m_1^{-\alpha}}{(m_1-m_{min})\sigma_e} 
-\nonumber \\ &
-\sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
+p(m_1,m_2,\epsilon) = &C(\alpha,k_m,m_{min},m_{max},M_{max},\epsilon)  
+\nonumber \\ & \sqrt{\frac{2}{\pi}} \frac{(m_2/m_1)^{k_m} m_1^{-\alpha} e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon} 
 \end{align}
 
+where $\alpha$ is the power law index, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum, and maximum masses of the binary components in the population respectively, and $\sigma_e$ is the orbital eccentricity distribution.
 The Eq.
 \ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
-For our analysis, we defined the constant of integration as equal to $\int_A dm_1 dm_2 p(m_1,m_2) dm_1 dm_2 = 1$.
+For our analysis, we defined the constant of integration equals to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
 Our detectors are very sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$; therefore, we will use $k_m=0$ throughout the studies.
@@
 \label{eq:plaw}
-p(m_1,m_2,e) =   \frac{m_1^{-\alpha}}{(m_1-m_{min})\sigma_e}  
-\sqrt{\frac{2}{\pi}} e^{(-1/2)(e/\sigma_e)^2}
+p(m_1,m_2,\epsilon) =   \sqrt{\frac{2}{\pi}} \frac{ m_1^{-\alpha}  e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon}  
 \end{align}
@@
 
-We choose the fixed parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$ to generate the population of $N=100$ sources from the power model given in Eq.
+We choose fixed parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$ to generate a population of $N=100$ sources from the power model given in Eq.
 \ref{eq:plawg}.
 We have generated two different populations from the fiducial model to show the significance of the eccentricity on the population.
-The selection of all the parameters is the same for both populations instead of $\sigma_e$.
+The selection of all the parameters is the same for both populations instead of $\sigma_\epsilon$.
 This different eccentricity distribution in both models will show how the events can be missed by not including the eccentricity in our analysis.
@@
 \label{eq:scaling}
-M^{ecc} = \frac{M}{(1-\frac{157}{24}e^2)^{3/5}}
+M^{ecc} = \frac{M}{(1-\frac{157}{24}\epsilon^2)^{3/5}}
 \end{align}

### 2023-06-06 paper/paper.tex
Updates from Overleaf

@@
 \ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
-For our analysis, we defined the constant of integration equals to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
+For our analysis, we defined the constant of integration equal to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
 Our detectors are very sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$; therefore, we will use $k_m=0$ throughout the studies.
@@
 
-We choose fixed parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$ to generate a population of $N=100$ sources from the power model given in Eq.
-\ref{eq:plawg}.
-We have generated two different populations from the fiducial model to show the significance of the eccentricity on the population.
-The selection of all the parameters is the same for both populations instead of $\sigma_\epsilon$.
-This different eccentricity distribution in both models will show how the events can be missed by not including the eccentricity in our analysis.
+We have created four different populations by fixing the
+power law parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$.
+Each population has the same values for $\alpha$, $m_{min}$, and $m_{max}$ except $\sigma_\epsilon$.
+We selected four values of $\sigma_\epsilon = 0.05,0.1,0.15,0.2$ and each synthetic population contains $N=10000$ sources.
+After getting those populations, we find the probability for each event to be detected by computing the VT of each source.
+Finally, we will randomly pick $100$ sources for each population to perform our analysis.
 
-Furthermore, to compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we have created two more different populations by scaling the existing populations using the Eq.
+Our first and second weighted populations with $\sigma_\epsilon=0.05, 0.1$ are given in \ref{fig:pop3d0.05_0.1}.
+
+\begin{figure*}
+
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.05.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.1.png}
+\caption{\label{fig:pop3d0.05_0.1} Synthetic Population of EBBHs.
+The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represetns the population for $\sigma_\epsilon=0.1$}
+
+\end{figure*}
+
+\begin{figure*}
+
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.15.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.2.png}
+\caption{\label{fig:pop3d0.15_0.2} Synthetic Population of EBBHs.
+The left figure represents the population for $\sigma_\epsilon=0.15$ and right figure represetns the population for $\sigma_\epsilon=0.2$}
+
+\end{figure*}
+
+\subsection{Scaled Population}
+
+To compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we created four more populations by scaling the existing populations using the Eq.
 \ref{eq:scaling} defined in \cite{2021_scaling_paper}.
 The Eq.
-\ref{eq:scaling} removes the eccentric component from the population, scales the masses, and eventually, we get a non-eccentric population to compare with the existing eccentric population.
+\ref{eq:scaling} removes the eccentric component from the population, scales the masses, and eventually, we get non-eccentric populations to compare with the existing eccentric population.
 \begin{align}
@@
 
-\subsubsection{EBBH and NEBBH Population with $0\leq \sigma_e \leq 0.5$}
- The first population has the Gaussian distribution bounded between $0\leq \sigma_e \leq 0.5$ given in Fig. \ref{fig:pop3d05}.
+The first population is given in 
+given in Fig. \ref{fig:pop3d05}.
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d05.png}
-\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs for $0\leq \sigma_e \leq 0.5$}}
-\end{figure}
+\begin{figure*}
 
+\includegraphics[width=0.45\textwidth]{paper/figures/pop0.05.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop0.1.png}
+\caption{\label{fig:pop_0.05_0.1} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$.
+The right figure shows the same plot but for $\sigma_\epsilon=0.1$}
+
+\end{figure*}
+
+\begin{figure*}
+
+\includegraphics[width=0.45\textwidth]{paper/figures/pop0.15.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop0.2.png}
+\caption{\label{fig:pop_0.15_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.15$.
+The right figure shows the same plot but for $\sigma_\epsilon=0.2$}
+
+\end{figure*}
 We have the total $100$ events with a minimum and maximum primary mass of $17.73 M_\odot$ and $49.14 M_\odot$, respectively.
 Similarly, we have the secondary minimum and maximum masses $10.11 M_\odot$ and $46.68 M_\odot$ given in Fig \ref{fig:pop2d05}.
-
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05.png}
-\caption{\label{fig:pop2d05}\textbf{Mass distribution of EBBHs for $0\leq \sigma_e \leq 0.5$}}
-\end{figure}
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05scl.png}
-\caption{\label{fig:pop2dscal05}\textbf{Scaled Mass distribution of NEBBHs for  $0\leq \sigma_e \leq 0.5$}}
-\end{figure}
-
 The significant effect after the scaling is not only the mass shift, but one may notice in Fig.
@@
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d05diff.png}
-\caption{\label{fig:pop2d05diff}\textbf{Total Mass of Binaries for EBBHs and NEBBHs for $0\leq \sigma_e \leq 0.5$:} The blue dots show the EBBHs, the orange shows scaled NEBBHs, and green color shows the increasing difference between those masses with eccentricity}
-\end{figure}
-
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss05.png}
-\caption{\label{fig:population05}\textbf{The relation between eccentricity and mass loss}}
-\end{figure}
-
-\subsubsection{EBBH and NEBBH Population with $0\leq \sigma_e \leq 0.38$}
- The second population has the Gaussian distribution bounded between $0\leq \sigma_e \leq 0.38$ given in Fig. \ref{fig:pop3d038}.
-
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d038.png}
-\caption{\label{fig:pop3d038}\textbf{3D-Synthetic Population of EBBHs for $0\leq \sigma_e \leq 0.38$}}
-\end{figure}
+The second population is given in Fig. \ref{fig:pop3d038}.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038.png}
-\caption{\label{fig:pop2d038}\textbf{Mass distribution of EBBHs with $0\leq \sigma_e \leq 0.38$}}
-\end{figure}
+The second population is given in Fig. \ref{fig:pop3d038}.
 
@@
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038scl.png}
-\caption{\label{fig:pop2d038scl}\textbf{Scaled mass distribution of NEBBHs for $0\leq \sigma_e \leq 0.38$}}
-\end{figure}
+\section{Results}
 
-Selecting a reduced range of $\sigma_e$ leads to avoiding the missing events.
-Although mass gets scaled, we did not see any missing events in Fig.
-\ref{fig:pop2d038scl}.
-The mass scaling effect is visible in Fig.
-\ref{fig:pop2d038diff} and reflects the importance of eccentricity.
+we can discuss the corner plots here.
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/pop2d038diff.png}
-\caption{\label{fig:pop2d038diff}\textbf{Total Mass of Binaries for EBBHs and NEBBHs for $0\leq \sigma_e \leq 0.38$:} The blue dots show the EBBHs, the orange shows scaled NEBBHs, and green color shows the increasing difference between those masses with eccentricity}
-\end{figure}
+\begin{figure*}
 
-\begin{figure}[H]
-\includegraphics[width=0.45\textwidth]{paper/figures/massloss038.png}
-\caption{\label{fig:population038}\textbf{The relation between eccentricity and mass loss}}
-\end{figure}
+\includegraphics[width=0.95\textwidth]{paper/figures/corner0.05.png}
+\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs}}
 
-\section{Inference of Synthetic BBH Population}
+\end{figure*}
+Same population.
+\begin{figure*}
+
+\includegraphics[width=0.95\textwidth]{paper/figures/corner0.05scl.png}
+\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs}}
+
+\end{figure*}
 

### 2023-06-09 paper/paper.tex

@@
 
-Our first and second weighted populations with $\sigma_\epsilon=0.05, 0.1$ are given in \ref{fig:pop3d0.05_0.1}.
+Our weighted populations with $\sigma_\epsilon=0.05, 0.1$ and $\sigma_\epsilon=0.15, 0.2$ are given in Fig. \ref{fig:pop3d0.05_0.1} and Fig. \ref{fig:pop3d0.15_0.2} respectively.
+
+\begin{table*}[]
+    \centering
+    \begin{tabular}{|ccccccc|}
+        \hline
+        $\sigma_\epsilon$ & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
+        0.05 & 16.81 & 49.81 & 10.8 & 46.82 & 0.01 & 0.05\\ \hline
+        0.1 & 19.03 & 49.93 & 10.1 & 48.64 & 0.05 & 0.1\\ \hline
+        0.15 & 10.04 & 49.9 & 10.01 & 49.2 & 0.1 & 0.15\\ \hline
+        0.2 & 19.6 & 49.97 & 10.73 & 48.39 & 0.15 & 0.2\\ \hline
+    \end{tabular}
+    \caption{Weighted Population Properties}
+    \label{tab:pop_prop}
+\end{table*}
 
@@
 \caption{\label{fig:pop3d0.15_0.2} Synthetic Population of EBBHs.
-The left figure represents the population for $\sigma_\epsilon=0.15$ and right figure represetns the population for $\sigma_\epsilon=0.2$}
+The left figure represents the population for $\sigma_\epsilon=0.15$ and right figure represents the population for $\sigma_\epsilon=0.2$}
 
@@
 
-The first population is given in 
-given in Fig. \ref{fig:pop3d05}.
+\begin{table*}[]
+    \centering
+    \begin{tabular}{|ccccccc|}
+        \hline
+        $\sigma_\epsilon$ & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
+        0.05 & 16.68 & 49.64 & 10.77 & 46.75 & 0.01 & 0.05\\ \hline
+        0.1 & 18.95 & 49.51 & 9.89 & 48.33 & 0.05 & 0.1\\ \hline
+        0.15 & 9.13 & 49.55 & 9.1 & 49.11 & 0.1 & 0.15\\ \hline
+        0.2 & 18.2 & 49.79 & 10.3 & 47.08 & 0.15 & 0.2\\ \hline
+    \end{tabular}
+    \caption{Scaled Population Properties}
+    \label{tab:popscl_prop}
+\end{table*}
 
@@
 \ref{fig:pop2d05diff}, and we also notice that sources start missing after $e \gtrsim 0.38$.
-
-The second population is given in Fig. \ref{fig:pop3d038}.
-
-We have the total $100$ events with a minimum and maximum primary mass of $15.12 M_\odot$ and $49.98 M_\odot$, respectively.
-Similarly, we have the secondary minimum and maximum masses $10.12 M_\odot$ and $46.01 M_\odot$ given in Fig \ref{fig:pop2d038}.
-
-The second population is given in Fig. \ref{fig:pop3d038}.
-
-Similarly, to get the non-eccentric population, we apply the Eq.
-\ref{eq:scaling} on the population given in Fig.
-\ref{fig:pop2d038} and get the new population with scaled masses.
-We get the new minimum and maximum values for the primary and secondary masses in Fig.
-\ref{fig:pop2d038scl}.
 

### 2023-06-09 paper/paper.tex
Updates from Overleaf

@@
 
-\begin{figure*}
+\begin{figure}[H]
 
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.05.png}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.1.png}
-\caption{\label{fig:pop3d0.05_0.1} Synthetic Population of EBBHs.
-The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represetns the population for $\sigma_\epsilon=0.1$}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.2.png}
+\caption{\label{fig:pop3d0.2} Synthetic Population of EBBHs. The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represetns the population for $\sigma_\epsilon=0.1$}
 
-\end{figure*}
-
-\begin{figure*}
-
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.15.png}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.2.png}
-\caption{\label{fig:pop3d0.15_0.2} Synthetic Population of EBBHs.
-The left figure represents the population for $\sigma_\epsilon=0.15$ and right figure represents the population for $\sigma_\epsilon=0.2$}
-
-\end{figure*}
+\end{figure}
 
@@
 \includegraphics[width=0.45\textwidth]{paper/figures/pop0.05.png}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop0.1.png}
-\caption{\label{fig:pop_0.05_0.1} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$.
-The right figure shows the same plot but for $\sigma_\epsilon=0.1$}
-
-\end{figure*}
-
-\begin{figure*}
-
-\includegraphics[width=0.45\textwidth]{paper/figures/pop0.15.png}
 \includegraphics[width=0.45\textwidth]{paper/figures/pop0.2.png}
-\caption{\label{fig:pop_0.15_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.15$.
+\caption{\label{fig:pop_0.05_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$.
 The right figure shows the same plot but for $\sigma_\epsilon=0.2$}
@@
 \end{figure*}
-We have the total $100$ events with a minimum and maximum primary mass of $17.73 M_\odot$ and $49.14 M_\odot$, respectively.
-Similarly, we have the secondary minimum and maximum masses $10.11 M_\odot$ and $46.68 M_\odot$ given in Fig \ref{fig:pop2d05}.
 
@@
 
-\includegraphics[width=0.95\textwidth]{paper/figures/corner0.05.png}
-\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs}}
+\includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.05.png}
+\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHS with $\sigma_\epsilon=0.05$}}
 
@@
 
-\includegraphics[width=0.95\textwidth]{paper/figures/corner0.05scl.png}
-\caption{\label{fig:pop3d05}\textbf{3D-Synthetic Population of EBBHs}}
+\includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.2.png}
+\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHs with $\sigma_\epsilon=0.2$}}
 

### 2023-06-09 paper/paper.tex
Inference tables
Updates from Overleaf

@@
     \centering
-    \begin{tabular}{|ccccccc|}
+    \begin{tabular}{c|cccccc}
         \hline
-        $\sigma_\epsilon$ & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
-        0.05 & 16.81 & 49.81 & 10.8 & 46.82 & 0.01 & 0.05\\ \hline
-        0.1 & 19.03 & 49.93 & 10.1 & 48.64 & 0.05 & 0.1\\ \hline
-        0.15 & 10.04 & 49.9 & 10.01 & 49.2 & 0.1 & 0.15\\ \hline
-        0.2 & 19.6 & 49.97 & 10.73 & 48.39 & 0.15 & 0.2\\ \hline
+        Populations & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
+        First 1 & 16.81 & 49.81 & 10.8 & 46.82 & 0.01 & 0.05\\ \hline
+        Second 2 & 19.03 & 49.93 & 10.1 & 48.64 & 0.05 & 0.1\\ \hline
+        Third 3 & 10.04 & 49.9 & 10.01 & 49.2 & 0.1 & 0.15\\ \hline
+        Forth 4 & 19.6 & 49.97 & 10.73 & 48.39 & 0.15 & 0.2\\ \hline
     \end{tabular}
@@
     \centering
-    \begin{tabular}{|ccccccc|}
+    \begin{tabular}{c|cccc}
         \hline
-        $\sigma_\epsilon$ & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
-        0.05 & 16.68 & 49.64 & 10.77 & 46.75 & 0.01 & 0.05\\ \hline
-        0.1 & 18.95 & 49.51 & 9.89 & 48.33 & 0.05 & 0.1\\ \hline
-        0.15 & 9.13 & 49.55 & 9.1 & 49.11 & 0.1 & 0.15\\ \hline
-        0.2 & 18.2 & 49.79 & 10.3 & 47.08 & 0.15 & 0.2\\ \hline
+        Populations & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ \\ \hline
+        First & 16.68 & 49.64 & 10.77 & 46.75 \\ \hline
+        Second & 18.95 & 49.51 & 9.89 & 48.33 \\ \hline
+        Third & 9.13 & 49.55 & 9.1 & 49.11  \\ \hline
+        Forth & 18.2 & 49.79 & 10.3 & 47.08  \\ \hline
     \end{tabular}
@@
 
-\section{Results}
+\section{Population Inference}
 
+\begin{table*}[]
+    \centering
+    \begin{tabular}{c|ccccc}
+        \hline
+       Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
+      First& $1.88^{+0.18}_{-0.59}$ & $-1.41^{+1.29}_{-0.97}$ & $5.47^{+1.74}_{-1.74}$ & $69.67^{+10.20}_{-10.75}$ & $0.05^{+0.35}_{-0.02}$ \\ \hline
+      Second & $1.89^{+0.15}_{-0.56}$ & $-0.99^{+1.01}_{-0.85}$ & $5.52^{+1.75}_{-1.80}$ & $69.74^{+10.64}_{-10.95}$ & $0.07^{+0.33}_{-0.02}$ \\ \hline
+      Third & $1.90^{+0.14}_{-0.53}$ & $-0.94^{+0.94}_{-0.82}$ & $5.63^{+2.02}_{-1.92}$ & $69.73^{+10.92}_{-11.07}$ & $0.10^{+0.30}_{-0.02}$  \\ \hline
+      Forth & $1.91^{+0.18}_{-0.55}$ & $-1.89^{+1.44}_{-1.00}$ & $5.48^{+1.87}_{-1.90}$ & $69.26^{+11.43}_{-10.96}$ & $0.14^{+0.29}_{-0.03}$  \\ \hline
+    \end{tabular}
+    \caption{Population Inference for EBBHs}
+    \label{tab:inference_EBBHS}
+\end{table*}
+
+\begin{table*}[]
+    \centering
+    \begin{tabular}{c|cccc}
+        \hline
+        Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ \\ \hline
+      First & $1.89^{+0.17}_{-0.43}$ & $-1.68^{+1.08}_{-0.80}$ & $5.49^{+2.38}_{-2.41}$ & $68.83^{+14.66}_{-13.13}$  \\ \hline
+      Second & $1.89^{+0.13}_{-0.43}$ & $-0.90^{+0.72}_{-0.68}$ & $5.40^{+2.29}_{-2.40}$ & $68.61^{+14.69}_{-13.33}$  \\ \hline
+      Third & $1.89^{+0.13}_{-0.44}$ & $-0.57^{+0.65}_{-0.67}$ & $5.72^{+2.48}_{-2.40}$ & $68.74^{+14.30}_{-12.99}$   \\ \hline
+      Forth & $1.89^{+0.14}_{-0.45}$ & $-1.20^{+0.81}_{-0.71}$ & $5.32^{+2.33}_{-2.38}$ & $68.83^{+14.51}_{-12.95}$  \\ \hline
+    \end{tabular}
+    \caption{Population Inference for NEBBHs}
+    \label{tab:inference_NEBBHS}
+\end{table*}
 we can discuss the corner plots here.
@@
 \includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.05.png}
-\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHS with $\sigma_\epsilon=0.05$}}
+\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHS for $\sigma_\epsilon=0.05$}}
 
@@
 \includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.2.png}
-\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHs with $\sigma_\epsilon=0.2$}}
+\caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHs for $\sigma_\epsilon=0.2$}}
 

### 2023-06-11 new_powerlaw_mcmc.py
Adding Gaussian Uncertainity in the injections

@@ -80,7 +80,7 @@ def expval_mc(
     )
 
     if return_err:
-        return rate*I*124, err_abs, err_rel
+        return rate*I, err_abs, err_rel
     else:
         return rate*I*124
 

### 2023-06-11 original_pop_code/original_powerlaw_mcmc.py

@@ -70,7 +70,7 @@ def expval_mc(
     if return_err:
         return rate*I, err_abs, err_rel
     else:
-        return rate*I
+        return rate*I*124
 
 
 def VT_interp(m1_m2, raw_interpolator, **kwargs):

### 2023-06-11 population-genrator/genrating.py

@@ -9,7 +9,7 @@ import syn_pop_prob
 
 # syn_pop_prob.joint_rvs(N,alpha,m_min,m_max,M_max=m1+m2,sigma_e)
 
-array = syn_pop_prob.joint_rvs(10000,-2,10,50,100,0.2)
+array = syn_pop_prob.joint_rvs(100000,-3,10,50,100,0.2)
 
 
 col_names = ["m1_source", "m2_source", "ecc"]

### 2023-06-11 population-genrator/real_PE.py

@@ -1,0 +1,44 @@
+import os
+import numpy as np
+from scipy.stats import truncnorm
+
+# Specify the path to the folder containing the .dat files
+folder_path = './weighted_events'
+
+# Specify the path to the folder where the copies will be saved
+output_folder = './rPE_events'
+
+# Create the output folder if it doesn't exist
+os.makedirs(output_folder, exist_ok=True)
+
+# Set the number of copies to generate
+num_copies = 4000
+
+# Set the standard deviation factor for Gaussian uncertainty
+std_dev_factor = 25
+col_names = ["m1_source", "m2_source","ecc"]
+# Iterate over all .dat files in the folder
+for filename in os.listdir(folder_path):
+    if filename.endswith(".dat"):
+        file_path = os.path.join(folder_path, filename)
+
+        # Load the data from the .dat file
+        data = np.loadtxt(file_path)
+
+        # Extract the values from each column
+        m1 = data[0, 0]
+        m2 = data[0, 1]
+        ecc = data[0, 2]
+        # Initialize an empty array to store the copies
+        copies = np.zeros((num_copies, 3))
+        # Generate copies with Gaussian uncertainty for each column
+        for i in range(num_copies):
+            copies[i, 0] = np.random.normal(m1, std_dev_factor)
+            copies[i, 1] = np.random.normal(m2, std_dev_factor)
+            copies[i, 2] = truncnorm.rvs(0,1,loc=ecc,scale=0.15, size=1)
+        copies[0,:] = m1,m2,ecc
+        # Create the output file path and save the copies
+        output_filename = os.path.splitext(filename)[0] + ".dat"
+        output_path = os.path.join(output_folder, output_filename)
+        np.savetxt(output_path, copies, delimiter='\t',header="\t".join(col_names))
+

### 2023-06-11 population-genrator/run_complete.sh

@@ -1,5 +1,7 @@
 # Script for population genration
 rm population.dat ;
+rm scaled_population.dat
+rm weighted_population.dat
 rm -r __pycache__ ;
 python genrating.py
 
@@ -8,6 +10,12 @@ rm -r __pycache__ ;
 rm -r weighted_events ;
 mkdir weighted_events;
 python weighting.py --dat population.dat --vt vt_m1_m2.hdf5 --n-out 100
+
+#Adding Gaussian Uncertainities in the injections
+
+rm -r rPE_events;
+mkdir rPE_events;
+python real_PE.py;
 
 #script for scaling
 rm -r scaled_events; mkdir scaled_events ; python scaling.py

### 2023-06-11 population-genrator/scaling.py

@@ -2,7 +2,7 @@ import numpy as np
 import os
 
 # Directory containing the input files
-input_dir = './weighted_events/'
+input_dir = './rPE_events/'
 
 # Directory to save the output files
 output_dir = './scaled_events/'
@@ -16,10 +16,11 @@ for i in range(1, 101):
     data = np.loadtxt(file_path)
 
     # Extract the columns
-    col1 = data[:, 0]
-    col2 = data[:, 1]
-    col3 = data[:, 2]
+    col1 = data[:,0]
+    col2 = data[:,1]
+    col3 = data[:,2] # need to discuss that should I scale with same ecc or with error
     scale_factor = ((1-(157/24)*col3**2)**(3/5))
+    print(col3)
     # Scaling equation: scaled_value = column_value * scale_factor
     scaled_col1 = col1 * scale_factor
     scaled_col2 = col2 * scale_factor
@@ -73,9 +74,9 @@ print(f"Total files with NaN or zero values: {count}")
 data01 = np.loadtxt('weighted_population.dat')
 
 # Extract the columns
-col01 = data01[:, 0]
-col02 = data01[:, 1]
-col03 = data01[:, 2]
+col01 = data01[:,0]
+col02 = data01[:,1]
+col03 = data01[:,2]
 scale_factor01 = ((1-(157/24)*col03**2)**(3/5))
 
 

### 2023-06-26 paper/paper.tex
Updates from Overleaf

@@
 \title{Population Inference of Non-Spinning Eccentric Binary Black Holes}
-\author{M. Zeeshan}
+%\author{R. O'Shaughnessy}
 \affiliation{\RIT}
-\author{R. O'Shaughnessy}
+%\author{M. Zeeshan}
+
 \affiliation{\RIT}
@@
     \item GWs are important
-    \item ecccentricity is a signature that can be measued with GWs
+    \item eccentricity is a signature that can be measured with GWs
     \item it is a unique signature of strong events
@@
     \item population inference of eccentricity is important.
-    \item some events may have unabmoguis eccentricty (refs to PE, richard will send them), but many will have only faint signatures or none at all.
-    \item critical for astrophysoics to acces distribution of eccentricites and any subpopulation
+    \item Some events may have unabmoguis eccentricity (refs to PE, Richard will send them), but many will have only faint signatures or none at all.
+    \item critical for astrophysics to acces distribution of eccentricities and any subpopulation
 \end{itemize}
@@
 \begin{itemize}
-    \item What other people have done with eccentricty, 
-    \item cite any preious work realted to eccentricty with population.
-    \item mostly it's real events or high mass evenets
-    \item what sort of eccentricy we expect to appear
+    \item What other people have done with eccentricity, 
+    \item cite any previous work related to eccentricity with population.
+    \item mostly it's real events or high mass events
+    \item what sort of eccentricity we expect to appear
 \end{itemize}
 \begin{itemize}
-    \item Explain,what and how you are doing here.
-    \item assess how well eccentricity of population  can be inferreed using low mass BHS where eccentricity measurement should be much more precise.
+    \item Explain, what and how you are doing here.
+    \item assesses how well the eccentricity of the population  can be inferred using low mass BHS where eccentricity measurement should be much more precise.
     \item for simplicity using a proof of concept population model to emphasize eccentric inference
@@
 
-Most of the eccentric binary black hole (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary have mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
+Most of the eccentric binary black holes (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
 \cite{mass_ecc_limit_2018}.
@@
     \end{tabular}
-    \caption{Weighted Population Properties}
+    \caption{Population Properties of EBBHs}
     \label{tab:pop_prop}
@@
 
-\subsection{Scaled Population}
+\subsection{Synthetic parameter inference with eccentricity}
 
@@
     \end{tabular}
-    \caption{Scaled Population Properties}
+    \caption{Population Properties of NEBBHs}
     \label{tab:popscl_prop}

### 2023-06-27 paper/paper.tex
Rewriting the method and synthetic population section

@@
 \begin{document}
-
+\renewcommand{\arraystretch}{1.5}
 \title{Population Inference of Non-Spinning Eccentric Binary Black Holes}
@@
 The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters.
-The intrinsic parameters mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary.
+The intrinsic parameters such as mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary.
 The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.
@@
 
-Once we have the population models, our next obstacle is constraining them with observational data.
-While constraining the model, we must consider the sensitivity of LIGO-VIRGO-KAGRA (LVK) and uncertainty in each detection.
-Therefore, the HBM is a good choice to constrain the population model.
+We use HBM to constrain the population models with gravitational wave signals. It is also a good choice to consider the sensitivity of LIGO-VIRGO-KAGRA (LVK) and uncertainty in each detection.
 
-In the HBM, we have the N number of discrete detection, in our case, from LVK.
-Those detections provide merger data denoted as $d_1,d_2,d_3,...,d_N$ where each $d_i$ shows the BBH.
-Each $d_i$ or merger has mass, spin, and eccentricity properties.
+In HBM, we have the N number of discrete detections.
+Those detections provide merger data denoted as $d_1,d_2,d_3,...,d_N$ where each $d_i$ shows a BBH merger.
+Each $d_i$ has mass, spin, and eccentricity properties.
 These properties, often called parameters, are denoted by $\lambda_1,\lambda_2,\lambda_3,...,\lambda_i$.
 Each parameter has its uncertainty, and we express it by the probability of the data given the parameter value.
-We also refer to it as the likelihood function $\mathcal{L}(\lambda)=p(d|\lambda)$ of one event/detection.
-The given value of $\lambda$ to calculate the likelihood is obtained from the mathematical model of the waveform.
-Once you have a likelihood function, you can use a prior (usually, a uniform) to find the posterior probability using the Bayes theorem.
+We also refer to it as the likelihood function $\mathcal{L}(\lambda)=p(d|\lambda)$ of one event.
+A given value of $\lambda$ is obtained from a mathematical model commonly known as a waveform to calculate a likelihood.
+Once you have a likelihood function, you may use a uniform prior or any other prior to find a posterior probability using the Bayes theorem as given in Eq.
+\ref{eq:Bayes_ind}
 
@@
 \label{eq:Bayes_ind}    
-p(\lambda|d) \propto p(d|\lambda) p(\lambda)
+p(\lambda|d) \propto p(d|\lambda) p(\lambda).
 \end{equation}
@@
 This posterior probability will constrain the properties of each binary, such as mass, spin, and eccentricity.
-We may get the inference of those parameters using Rapid parameter Inference on gravitational wave sources via Iterative Fitting (RIFT): an open source code for parameter estimation of the binary sources \cite{rift_2018}.
+We may infer those parameters using rapid parameter inference on gravitational wave sources via iterative fitting (RIFT): an open source code for parameter estimation (PE) of the binary sources \cite{rift_2018}.
 
@@
 
-Now having the parameters of individual sources, we are ready to make a population inference.
-Following the Bayesian framework for inference, our first task is to find the likelihood of the population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ equivalent to the probability of the individual sources given the population parameter $\Lambda$ written as follows.
+Now having the PE of individual sources, we will follow the Bayesian framework for population inference.
+The likelihood of a population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ is equivalent to the probability of the individual sources given the population parameter $\Lambda$ is written as follows.
 
@@
 
-The second step is to find the posterior probability, and we can do this by using the above likelihood in the Bayes theorem as given below.
+We will use the likelihood given in Eq. \ref{eq:likelihood_pop} in the Bayes theorem given in Eq. \ref{eq:Bayes} to find the posterior probability.
 
@@
 \label{eq:Bayes}    
-p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)}
+p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)},
 \end{equation}
 
-where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or sometime known as evidence.
+where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or also known as evidence.
 
-To conduct our specific analysis for mass and eccentricity distribution, we will use the inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
+To conduct our analysis for mass and eccentricity distribution, we will use the inhomogeneous Poisson process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
 
@@
 
-Where $\mu(\mathcal{R},\Lambda)$ is the expected number of detection under the given population parametrization $\Lambda$ with the overall rate $\mathcal{R}$.
+where $\mu(\mathcal{R},\Lambda)$ is the expected number of detection under the given population parametrization $\Lambda$ with the overall rate $\mathcal{R}$.
 $\ell_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
-Finally, we will get our posterior as $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$.
+Finally, we will get our posterior as $p(\mathcal{R},\Lambda | D)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$ by choosing a prior $p(\mathcal{R},\Lambda)$.
 
 These calculations are analytically intractable and must be performed numerically.
-Specifically, we will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} method to find the posterior distribution of population parameters.
-This method draws samples from the targeted distribution for $\Lambda$; in our case, it's the power law model given in Eq.
-\ref{eq:plawg}, then compares it with the given data (collection of individual events) and store the best-fit sample.
+Specifically, we will use Goodman and Weare's affine invariant Markov chain Monte Carlo (MCMC) \cite{mcmc_paper} to find the posterior distribution of population parameters.
+This method draws samples from the targeted distribution for $\Lambda$, in our case, it's the power law model given in Eq.
+\ref{eq:plawg}, then compares it with the given data (collection of individual events) and stores the best-fit sample.
 We may iterate this as we need and store multiple sample values untill they converge.
 This method is in a Python package called EMCEE \cite{emcee_paper}.
-
-%\begin{itemize}
-%    \item We have the parameter estimation from RIFT or any other method. It will give us a distribution for each parameter, and we may take the mean value of it.
- %   \item Our first task is to find the likelihood function for the population parameters: such as mass and eccentricity.
-%    \item We denote population parameters with \\ $\Lambda \equiv  (\alpha, \mathcal{R}, m_{min}, m_{max}, e)$.
-%    \item uncertainty in $\Lambda$ is a likelihood which can be written as $\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)$
-%    \item By applying the Bayes theorem, we can turn above likelihood into posterior probability of a parameter.
-%\begin{equation}
-%\label{eq:Bayes}    
-%p(\Lambda|d_1,d_2,...,d_N)= \frac{p(\Lambda)p(d_1,d_2,...,d_N|\Lambda)}{p(d_1,d_2,...,d_N)}
-%\end{equation}
-
-%    \item where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization/evidence, and $p(d_1,d_2,d_3,...,d_N|\Lambda)$ is likelihood.
-
-%   \item We will use inhomogeneous Poisson Process scaled by rate $\mathcal{R} = \frac{dN}{dtdV_c}$ and parameterize by $\Lambda$ to find the likelihood  $\mathcal{L}(\mathcal{R},\Lambda)\equiv p(D|\mathcal{R},\Lambda)$ of an astrophysical population given the merger rate and parameter $\Lambda$.
-%    {\color{red}I think these values of $\Lambda$ are provided by the power law model}. \\
-%\begin{equation}
-%\label{eq: likelihood}
-%\mathcal{L}(\mathcal{R},\Lambda) \propto e^{-%\mu(\mathcal{R},\Lambda)}\prod_{n=1}^N\int d\lambda %\ell_n(\lambda) \mathcal{R} p(\lambda|\Lambda)    
-%\end{equation}
-
-%    \item where $\mu(\mathcal{R},\Lambda)$ are the expected number of detections under the given population parametrization $\Lambda$ with overall rate $\mathcal{R}$.
-$l_n(\lambda)=p(d_n|\lambda)$ is the likelihood of the data $d_n$ given binary parameter.
- %   \item now we can find the posterior by Baye's theorem. \\
- %   $p(\mathcal{R},\Lambda)\propto p(\mathcal{R},\Lambda)  \mathcal{L}(\mathcal{R},\Lambda)$
-%    \item calculation of this posterior is computationally expensive, therefore, we will use MCMC method to find the posterior distribution.
-
-%\end{itemize}
 
@@
 \label{eq:mu}
-  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda  
+  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda ,
 \end{equation}
@@
 
-%\begin{itemize}
-%    \item Estimation of VT (sensitivity of the LVK) only depend on the mass.
-For the sake of simplicity {\color{red}(physics reason)}, we consider only non-spinning, non-eccentric and nonprecessing BH Binaries.
-%    \item Sources are detected if SNR is greater than 8.
-%    \item V is characteristic volume in which source can be detected by LVK. It is orientation averaged sensitive 3-volume.
-%\begin{equation}
-%\label{eq:volume}
-%V(\lambda) = \int P((<D(z))/D_n(\lambda))\frac{dV_c}{dz}\frac{dz}{1+z}
-%\end{equation}    
-%    \item where $D(z)$ is the luminosity distance for redshift z, $V_c$ is the comoving volume.
-
-%    \item to compute the $\mu(\mathcal{R},\Lambda)$, we use previously defined Volume in $Gpc^{-3}yr^{-1}$ and observing at this sensitivity for the time $T$, the average number of detection would be as follows: \\
-%\begin{equation}
-%\label{eq:mu}
-%  \mu(\mathcal{R},\Lambda) = \int(VT)\lambda \mathcal{R}p(\lambda|\Lambda)d\lambda  
-%\end{equation}
-%    \item where $p(\lambda|\Lambda)$ is the probability density function for a random binary in the universe to have intrinsic paramter $\lambda$.
-Keep in mind that $\lambda$ is equal to all the 15 parameters of a binary.
-%\end{itemize}
-
 \subsection{Power law Model}
 There are various weak and pure phenomenological population models proposed in previous studies \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model,2018talbot_bbh_model}.
-However, our analysis used the pure truncated power law defined in \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model} and added the Gaussian eccentricity.
+However, our analysis used the pure truncated power law defined in \cite{2016PRXAbbot_BBH_model,2017FishBach_BBH_model} and Gaussian eccentricity.
 This model computes the intrinsic probability of $m_1$, $m_2$, and $\epsilon$.
@@
 p(m_1,m_2,\epsilon) = &C(\alpha,k_m,m_{min},m_{max},M_{max},\epsilon)  
-\nonumber \\ & \sqrt{\frac{2}{\pi}} \frac{(m_2/m_1)^{k_m} m_1^{-\alpha} e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon} 
+\nonumber \\ & \sqrt{\frac{2}{\pi}} \frac{(m_2/m_1)^{k_m} m_1^{-\alpha} e^{-(\epsilon/\sqrt{2}\sigma_\epsilon)^2}}{(m_1-m_{min})\sigma_\epsilon},
 \end{align}
 
-where $\alpha$ is the power law index, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum, and maximum masses of the binary components in the population respectively, and $\sigma_e$ is the orbital eccentricity distribution.
+where $\alpha$ is the power law index, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum and maximum masses of the binary components in the population, and $\sigma_e$ is the orbital eccentricity distribution.
 The Eq.
 \ref{eq:plawg} represent a truncated power law for primary mass $m_1$ with index $-\alpha$ and conditional power law distribution $p(m_2|m_1)$ for secondary mass $m_2$ using simple power law, and Gaussian distribution for orbital eccentricity $e$.
-For our analysis, we defined the constant of integration equal to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
-Our detectors are very sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$; therefore, we will use $k_m=0$ throughout the studies.
-Therefore, we have our reduced form of the truncated power law in Eq.
+For our analysis, we defined a constant of integration equal to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
+Our detectors are sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$, therefore, we will use $k_m=0$ throughout the studies.
+As a result, we have our reduced form of the truncated power law in Eq.
 \ref{eq:plaw}
@@
 
-\subsection{Fiducial Model for Synthetic Population}
+\section{Synthetic Population}
 
-We have created four different populations by fixing the
-power law parameters $\alpha = -2$, $m_{min} = 10$, $m_{max}=50$.
+We have created four synthetic populations by choosing the power law parameters $\alpha = -1$, $m_{min} = 10$, $m_{max}=50$.
 Each population has the same values for $\alpha$, $m_{min}$, and $m_{max}$ except $\sigma_\epsilon$.
-We selected four values of $\sigma_\epsilon = 0.05,0.1,0.15,0.2$ and each synthetic population contains $N=10000$ sources.
-After getting those populations, we find the probability for each event to be detected by computing the VT of each source.
-Finally, we will randomly pick $100$ sources for each population to perform our analysis.
+We selected four values for $\sigma_\epsilon = 0.05, 0.1, 0.15, 0.2$, and  generated synthetic population containing $N=10000$ sources.
 
-Our weighted populations with $\sigma_\epsilon=0.05, 0.1$ and $\sigma_\epsilon=0.15, 0.2$ are given in Fig. \ref{fig:pop3d0.05_0.1} and Fig. \ref{fig:pop3d0.15_0.2} respectively.
+\subsection{Synthetic population with eccentricity}
+
+After generating four populations, we find the probability for each event to be detected by computing the VT of each source.
+Finally, we did the weighting based on these VTs and randomly picked $N=100$ sources from each population to perform our analysis.
+Our weighted populations with $\sigma_\epsilon=0.05$ and $\sigma_\epsilon=0.2$ are given in Fig.
+\ref{fig:pop3d0.05_0.2}.
+The complete description of each EBBHs population is given in Table \ref{tab:pop_prop}.
+
+\begin{figure*}[]
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d_0.05.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop3d_0.2.png}
+\caption{\label{fig:pop3d0.05_0.2} Synthetic Population of EBBHs.
+The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represents the population for $\sigma_\epsilon=0.2$}
+\end{figure*}
 
@@
     \begin{tabular}{c|cccccc}
-        \hline
-        Populations & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
-        First 1 & 16.81 & 49.81 & 10.8 & 46.82 & 0.01 & 0.05\\ \hline
-        Second 2 & 19.03 & 49.93 & 10.1 & 48.64 & 0.05 & 0.1\\ \hline
-        Third 3 & 10.04 & 49.9 & 10.01 & 49.2 & 0.1 & 0.15\\ \hline
-        Forth 4 & 19.6 & 49.97 & 10.73 & 48.39 & 0.15 & 0.2\\ \hline
+        \hline \hline
+        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{2_{max}} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
+        First  & 12.19 & 49.89 & 10.17 & 46.67 & 0.0004 & 0.0494\\ \hline
+        Second  & 15.42 & 49.79 & 10.46 & 47.59 & 0.0003 & 0.0999\\ \hline
+        Third  & 16.17 & 49.83 & 10.69 & 47.52 & 0.0002 & 0.148\\ \hline
+        Forth  & 12.52 & 49.89 & 10.08 & 49.51 & 0.0012 & 0.1984\\ \hline
     \end{tabular}
@@
 \end{table*}
+To make our study more realistic, we must add the measurement error in each source.
+The chirp mass and symmetric mass ratio are well-constrained compared to the primary and secondary masses of BBHs.
+Therefore, we compute the chirp mass and symmetric mass ratio of each event in the population by using the following relations.
 
-\begin{figure}[H]
+\begin{equation}
+    M_c^T = \frac{(m_1 m_2)^{3/5}}{(m_1+m_2)^{1/5}},
+\end{equation}
 
-\includegraphics[width=0.45\textwidth]{paper/figures/pop3d0.2.png}
-\caption{\label{fig:pop3d0.2} Synthetic Population of EBBHs. The left figure represents the population for $\sigma_\epsilon=0.05$ and right figure represetns the population for $\sigma_\epsilon=0.1$}
+\begin{equation}
+    \eta^T = \frac{(m_1 m_2)}{(m_1+m_2)^2},
+\end{equation}
+where $M_c^T$ and $\eta^T$ are found using the primary and secondary masses of each event generated by the power law model.
+Furthermore, using the following relations, We add the measurement errors in the $M_c^T$ and $\eta^T$.
+\begin{equation}
+    M_c = M_c^T\left( 1+\alpha (r_0+r ) \frac{12}{\rho}\right),
+\end{equation}
 
-\end{figure}
+\begin{equation}
+\eta = \eta^T\left( 1+0.03 (r_0'+r') \frac{12}{\rho}\right),   
+\end{equation}
+where $r_0$ and $r_0'$ are the random numbers drawn from the standard normal distribution, which will shift the mean of the $M_c$ and $\eta$ distribution with respect to $M_c^T$ and $\eta^T$.
+The $r$ and $r'$ are the independent and identically distributed arrays of those randomly generated numbers to spread the distribution.
+The measurement uncertainty is inversely proportional to signal-to-noise ratio $\rho$, drawn from the distribution $p(\rho) \propto \rho^{-4}$, which holds for isotropically distributed sources in a static universe, subject to the threshold $\rho\geq 8$ for detection.
+The $\alpha =0.1$ is used for the scaling inspired by analyses of mock data with the LALINFERENCE pipeline \cite{alpha_error_2015} and includes the impact of correlation with parameters describing arbitrary remnant spins.
 
-\subsection{Synthetic parameter inference with eccentricity}
+Finally, after adding the measurement errors in the $M_c$ and $\eta$, we will convert them back to $m_1$ and $m_2$ to perform our analysis.
+We used the following relation for conversion, and it will provide the masses based on the condition $m_1\geq m_2$.
+
+\begin{equation}
+    m_1 = \frac{1}{2} M_c \eta^{-3/5} (1+\sqrt{\eta_v}),
+\end{equation}
+
+\begin{equation}
+    m_2 = \frac{1}{2} M_c \eta^{-3/5} (1-\sqrt{\eta_v}), 
+\end{equation}
+where $\eta_v = 1-4\eta$, we kept the samples with non-negative values and ignored the negative samples to avoid the square root issues.
+We also added the absolute error in the eccentricity by using the truncated normal distribution scaling at $0.05$ to keep the eccentricity positive.
+
+\subsection{Synthetic population without eccentricity}
 
@@
     \begin{tabular}{c|cccc}
-        \hline
-        Populations & $m^1_{min} [M_\odot] $ & $m^1_{max} [M_\odot]$ & $m^2_{min} [M_\odot]$ & $m^2_{max} [M_\odot]$ \\ \hline
-        First & 16.68 & 49.64 & 10.77 & 46.75 \\ \hline
-        Second & 18.95 & 49.51 & 9.89 & 48.33 \\ \hline
-        Third & 9.13 & 49.55 & 9.1 & 49.11  \\ \hline
-        Forth & 18.2 & 49.79 & 10.3 & 47.08  \\ \hline
+        \hline \hline
+        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{_2{max}} [M_\odot]$ \\ \hline
+        First & 12.09 & 49.6 & 10.17 & 46.27 \\ \hline
+        Second & 15.26 & 49.77 & 10.3 & 47.56 \\ \hline
+        Third & 15.96 & 49.5 & 10.62 & 46.60  \\ \hline
+        Forth & 11.86 & 49.76 & 9.67 & 46.99  \\ \hline
     \end{tabular}
@@
 
-\includegraphics[width=0.45\textwidth]{paper/figures/pop0.05.png}
-\includegraphics[width=0.45\textwidth]{paper/figures/pop0.2.png}
-\caption{\label{fig:pop_0.05_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$.
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d_0.05.png}
+\includegraphics[width=0.45\textwidth]{paper/figures/pop2d_0.2.png}
+\caption{\label{fig:pop2d_0.05_0.2} The left figure shows the primary mass vs secondary mass of the EBBHs and NEBBHs with $\sigma_\epsilon =0.05$.
 The right figure shows the same plot but for $\sigma_\epsilon=0.2$}
@@
 To get the non-eccentric population, we apply the Eq.
-\ref{eq:scaling} on the population given in Fig.
-\ref{fig:pop2d05} and get the new population with scaled masses.
-We get the new minimum and maximum values for the primary and secondary masses in Fig.
-\ref{fig:pop2dscal05}.
+\ref{eq:scaling} on the population given in the table.
+\ref{tab:pop_prop}.
+After applying the scaling equation, we get the new populations with scaled masses, giving us the new minimum and maximum values for the primary and secondary masses given in table \ref{tab:popscl_prop}.
 
-The significant effect after the scaling is not only the mass shift, but one may notice in Fig.
-\ref{fig:pop2dscal05} that $14$ sources were missed out of $100$.
-This reflects that we may miss the various sources by ignoring the effect of eccentricity.
-The mass scaling effect is visible in Fig.
-\ref{fig:pop2d05diff}, and we also notice that sources start missing after $e \gtrsim 0.38$.
+The significant effect after scaling is the mass shift, which can be observed in Fig.
+\ref{fig:pop2d_0.05_0.2}.
+In this figure, the left-hand side shows the mass shift of the first population generated with smaller $\sigma_\epsilon =0.05$, which leads to a lesser mass shift.
+However, on the right hand of Fig.
+\ref{fig:pop2d_0.05_0.2}, one can see the significant mass shift after removing the eccentric component.
+This also reflects that we may miss the various sources by ignoring the effect of eccentricity and we notice that sources start missing, which has eccentricity $\epsilon>0.38$.
 
 \section{Population Inference}
-
+%\renewcommand{\arraystretch}{2}
 \begin{table*}[]
@@
     \begin{tabular}{c|ccccc}
-        \hline
+        \hline \hline
        Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
-      First& $1.88^{+0.18}_{-0.59}$ & $-1.41^{+1.29}_{-0.97}$ & $5.47^{+1.74}_{-1.74}$ & $69.67^{+10.20}_{-10.75}$ & $0.05^{+0.35}_{-0.02}$ \\ \hline
-      Second & $1.89^{+0.15}_{-0.56}$ & $-0.99^{+1.01}_{-0.85}$ & $5.52^{+1.75}_{-1.80}$ & $69.74^{+10.64}_{-10.95}$ & $0.07^{+0.33}_{-0.02}$ \\ \hline
-      Third & $1.90^{+0.14}_{-0.53}$ & $-0.94^{+0.94}_{-0.82}$ & $5.63^{+2.02}_{-1.92}$ & $69.73^{+10.92}_{-11.07}$ & $0.10^{+0.30}_{-0.02}$  \\ \hline
-      Forth & $1.91^{+0.18}_{-0.55}$ & $-1.89^{+1.44}_{-1.00}$ & $5.48^{+1.87}_{-1.90}$ & $69.26^{+11.43}_{-10.96}$ & $0.14^{+0.29}_{-0.03}$  \\ \hline
+      First& $1.42^{+0.58}_{-2.64}$ & $-0.88^{+3.17}_{-1.89}$ & $8.36^{+7.49}_{-3.51}$ & $58.44^{+10.21}_{-24.31}$ & $0.30^{+0.54}_{-0.25}$ \\ \hline
+      Second & $1.36^{+0.63}_{-2.65}$ & $-0.93^{+3.29}_{-1.92}$ & $8.71^{+7.36}_{-3.81}$ & $57.84^{+10.81}_{-25.53}$ & $0.33^{+0.51}_{-0.25}$ \\ \hline
+      Third & $1.39^{+0.58}_{-2.69}$ & $-0.69^{+2.98}_{-1.95}$ & $9.21^{+6.31}_{-3.99}$ & $57.85^{+10.73}_{-27.04}$ & $0.33^{+0.51}_{-0.23}$  \\ \hline
+      Forth & $1.27^{+0.72}_{-2.68}$ & $-0.90^{+3.30}_{-1.98}$ & $8.63^{+7.85}_{-3.72}$ & $57.76^{+11.03}_{-25.82}$ & $0.37^{+0.47}_{-0.23}$  \\ \hline
     \end{tabular}
@@
     \begin{tabular}{c|cccc}
-        \hline
+        \hline \hline
         Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ \\ \hline
-      First & $1.89^{+0.17}_{-0.43}$ & $-1.68^{+1.08}_{-0.80}$ & $5.49^{+2.38}_{-2.41}$ & $68.83^{+14.66}_{-13.13}$  \\ \hline
-      Second & $1.89^{+0.13}_{-0.43}$ & $-0.90^{+0.72}_{-0.68}$ & $5.40^{+2.29}_{-2.40}$ & $68.61^{+14.69}_{-13.33}$  \\ \hline
-      Third & $1.89^{+0.13}_{-0.44}$ & $-0.57^{+0.65}_{-0.67}$ & $5.72^{+2.48}_{-2.40}$ & $68.74^{+14.30}_{-12.99}$   \\ \hline
-      Forth & $1.89^{+0.14}_{-0.45}$ & $-1.20^{+0.81}_{-0.71}$ & $5.32^{+2.33}_{-2.38}$ & $68.83^{+14.51}_{-12.95}$  \\ \hline
+      First & $1.85^{+0.14}_{-1.23}$ & $-1.37^{+1.43}_{-0.94}$ & $7.54^{+3.68}_{-3.04}$ & $63.29^{+8.00}_{-13.08}$  \\ \hline
+      Second & $1.83^{+0.14}_{-1.38}$ & $-1.12^{+1.47}_{-0.91}$ & $7.69^{+3.88}_{-3.15}$ & $61.91^{+8.59}_{-13.87}$  \\ \hline
+      Third & $1.84^{+0.15}_{-1.55}$ & $-0.38^{+1.03}_{-0.96}$ & $8.07^{+3.44}_{-3.27}$ & $61.70^{+8.49}_{-24.26}$   \\ \hline
+      Forth & $1.86^{+0.16}_{-1.85}$ & $-0.31^{+1.17}_{-1.14}$ & $6.99^{+8.77}_{-2.58}$ & $61.15^{+8.72}_{-16.84}$  \\ \hline
     \end{tabular}
@@
 
-\includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.05.png}
+\includegraphics[width=0.95\textwidth]{paper/figures/cor_0.05.png}
 \caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHS for $\sigma_\epsilon=0.05$}}
@@
 
-\includegraphics[width=0.95\textwidth]{paper/figures/corfig_0.2.png}
+\includegraphics[width=0.95\textwidth]{paper/figures/cor_0.2.png}
 \caption{\label{fig:pop3d05}\textbf{Corner Plots for EBBHs and BBHs for $\sigma_\epsilon=0.2$}}

### 2023-07-06 paper/paper.tex

@@
 \begin{abstract}
-Yet to work. 
+The astrophysical mergers produces gravitational waves which carries the plethora of information of their generator.
+We use those gravitational waves signals to make the inference of population properties.
+In this study, we made the inference of non-spinning, eccentric binary black holes.
+As the current GWs signals do not include the eccentricity in their parameters estimation, we made the four synthetic population with different eccentricity using power law model.
+We used the MCMC method to constrain the model and successfully recovered the eccentricity from the population.
 \end{abstract}
@@
 \section{Introduction}
-Big Picture
-\begin{itemize}
-    \item GWs are important
-    \item eccentricity is a signature that can be measured with GWs
-    \item it is a unique signature of strong events
-\end{itemize}
-Why Population, smaller picture
-\begin{itemize}
-    \item population inference of eccentricity is important.
-    \item Some events may have unabmoguis eccentricity (refs to PE, Richard will send them), but many will have only faint signatures or none at all.
-    \item critical for astrophysics to acces distribution of eccentricities and any subpopulation
-\end{itemize}
-Previous work
-\begin{itemize}
-    \item What other people have done with eccentricity, 
-    \item cite any previous work related to eccentricity with population.
-    \item mostly it's real events or high mass events
-    \item what sort of eccentricity we expect to appear
-\end{itemize}
-\begin{itemize}
-    \item Explain, what and how you are doing here.
-    \item assesses how well the eccentricity of the population  can be inferred using low mass BHS where eccentricity measurement should be much more precise.
-    \item for simplicity using a proof of concept population model to emphasize eccentric inference
-\end{itemize}
+Black holes are intriguing and compact objects in the universe \cite{Frolic_BH_Book_2011}.
+They get interesting and enlightening when they coalesce and emit gravitational waves (GWs) \cite{Indrajit_GW_intro_1999}.
+These waves carry insightful information about the BBH system, such as mass, spin, period, eccentricity, distance, location, formation, and evolution.
+In addition, gravitational waves may be the only way to detect those chaotic mergers in space because there was no physical evidence before LIGO's \cite{LIGO_2015} first detection of a BBH merger.
+Afterward, LIGO-VIRGO-KAGRA (LVK) \cite{LIGO_2015,VIRGO_2012,virgo-2015,kagra-2013} are detecting  gravitational waves from various compact binaries such as binary black holes (BBHs), binary neutron stars (BNS), and binary made of neutron stars and black holes.
+The sensitivity of detectors is increasing with time, leading to a higher detection in each observing run, and we will have hundreds of events in future runs \cite{detection_rate_2016,detection_rate_2015}.
 
-We discussed the various formation scenario of EBBHs in the previous section; each one has a distinct signature in the GW signals.
-Hence, it's trivial to observe the effect of each scenario or the mixed effect on the merging population.
-It leads us to make the population model, which considers each scenario and allows distinct merger rates for them.
+Among all the confirmed detection, most of them shows a nearly circular orbit in LVK frequency band.
+However, we have growing evidences that suggests BBHs such as GW15901 may exhibit unambiguous eccentricity or faint signatures \cite{Gamba_2022_GW190521_dynamical,yumeng-2023,Isobel-2022} which can give us better details about the formation scenario.
+In addition to have faint signature in LIGO frequnecy band, we may have the higher eccentric mergers at the lower frequency searches \cite{sesana-2016,chen-2017}.
+There are various scenarios which may develop eccentricity in BBHs orbits, such as stellar scattering, dynamic interactions in dense environments, or a third object in a binary.
+The existence of eccentricity makes things difficult to comprehend and challenges the current understanding of the formation, environment, and evolution of such compact objects.
+We also have the reasonable disagreement on the masses, spin, event rate and formation scenarios \cite{LSC-BBH-2016,LSC-GW150914-2016} which leads to new models \cite{Mandel-2016,marchant-2016}.
+The increment of discovery rate of GWs is constantly challenging the theoretical models and pushing the researcher to develop more sophisticated models.
 
-Most of the eccentric binary black holes (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
+Therefore, identifying and understanding the populations of EBBHs with growing number of detection of compact binaries will give us more insightful information about formation and evolution of massive stars from birth to death over cosmic time.
+
+Eccentricity is the sharply decreasing function of GWs frequency. \cite{peters-1964}.
+
+The GWs produced by the dynamical formed BBHs can carry the non-negligible eccentricity within the LIGO frequency band \cite{samsing-2018, Rodriguez-2018,Antonini-2014}.
+
+Currently, LVK is using the ciruclar wave form.
+However, studies shows that we need to consider the multiple formation channels to get the better understaing because, each channel does not contribute more than 70\% of the observation sample of binary black holes \cite{zevin-2021}.
+
+Till now, no modelled search for eccentric signals due to lack of inspiral-merger-ringdown waveform models with eccentricity.
+
+Unmodelled searches mostly efficient for higher mass system whereas effect of eccentricity will be more in the low mass systems.
+
+Despite this, little is known about the formation channels of these mergers.
+It is difficult to characterize the formation channel of a given merger from the parameters that are usually measured, mass and spin.
+Unfortunately, many different formation scenarios can produce these combinations.
+One possible way of determining how a binary may have formed is to better constrain the dynamics of the merger, in addition to these parameters.
+
+Specifically, orbital eccentricity may help distinguish different populations of compact binaries.
+Slow-forming, low-energy binaries tend to start with low eccentricities, and gravitational wave radiation further circularizes their orbits \cite{Peters-1964}.
+However, binaries formed in more violent, energetic environments, such as globular clusters, can have large eccentricities.
+These binaries are expected to retain measurable eccentricities after entering the LIGO sensitivity band, potentially making orbital eccentricity an important parameter in determining formation channel \cite{Rod-2018,zevin-samsing-2019}.
+
+Recent works investigates how measurements of black hole mass and spin distributions can elucidate the population properties of binary black holes \cite{Dan_2019,Michael-2015,Mandel_2017_Errors,samsing-hamers-2019,belczynski-2016,Colm-2017,Akinobu-2017,Michael-zevin-2017,farr-2017-nature,Richard-2017-natal-kicks,Dan-Richard-2018}.
+we also have the research to constrain the population models with eccentricity \cite{lower-2018-ecc-pop}.
+It has also been suggested that the future
+Laser Interferometer Space Antenna (LISA) will be able to
+observe nearby stellar-mass BBH during the early inspiral
+phase. These observations would allow for long-term
+tracking of BBH orbital properties which can be used to
+infer the formation mechanism \cite{Breivik-2016}.
+
+In Sec.
+\ref{sec:methods} we described the Bayesian statistical methods which we used to make the population inference, and brief method of volume-time estimation.
+We also explained the previously constructed\cite{fishbach-2017,2018talbot_bbh_model} power law model with modification for eccentricity by using truncated normal distribution.
+we used this model to generate synthetic population and it's inference.
+The Sec.
+\ref{sec:syn_pop} describes that how we have created synthetic population using power law model and then adding error into each event to make population closer to real population detectable by LVK.
+Secodnly, we explained about the scaling by removing the eccentricity from the synthetic events to make the comparison between EBBHs and NEBBHs population.
+
+%In addition to electromagnetic waves spectrum, the gravitational waves also provides the completely new domain to study the formation and environment of such compact object.
+
+%Eccentricity is the unique signature of strong events and can be measured with GWs.
+
+%Most of the eccentric binary black holes (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
 \cite{mass_ecc_limit_2018}.
@@
 The coalescing BBH can be completely described by three intrinsic and seven extrinsic parameters.
-The intrinsic parameters such as mass of binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary.
+The intrinsic parameters, such as the mass of the binary component ($m_i$), spin $( \chi_i)$, and eccentricity $\epsilon$  are subject to the orbital evolution of the binary.
 The extrinsic parameters are orientation (orbital phase, polarization, and inclination), sky location (right ascension and declination), luminosity distance, and coalescence time depend on the observer.
@@
 \end{equation}
-
+%
 where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or also known as evidence.
@@
 \end{equation}
-
+%
 where $\mu(\mathcal{R},\Lambda)$ is the expected number of detection under the given population parametrization $\Lambda$ with the overall rate $\mathcal{R}$.
@@
 We may iterate this as we need and store multiple sample values untill they converge.
-This method is in a Python package called EMCEE \cite{emcee_paper}.
+The specific implementation we use is a Python package called EMCEE \cite{emcee_paper}.
 
@@
 
-To make our study realistic, including the sensitivity of the LVK instruments is significant.
-This sensitivity is defined by VT estimation.
-$V$ is the characteristic volume with units $Gpc^{-3}yr^{-1}$, which refers to the possible detection region in the sky for the LVK \cite{Volume_1993}, and $T$ is the time duration of making observations at this sensitivity.
-Existing LVK instruments' sensitivity depends only on the mass; therefore, we will get the same VT calculated previously \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries.
+To make our study realistic, we include the sensitivity of the LVK instruments.
+This sensitivity is defined by time-volume to which a census of gravitational wave events is sensitive: inferring the product $VT$.
+In this expression, $V$ is the characteristic volume with units $Gpc^{3}$, which refers to the possible detection region in the sky for the LVK \cite{Volume_1993}, and $T$ is the time duration of making observations at this sensitivity.
+In practice, $VT$ reflects a suitable time-averaged or cumulative sensitivity, as the true network and sensitivity varies over time.
+Existing LVK instruments' sensitivity depends primarily on the mass and to a lesser extent on binary spin and (if present) modest eccentricity.
+Since we neglect spin in this work, we assume the network will have the same  VT versus mass as was previously estimated  \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries.
 Hence, we briefly explain the calculations, see \cite{Dan_2019} for details.
@@
 \end{align}
-
+%
 where $\alpha$ is the power law index, $\mathcal{R}$ is the merger rate, $m_{min}, m_{max}$ are the minimum and maximum masses of the binary components in the population, and $\sigma_e$ is the orbital eccentricity distribution.
@@
 For our analysis, we defined a constant of integration equal to $\int_V dm_1 dm_2 d\epsilon p(m_1,m_2,\epsilon) = 1$.
-Our detectors are sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$, therefore, we will use $k_m=0$ throughout the studies.
+Our detectors are sensitive to high-mass BBHs, particularly $M_{max}\geq 200 M_\odot$.
+Therefore, we will use $k_m=0$ throughout the studies.
 As a result, we have our reduced form of the truncated power law in Eq.
@@
 \section{Synthetic Population}
-
+\label{sec:syn_pop}
 We have created four synthetic populations by choosing the power law parameters $\alpha = -1$, $m_{min} = 10$, $m_{max}=50$.
 Each population has the same values for $\alpha$, $m_{min}$, and $m_{max}$ except $\sigma_\epsilon$.
-We selected four values for $\sigma_\epsilon = 0.05, 0.1, 0.15, 0.2$, and  generated synthetic population containing $N=10000$ sources.
+We selected four values for $\sigma_\epsilon = 0.05, 0.1, 0.15, 0.2$, and  generated synthetic population containing $10000$ sources.
 
@@
 Finally, we did the weighting based on these VTs and randomly picked $N=100$ sources from each population to perform our analysis.
-Our weighted populations with $\sigma_\epsilon=0.05$ and $\sigma_\epsilon=0.2$ are given in Fig.
+Our weighted populations with $\sigma_\epsilon=0.05$ and $\sigma_\epsilon=0.2$ are shown in Fig.
 \ref{fig:pop3d0.05_0.2}.
-The complete description of each EBBHs population is given in Table \ref{tab:pop_prop}.
+The complete description of each EBBHs population is provided in Table \ref{tab:pop_prop}.
 
@@
 \end{table*}
+
 To make our study more realistic, we must add the measurement error in each source.
+Rather than generate synthetic gravitational wave sources and perform full Bayesian inference, following previous work \cite{Mandel_2017_Errors} we generate mock measurement errors motivated by real parameter inference investigations.
 The chirp mass and symmetric mass ratio are well-constrained compared to the primary and secondary masses of BBHs.
-Therefore, we compute the chirp mass and symmetric mass ratio of each event in the population by using the following relations.
+Therefore, we compute them for each event in a population by using the following relations.
 
@@
 \end{equation}
+%
 where $M_c^T$ and $\eta^T$ are found using the primary and secondary masses of each event generated by the power law model.
 Furthermore, using the following relations, We add the measurement errors in the $M_c^T$ and $\eta^T$.
+
 \begin{equation}
@@
 where $\eta_v = 1-4\eta$, we kept the samples with non-negative values and ignored the negative samples to avoid the square root issues.
+
 We also added the absolute error in the eccentricity by using the truncated normal distribution scaling at $0.05$ to keep the eccentricity positive.
+\textbf{this is important - refer to other papers.  THIS IS REASONABLE FOR LOW MASS BUT VERY OPTIMISTIC FOR HIGH MASS}
+STRETCH GOAL: what if you ran again, but with large error?
 
@@
 
-To compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we created four more populations by scaling the existing populations using the Eq.
-\ref{eq:scaling} defined in \cite{2021_scaling_paper}.
-The Eq.
-\ref{eq:scaling} removes the eccentric component from the population, scales the masses, and eventually, we get non-eccentric populations to compare with the existing eccentric population.
+To compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we estimate how our sources would be characterized by parameter inferences which omitted the effects of eccentricity.
+Following \cite{2021_scaling_paper}, the following effective chirp mass is well-constrained by observations dominated by the inspiral of a slightly eccentric binary:
 \begin{align}
@@
 \end{align}
+%
+Our ansatz for source identification and characterization is that the best-fitting parameters and posteriors are directly related to the true posteriors, except that the recovered chirp mass is given by  Eq.
+\ref{eq:scaling}.
+This process removes the eccentric component from the population only by scaling the masses and omitting the eccentricity.
 
-\begin{table*}[]
-    \centering
-    \begin{tabular}{c|cccc}
-        \hline \hline
-        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{_2{max}} [M_\odot]$ \\ \hline
-        First & 12.09 & 49.6 & 10.17 & 46.27 \\ \hline
-        Second & 15.26 & 49.77 & 10.3 & 47.56 \\ \hline
-        Third & 15.96 & 49.5 & 10.62 & 46.60  \\ \hline
-        Forth & 11.86 & 49.76 & 9.67 & 46.99  \\ \hline
-    \end{tabular}
-    \caption{Population Properties of NEBBHs}
+%\begin{table*}[]
+%    \centering
+%    \begin{tabular}{c|cccc}
+%        \hline \hline
+%        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{_2{max}} [M_\odot]$ \\ \hline
+%        First & 12.09 & 49.6 & 10.17 & 46.27 \\ \hline
+%        Second & 15.26 & 49.77 & 10.3 & 47.56 \\ \hline
+%        Third & 15.96 & 49.5 & 10.62 & 46.60  \\ \hline
+%        Forth & 11.86 & 49.76 & 9.67 & 46.99  \\ \hline
+%    \end{tabular}
+%    \caption{Population Properties of NEBBHs}
     \label{tab:popscl_prop}
-\end{table*}
+%\end{table*}
 
@@
 
-To get the non-eccentric population, we apply the Eq.
+%To get the non-eccentric population, we apply the Eq.
 \ref{eq:scaling} on the population given in the table.
@@
 \section{Population Inference}
+\label{sec:pop_inference}
 %\renewcommand{\arraystretch}{2}
@@
 \end{table*}
+
+Figure A shows the results of population inference under the most conservative case with $\sigma_\epsilon=0.05$.  In this figure, the ZZ color shows the results of QQ.
+%
+Most notably, this figure shows that even in the limit of extremely small ecctricity, the impact of eccentricity can be measured from a large population of observations, comparable to the current size.
+This result depends of course on our assumption that all binaries are equally likely to have eccentricity, and that our measurement uncertainty in eccentricity is both indpeendent of mass and fairly small, both optimisitic assumptions.
+In fact, the overall uncertainty in $\sigma_\epsilon$ is expected: given 100 events, each with a measurement error of $0.05$, we expect $\sigma_\epsilon$ to be measured to of order $0.05/\sqrt{N}$ (handwavy -- really a chisquared distribution,etc).
+SUSPECT THAT IF YOU MAKE THIS JUST 3X bigger, STILL GET RESULT -- you will almost always get this result
+
+Also noteworthy is the impact of neglecting eccentricity.
+If eccentricity is neglected, then the population inference is biased away from true values: the black contours in the first column do not include the true values for $m_{max}, m_{min}$ example.
+
+Figure B shows the results of population inference under the most optimistic case with $\sigma_\epsilon=0.2$.  In this figure, the ZZ color shows the results of QQ.
+
 we can discuss the corner plots here.
+
+TAKEAWAYS
+
+* Eccentricity can be measured
+
+* If not measured, biases
 
@@
 
+* summarize your calculations
+
+* summarize key takeaways
+
+** we WILL find this, measuremeent error can't be too large/infinite!
+
+* contrast /compare with other papers
+
+* some comments on what it all means for the big picture (dynamical evolution tests for example could be done)
+
 \begin{acknowledgements}

### 2023-07-09 new_powerlaw_mcmc.py
Making Chain Plots to check the convergence

@@ -579,6 +579,20 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
+         import matplotlib.pyplot as plt
+         fig, axes = plt.subplots(5, figsize=(10, 7), sharex=True)
+         samples = posterior_pos
+         labels = ["log_rate", "alpha", "m_min","m_max","sigma_ecc"]
+         for i in range(ndim):
+            ax = axes[i]
+            ax.plot(samples[:, :, i].T,'-',color='k', alpha=0.3)
+            ax.set_xlim(0, len(samples))
+            ax.set_ylabel(labels[i])
+            plt.savefig("chain.png")
+         axes[-1].set_xlabel("step number");
+
+        # plt.plot(posterior_pos[:,:,0].T, '-', color='k')
+        # plt.savefig('chain1.png')
       # irint(nsamples,n_walkers,ndim)
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate

### 2023-07-09 original_pop_code/original_powerlaw_mcmc.py

@@ -544,7 +544,16 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
-
+         fig, axes = plt.subplots(5, figsize=(10, 7), sharex=True)
+         samples = posterior_pos
+         labels = ["log_rate", "alpha", "m_min","m_max","sigma_ecc"]
+         for i in range(ndim):
+            ax = axes[i]
+            ax.plot(samples[:, :, i].T,'-',color='k', alpha=0.3)
+            ax.set_xlim(0, len(samples))
+            ax.set_ylabel(labels[i])
+            plt.savefig("chain.png")
+         axes[-1].set_xlabel("step number");
 
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate

### 2023-07-09 run.sh

@@ -1,1 +1,1 @@
-time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout_10_120_20000.hdf5 --n-walkers 120 --n-samples 20000
+time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 120 --n-samples 20

### 2023-07-09 new_powerlaw_mcmc.py
Chain Plots: Adding True values

@@ -108,7 +108,7 @@ def log_prior_pop(
         alpha_min=None, alpha_max=None,
         m_min_min=None, m_min_max=None,
         m_max_min=None, m_max_max=None,
-        sigma_ecc_min=0, sigma_ecc_max=0.4,
+        sigma_ecc_min=0, sigma_ecc_max=0.5,
         sigma_ecc_scale = "uniform",
         log_rate_scale="uniform", alpha_scale="uniform",
         m_min_scale="uniform", m_max_scale="uniform",
@@ -218,7 +218,7 @@ def init_uniform(
 
         columns.append(m_max)
 
-    sigma_ecc = rand_state.uniform(0,1, nwalkers)
+    sigma_ecc = rand_state.uniform(0,0.5, nwalkers)
     columns.append(sigma_ecc)
 
     return numpy.column_stack(tuple(columns))
@@ -366,7 +366,7 @@ def _get_args(raw_args):
     )
     parser.add_argument(
         "--m-max-min",
-        type=float, default=20.0,
+        type=float, default=30.0,
         help="Minimum m_max allowed.",
     )
     parser.add_argument(
@@ -493,7 +493,7 @@ def _main(raw_args=None):
         cli_args.m_min_min, cli_args.m_min_max,
         cli_args.m_max_min, cli_args.m_max_max,
         log_rate_scale=cli_args.log_rate_initial_cond,
-        sigma_ecc_min=0, sigma_ecc_max=1,
+        sigma_ecc_min=0, sigma_ecc_max=0.5,
         alpha_scale=cli_args.alpha_initial_cond,
         m_min_scale=cli_args.m_min_initial_cond,
         m_max_scale=cli_args.m_max_initial_cond,
@@ -579,20 +579,30 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
+         #Making chain plots to check the convergence of the chains
          import matplotlib.pyplot as plt
-         fig, axes = plt.subplots(5, figsize=(10, 7), sharex=True)
+         inj_rate,inj_alpha,inj_m_min,inj_m_max,inj_sigma_ecc = (2,-1,10,50,0.1)
+         fig, axes = plt.subplots(5, figsize=(12, 12), sharex=True)
          samples = posterior_pos
-         labels = ["log_rate", "alpha", "m_min","m_max","sigma_ecc"]
[…]

### 2023-07-09 original_pop_code/original_powerlaw_mcmc.py

@@ -317,7 +317,7 @@ def _get_args(raw_args):
     )
     parser.add_argument(
         "--m-min-max",
-        type=float, default=10.0,
+        type=float, default=20.0,
         help="Maximum m_min allowed.",
     )
 
@@ -341,7 +341,7 @@ def _get_args(raw_args):
     )
     parser.add_argument(
         "--m-max-max",
-        type=float, default=100.0,
+        type=float, default=80.0,
         help="Maximum m_max allowed.",
     )
 
@@ -544,17 +544,28 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
-         fig, axes = plt.subplots(5, figsize=(10, 7), sharex=True)
+         #Making chain plots to check the convergence of the chains
+         import matplotlib.pyplot as plt
+         inj_rate,inj_alpha,inj_m_min,inj_m_max,inj_sigma_ecc = (2,-1,10,50,0.1)
+         fig, axes = plt.subplots(5, figsize=(12, 12), sharex=True)
          samples = posterior_pos
-         labels = ["log_rate", "alpha", "m_min","m_max","sigma_ecc"]
+         labels = ["$log_{10}(\mathcal{R})$",r"$\alpha$", "$m_{min}$","$m_{max}$","$\sigma_\epsilon$"]
          for i in range(ndim):
             ax = axes[i]
             ax.plot(samples[:, :, i].T,'-',color='k', alpha=0.3)
             ax.set_xlim(0, len(samples))
             ax.set_ylabel(labels[i])
+            axes[0].axhline(inj_rate,color='blue')
+            axes[1].axhline(inj_alpha,color='blue')
+            axes[2].axhline(inj_m_min,color='blue')
+            axes[3].axhline(inj_m_max,color='blue')
+            axes[4].axhline(inj_sigma_ecc,color='blue')
             plt.savefig("chain.png")
          axes[-1].set_xlabel("step number");
-
+         #plt.plot(posterior_pos[:,:,0].T, '-', color='k')
+         #plt.axhline(2,color='blue')
+         #plt.savefig('chain1.png')
+         
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate
 # from the log10(rate), and the normalization factor for the mass

### 2023-07-10 paper/paper.tex
Intro and Abstract first draft complete

@@
 \title{Population Inference of Non-Spinning Eccentric Binary Black Holes}
+%\author{M. Zeeshan}
+\affiliation{\RIT}
 %\author{R. O'Shaughnessy}
-\affiliation{\RIT}
-%\author{M. Zeeshan}
 
@@
 \begin{abstract}
-The astrophysical mergers produces gravitational waves which carries the plethora of information of their generator.
-We use those gravitational waves signals to make the inference of population properties.
-In this study, we made the inference of non-spinning, eccentric binary black holes.
-As the current GWs signals do not include the eccentricity in their parameters estimation, we made the four synthetic population with different eccentricity using power law model.
-We used the MCMC method to constrain the model and successfully recovered the eccentricity from the population.
+The astrophysical mergers produce gravitational waves (GWs), which carry insightful information of their sources such as Binary Black holes (BBHs) mergers.
+The LIGO-VIRGO-KAGRA (LVK) detects GWs signals, and we use those measurements to infer BBHs population properties such as mass, spin, and eccentricity distribution in the universe.
+Although mass and spin distribution gives information of the source formation, evolution, and environment but eccentricity is the unique and measurable signature of strong events.
+Hence, the current parameter estimation (PE) methods of GWs signals do not include the eccentricity due to the lack of waveform with eccentricity.
+Therefore, for our analysis, we generated four synthetic populations of non-spinning, non-precessing, lower mass $(10 M_\odot-50 M_\odot)$, eccentric binary black holes (EBBHs) using a modified power law model, each consisting of $100$ events with different eccentricity distribution $\sigma_\epsilon = (0.05,0.1,0.15,0.2)$.
+Furthermore, to compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we estimate how our sources would be characterized by parameter inferences which omitted the effects of eccentricity.
+We apply the Markov Chain Monte Carlo (MCMC) method to constrain the model parameters: event rate $\mathcal{R}$, $\alpha$, $m_{min}$, $m_{max}$, and $\sigma_\epsilon$.
+We are also able to recover the injected eccentricity for each population \textbf{(need to add more specfic information after runs)}.
+All the tools we used in this analysis are publicly available in a GitHub repository, and one can use them to infer the synthetic population of EBBHs.
 \end{abstract}
@@
 Black holes are intriguing and compact objects in the universe \cite{Frolic_BH_Book_2011}.
-They get interesting and enlightening when they coalesce and emit gravitational waves (GWs) \cite{Indrajit_GW_intro_1999}.
+They get exciting and enlightening when they coalesce and emit gravitational waves (GWs) \cite{Frolic_BH_Book_2011, Indrajit_GW_intro_1999}.
 These waves carry insightful information about the BBH system, such as mass, spin, period, eccentricity, distance, location, formation, and evolution.
-In addition, gravitational waves may be the only way to detect those chaotic mergers in space because there was no physical evidence before LIGO's \cite{LIGO_2015} first detection of a BBH merger.
-Afterward, LIGO-VIRGO-KAGRA (LVK) \cite{LIGO_2015,VIRGO_2012,virgo-2015,kagra-2013} are detecting  gravitational waves from various compact binaries such as binary black holes (BBHs), binary neutron stars (BNS), and binary made of neutron stars and black holes.
-The sensitivity of detectors is increasing with time, leading to a higher detection in each observing run, and we will have hundreds of events in future runs \cite{detection_rate_2016,detection_rate_2015}.
+Primarily, gravitational waves may be the only way to detect those chaotic mergers in space because there was no physical evidence before LIGO's \cite{LIGO_2015} first detection of a BBH merger named GW150914.
+Afterward, LIGO-VIRGO-KAGRA (LVK) \cite{LIGO_2015, VIRGO_2012, Virgo-2015,kagra-2013} are detecting gravitational waves from various compact binaries such as binary black holes (BBHs), binary neutron stars (BNS), and binary made of a neutron star and black hole.
+The researchers use those GWs signal to infer the intrinsic and extrinsic properties of associated binary such as mass, spin, and eccentricity.
 
-Among all the confirmed detection, most of them shows a nearly circular orbit in LVK frequency band.
-However, we have growing evidences that suggests BBHs such as GW15901 may exhibit unambiguous eccentricity or faint signatures \cite{Gamba_2022_GW190521_dynamical,yumeng-2023,Isobel-2022} which can give us better details about the formation scenario.
-In addition to have faint signature in LIGO frequnecy band, we may have the higher eccentric mergers at the lower frequency searches \cite{sesana-2016,chen-2017}.
-There are various scenarios which may develop eccentricity in BBHs orbits, such as stellar scattering, dynamic interactions in dense environments, or a third object in a binary.
-The existence of eccentricity makes things difficult to comprehend and challenges the current understanding of the formation, environment, and evolution of such compact objects.
-We also have the reasonable disagreement on the masses, spin, event rate and formation scenarios \cite{LSC-BBH-2016,LSC-GW150914-2016} which leads to new models \cite{Mandel-2016,marchant-2016}.
-The increment of discovery rate of GWs is constantly challenging the theoretical models and pushing the researcher to develop more sophisticated models.
+Mass and spin are widely used to understand a binary's evolution and formation channel, ignoring the significant parameter eccentricity.
+Because the existence of eccentricity makes things difficult to comprehend and challenges the current understanding of the formation, environment, and evolution of such compact objects, various scenarios may develop eccentricity in BBHs orbits, such as stellar scattering, dynamic interactions in dense environments, or interaction of the third object with a binary.
+Therefore, a more insightful method to understand different populations and determine how a binary may have formed is to constrain the eccentricity along with mass and spin \cite{Rod-2018, zevin-samsing-2019, samsing-2018, Rodriguez-2018, Antonini-2014}.
 
-Therefore, identifying and understanding the populations of EBBHs with growing number of detection of compact binaries will give us more insightful information about formation and evolution of massive stars from birth to death over cosmic time.
+Low energy and slow-forming binaries usually start with lesser eccentricities, and eventually, gravitational wave radiation circularizes their orbits \cite{Peters-1964} before entering the LVK frequency band.
+However, binaries formed in more violent, energetic environments, such as globular clusters, can have large eccentricities.
+These binaries retain measurable eccentricities after entering the LVK sensitivity band, potentially making orbital eccentricity an important parameter in determining formation channels \cite{Rod-2018,zevin-samsing-2019,samsing-2018, Rodriguez-2018,Antonini-2014}.
+We may also have stellar mass higher eccentric mergers at the lower frequency searches \cite{sesana-2016,chen-2017}, which can be measurable by detectors like Laser Interferometer Space Antenna (LISA) \cite{LISA-2017}.
+These observations would allow for long-term tracking of BBHs orbital properties, which can be used to infer the formation mechanism \cite{Breivik-2016}.
 
-Eccentricity is the sharply decreasing function of GWs frequency. \cite{peters-1964}.
+Although eccentricity is a signature that can be measured with GWs and is unique to extreme events, all the confirmed detection till now shows a nearly circular orbit in the LVK frequency band.
+The first potential reason for getting circular orbits is that eccentric effects are more evident in low-mass events, and current searches are more efficient for higher-mass mergers.
+Secondly, it may result from selection biases in the waveforms because LVK detectors only use circular waveforms for parameter estimation (PE), which better present binaries evolved in an isolated environment.
+Studies show that we must consider the multiple formation channels to understand the population better.
+Because a single channel does not contribute more than $70\%$ of the observation sample of BBHs \cite{zevin-2021}.
+We also have growing shreds of evidence that suggests BBHs like GW190521 may exhibit unambiguous eccentricity or faint signatures \cite{Gamba_2022_GW190521_dynamical,yumeng-2023,Isobel-2022}, which leads towards the dynamical formation of the source.
+So, it is critical for astrophysical implications to assess eccentricity distribution.
 
-The GWs produced by the dynamical formed BBHs can carry the non-negligible eccentricity within the LIGO frequency band \cite{samsing-2018, Rodriguez-2018,Antonini-2014}.
+The sensitivity of LVK detectors is increasing with time, leading to more detection in each observing run.
+We will have hundreds of events in future runs, and with the upcoming O4 run, we expect detection every other day \cite{detection_rate_2016,detection_rate_2015}.
+Interestingly, the new detections lead us to the reasonable disagreement on the masses, spin, event rate, and formation scenarios \cite{LSC-BBH-2016, LSC-GW150914-2016}, which pushes researchers to develop new models \cite{Mandel-2016,marchant-2016}, and also provide more information about the observable GWs \cite{Barausse-2018, Abbot-2016-schotastic,dvorkin-2016}.
+Additionally, researchers are actively producing the eccentric waveform models analytically \cite{Huerta-2014} and numerically \cite{gold-2016,hinder-2010, Healy-2022, Alessandro-2022, Campanelli-2009}.
+Those waveforms in the bank will give us real PE with eccentricity shortly.
+Therefore, identifying and understanding the populations of EBBHs with a growing number of detection will give us more insightful information about the formation and evolution of massive stars from birth to death over cosmic time.
 
-Currently, LVK is using the ciruclar wave form.
-However, studies shows that we need to consider the multiple formation channels to get the better understaing because, each channel does not contribute more than 70\% of the observation sample of binary black holes \cite{zevin-2021}.
+Recent works investigate how measurements of BBHs mass and spin distributions can elucidate the population properties of BBHs \cite{Dan_2019, Michael-2015, Mandel_2017_Errors, samsing-hamers-2019,  belczynski-2016, Colm-2017, Akinobu-2017, Michael-zevin-2017, farr-2017-nature, Richard-2017-natal-kicks, Dan-Richard-2018,Abbot-2019-pop}.
+Researchers also study the population of BBHs with eccentricity \cite{lower-2018-ecc-pop,Fang-2019}.
+However, most research is inclined towards the high mass range or real events \cite{wu-2020}.
+Therefore, this paper focused on the non-spinning, non-precessing, lower mass $(10 M_\odot - 50 M_\odot)$, and lesser eccentric $(0-0.2)$ merger.
+In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2=100M_\odot$.
+We present the correlation of event rate, mass, and eccentricity distribution because studies show that binary parameters and event rate are strongly coupled.
+We also infer how well we can recover the mass, eccentricity, and event rate using the $100$ eccentric events.
+To extend our analysis, we compared constrained parameters using the EBBBhs and NEBBHs, which show a considerable difference.
 
-Till now, no modelled search for eccentric signals due to lack of inspiral-merger-ringdown waveform models with eccentricity.
+We organized this paper as follows.
+In Sec.
+\ref{sec:methods}, we described the Bayesian statistical methods used to make the population inference.
+Briefly, we also described the volume-time estimation to accommodate the LVK sensitivity.
+Using a truncated normal distribution, we modified the previously constructed\cite{fishbach-2017,2018talbot_bbh_model} power law model for eccentricity.
+We used this model to generate a synthetic population and then made the inference using it.
+The Sec.
+\ref{sec:syn_pop} describes how we have created a synthetic population using the power law model and then added an error in each event to make the population closer to real events detectable by LVK.
+Secondly, we explained the scaling to remove the eccentricity from the synthetic events to compare the different constraints using EBBHs and NEBBHs populations.
+Sec.
+\ref{sec:pop_inference} discusses our results and explains how well we have constrained the parameters.
+Also, their accuracy and recovery comparison among EBBHs and NEBBHs.
+Finally, we summarize our findings in Sec.
+\ref{sec:conclude}.
 
-Unmodelled searches mostly efficient for higher mass system whereas effect of eccentricity will be more in the low mass systems.
-
-Despite this, little is known about the formation channels of these mergers.
-It is difficult to characterize the formation channel of a given merger from the parameters that are usually measured, mass and spin.
-Unfortunately, many different formation scenarios can produce these combinations.
-One possible way of determining how a binary may have formed is to better constrain the dynamics of the merger, in addition to these parameters.
-
-Specifically, orbital eccentricity may help distinguish different populations of compact binaries.
-Slow-forming, low-energy binaries tend to start with low eccentricities, and gravitational wave radiation further circularizes their orbits \cite{Peters-1964}.
-However, binaries formed in more violent, energetic environments, such as globular clusters, can have large eccentricities.
-These binaries are expected to retain measurable eccentricities after entering the LIGO sensitivity band, potentially making orbital eccentricity an important parameter in determining formation channel \cite{Rod-2018,zevin-samsing-2019}.
-
-Recent works investigates how measurements of black hole mass and spin distributions can elucidate the population properties of binary black holes \cite{Dan_2019,Michael-2015,Mandel_2017_Errors,samsing-hamers-2019,belczynski-2016,Colm-2017,Akinobu-2017,Michael-zevin-2017,farr-2017-nature,Richard-2017-natal-kicks,Dan-Richard-2018}.
-we also have the research to constrain the population models with eccentricity \cite{lower-2018-ecc-pop}.
-It has also been suggested that the future
-Laser Interferometer Space Antenna (LISA) will be able to
-observe nearby stellar-mass BBH during the early inspiral
-phase. These observations would allow for long-term
-tracking of BBH orbital properties which can be used to
-infer the formation mechanism \cite{Breivik-2016}.
-
-In Sec.
-\ref{sec:methods} we described the Bayesian statistical methods which we used to make the population inference, and brief method of volume-time estimation.
-We also explained the previously constructed\cite{fishbach-2017,2018talbot_bbh_model} power law model with modification for eccentricity by using truncated normal distribution.
-we used this model to generate synthetic population and it's inference.
-The Sec.
-\ref{sec:syn_pop} describes that how we have created synthetic population using power law model and then adding error into each event to make population closer to real population detectable by LVK.
-Secodnly, we explained about the scaling by removing the eccentricity from the synthetic events to make the comparison between EBBHs and NEBBHs population.
-
-%In addition to electromagnetic waves spectrum, the gravitational waves also provides the completely new domain to study the formation and environment of such compact object.
-
-%Eccentricity is the unique signature of strong events and can be measured with GWs.
-
-%Most of the eccentric binary black holes (EBBHs) are driven by low masses $(10M_\odot-30M_\odot)$ and if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.
-\cite{mass_ecc_limit_2018}.
-Therefore,  in this study, we  will consider binaries with non-spinning $(\chi_i = 0)$, eccentric $(0-0.5)$ and moderate mass range $(10M_\odot-50M_\odot)$.
-In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2=100M_\odot$.
+%if any component of the binary has mass less than $30 M_\odot$ then only $10\%$ of them are able to maintain eccentricity near the last stable orbit.  \cite{mass_ecc_limit_2018}.
 
@@
 
-We use HBM to constrain the population models with gravitational wave signals. It is also a good choice to consider the sensitivity of LIGO-VIRGO-KAGRA (LVK) and uncertainty in each detection.
-
+We use HBM to constrain the population models with gravitational wave signals.
 In HBM, we have the N number of discrete detections.
@@
 We also refer to it as the likelihood function $\mathcal{L}(\lambda)=p(d|\lambda)$ of one event.
-A given value of $\lambda$ is obtained from a mathematical model commonly known as a waveform to calculate a likelihood.
-Once you have a likelihood function, you may use a uniform prior or any other prior to find a posterior probability using the Bayes theorem as given in Eq.
+To calculate likelihood, given value of $\lambda$ is obtained from a mathematical model commonly known as a waveform .
+Once we have a likelihood function, you may use a uniform prior or any informative prior to find a posterior probability using the Bayes theorem as given in Eq.
 \ref{eq:Bayes_ind}
@@
 Now having the PE of individual sources, we will follow the Bayesian framework for population inference.
-The likelihood of a population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ is equivalent to the probability of the individual sources given the population parameter $\Lambda$ is written as follows.
+The likelihood of a population parameter $\Lambda$, also considered as uncertainty in $\Lambda$ is equivalent to the probability of the individual sources given the population parameter $\Lambda$ is written as follows
 
@@
 \label{eq:likelihood_pop}    
-\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda)
+\mathcal{L}(\Lambda)\equiv p(d_1,d_2,d_3,...,d_N|\Lambda).
 \end{equation}
 
-We will use the likelihood given in Eq. \ref{eq:likelihood_pop} in the Bayes theorem given in Eq. \ref{eq:Bayes} to find the posterior probability.
+We will use likelihood provided in Eq. \ref{eq:likelihood_pop} in the Bayes theorem defined in Eq. \ref{eq:Bayes} to find posterior probability.
 
@@
 Existing LVK instruments' sensitivity depends primarily on the mass and to a lesser extent on binary spin and (if present) modest eccentricity.
-Since we neglect spin in this work, we assume the network will have the same  VT versus mass as was previously estimated  \cite{Dan_2019} for non-spinning, non-eccentric, and no-precessing binaries.
+Since we neglect spin in this work, we assume the network will have the same  VT versus mass as was previously estimated  \cite{Dan_2019} for non-spinning, non-eccentric, and non-precessing binaries.
 Hence, we briefly explain the calculations, see \cite{Dan_2019} for details.

### 2023-07-11 new_powerlaw_mcmc.py
Numpy and chain plot update

@@ -579,30 +579,20 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
-         #Making chain plots to check the convergence of the chains
          import matplotlib.pyplot as plt
-         inj_rate,inj_alpha,inj_m_min,inj_m_max,inj_sigma_ecc = (2,-1,10,50,0.1)
          fig, axes = plt.subplots(5, figsize=(12, 12), sharex=True)
          samples = posterior_pos
          labels = ["$log_{10}(\mathcal{R})$",r"$\alpha$", "$m_{min}$","$m_{max}$","$\sigma_\epsilon$"]
          for i in range(ndim):
             ax = axes[i]
-            ax.plot(samples[:, :, i].T,'-',color='k', alpha=0.3)
+            ax.plot(samples[:, :, i],'-', alpha=0.3)
             ax.set_xlim(0, len(samples))
             ax.set_ylabel(labels[i])
-            axes[0].axhline(inj_rate,color='blue')
-            axes[1].axhline(inj_alpha,color='blue')
-            axes[2].axhline(inj_m_min,color='blue')
-            axes[3].axhline(inj_m_max,color='blue')
-            axes[4].axhline(inj_sigma_ecc,color='blue')
             plt.savefig("chain.png")
          axes[-1].set_xlabel("step number");
-         #plt.plot(posterior_pos[:,:,0].T, '-', color='k')
-         #plt.axhline(2,color='blue')
-         #plt.savefig('chain1.png')
-      
-      
-      
+
+        # plt.plot(posterior_pos[:,:,0].T, '-', color='k')
+        # plt.savefig('chain1.png')
       # irint(nsamples,n_walkers,ndim)
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate

### 2023-07-11 original_pop_code/original_powerlaw_mcmc.py

@@ -544,28 +544,21 @@ def _main(raw_args=None):
             dtype=numpy.float64,
             verbose=cli_args.verbose,
         )
-         #Making chain plots to check the convergence of the chains
          import matplotlib.pyplot as plt
-         inj_rate,inj_alpha,inj_m_min,inj_m_max,inj_sigma_ecc = (2,-1,10,50,0.1)
-         fig, axes = plt.subplots(5, figsize=(12, 12), sharex=True)
+         fig, axes = plt.subplots(4, figsize=(12, 12), sharex=True)
          samples = posterior_pos
-         labels = ["$log_{10}(\mathcal{R})$",r"$\alpha$", "$m_{min}$","$m_{max}$","$\sigma_\epsilon$"]
+         labels = ["$log_{10}(\mathcal{R})$",r"$\alpha$", "$m_{min}$","$m_{max}$"]
          for i in range(ndim):
             ax = axes[i]
-            ax.plot(samples[:, :, i].T,'-',color='k', alpha=0.3)
+            ax.plot(samples[:, :, i],'-', alpha=0.3)
             ax.set_xlim(0, len(samples))
             ax.set_ylabel(labels[i])
-            axes[0].axhline(inj_rate,color='blue')
-            axes[1].axhline(inj_alpha,color='blue')
-            axes[2].axhline(inj_m_min,color='blue')
-            axes[3].axhline(inj_m_max,color='blue')
-            axes[4].axhline(inj_sigma_ecc,color='blue')
             plt.savefig("chain.png")
          axes[-1].set_xlabel("step number");
-         #plt.plot(posterior_pos[:,:,0].T, '-', color='k')
-         #plt.axhline(2,color='blue')
-         #plt.savefig('chain1.png')
-         
+
+        # plt.plot(posterior_pos[:,:,0].T, '-', color='k')
+        # plt.savefig('chain1.png')
+
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate
 # from the log10(rate), and the normalization factor for the mass

### 2023-07-11 original_pop_code/prob.py

@@ -194,8 +194,8 @@ def joint_pdf(
     tmp_TS = numpy.empty(TS, dtype=numpy.float64)
 
     # Initialize two shape ``T+S`` index arrays, to be reused.
-    tmp_TS_i1 = numpy.empty(TS, dtype=numpy.bool)
-    tmp_TS_i2 = numpy.empty(TS, dtype=numpy.bool)
+    tmp_TS_i1 = numpy.empty(TS, dtype=bool)
+    tmp_TS_i2 = numpy.empty(TS, dtype=bool)
 
     # Create array of booleans determining which combinations of masses and
     # parameters do not correspond to zero probability and also have
@@ -408,7 +408,7 @@ def pdf_const(alpha, m_min, m_max, M_max, out=None, where=True):
         result = out
 
     # Initialize temporary shape ``S`` array to hold booleans.
-    tmp_i = numpy.zeros(S, dtype=numpy.bool)
+    tmp_i = numpy.zeros(S, dtype=bool)
 
     # Determine where the special case ``beta == 0`` occurs, and evaluate the
     # normalization constant there.
@@ -444,7 +444,7 @@ def _pdf_const_special(
     S = m_min.shape
 
     # Initialize temporary shape ``S`` array to hold booleans.
-    tmp_i = numpy.zeros(S, dtype=numpy.bool)
+    tmp_i = numpy.zeros(S, dtype=bool)
 
     # Separately handle populations that are affected by the M_max cutoff and
     # those that are not.
@@ -568,7 +568,7 @@ def _pdf_const_nonspecial(
     # ``False``.
     tmp1 = numpy.zeros(S, dtype=numpy.float64)
     tmp2 = numpy.zeros(S, dtype=numpy.float64)
-    tmp_i = numpy.zeros(S, dtype=numpy.bool)
+    tmp_i = numpy.zeros(S, dtype=bool)
 
     # Compute the indices where ``beta`` is an integer.
     # Then, for each of those indices, compute the normalization constant using
@@ -625,7 +625,7 @@ def _pdf_const_nonspecial_nonintegral(
     # Initialize temporary shape ``S`` boolean array to all ``False``, as we
     # want to use ``where`` to only modify certain indices, and the ones left
     # un-modified need to be ``False``.
-    tmp_i = numpy.zeros(S, dtype=numpy.bool)
+    tmp_i = numpy.zeros(S, dtype=bool)
 
     # Determine which indices need to be computed with the ``M_max`` cutoff
     # in effect, and compute them.

### 2023-07-11 original_pop_code/utils_top.py

@@ -213,7 +213,7 @@ def quantile(a, q, axis=None, weights=None, out=None,
         a = asanyarray(a)
         wgt = np.asanyarray(weights)
 
-        if issubclass(a.dtype.type, (np.integer, np.bool_)):
+        if issubclass(a.dtype.type, (np.integer, bool_)):
             result_dtype = np.result_type(a.dtype, wgt.dtype, 'f8')
         else:
             result_dtype = np.result_type(a.dtype, wgt.dtype)

### 2023-07-11 run2.sh

@@ -1,0 +1,1 @@
+time python -mcProfile new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 20 --n-samples 30

### 2023-07-13 run.sh
Correcting run file

@@ -1,1 +1,1 @@
-time python new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat events/GW151012_GWTC-1.dat events/GW151226_GWTC-1.dat events/GW170104_GWTC-1.dat events/GW170608_GWTC-1.dat events/GW170729_GWTC-1.dat events/GW170809_GWTC-1.dat events/GW170814_GWTC-1.dat events/GW170818_GWTC-1.dat events/GW170823_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 120 --n-samples 20
+time python -mcProfile new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 20 --n-samples 30

### 2023-07-13 run.sh
Correcting run

@@ -1,1 +1,1 @@
-time python -mcProfile new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 20 --n-samples 30
+time python -mcProfile new_powerlaw_mcmc.py events/*.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 20 --n-samples 30

### 2023-07-13 run2.sh

@@ -1,1 +1,0 @@
-time python -mcProfile new_powerlaw_mcmc.py events/GW150914_GWTC-1.dat vt_m1_m2.hdf5 myout.hdf5 --n-walkers 20 --n-samples 30

### 2023-07-13 run_syn_new.sh

@@ -1,1 +1,0 @@
-time python -mcProfile new_powerlaw_mcmc.py syn_events/event_1.dat syn_events/event_2.dat syn_events/event_3.dat syn_events/event_4.dat syn_events/event_5.dat syn_events/event_6.dat syn_events/event_7.dat syn_events/event_8.dat syn_events/event_9.dat syn_events/event_10.dat syn_events/event_11.dat syn_events/event_12.dat syn_events/event_13.dat syn_events/event_14.dat syn_events/event_15.dat syn_events/event_16.dat syn_events/event_17.dat syn_events/event_18.dat syn_events/event_19.dat syn_events/event_20.dat syn_events/event_21.dat syn_events/event_22.dat syn_events/event_23.dat syn_events/event_24.dat syn_events/event_25.dat syn_events/event_26.dat syn_events/event_27.dat syn_events/event_28.dat syn_events/event_29.dat syn_events/event_30.dat syn_events/event_31.dat syn_events/event_32.dat syn_events/event_33.dat syn_events/event_34.dat syn_events/event_35.dat syn_events/event_36.dat syn_events/event_37.dat syn_events/event_38.dat syn_events/event_39.dat syn_events/event_40.dat syn_events/event_41.dat syn_events/event_42.dat syn_events/event_43.dat syn_events/event_44.dat syn_events/event_45.dat syn_events/event_46.dat syn_events/event_47.dat syn_events/event_48.dat syn_events/event_49.dat syn_events/event_50.dat syn_events/event_51.dat syn_events/event_52.dat syn_events/event_53.dat syn_events/event_54.dat syn_events/event_55.dat syn_events/event_56.dat syn_events/event_57.dat syn_events/event_58.dat syn_events/event_59.dat syn_events/event_60.dat syn_events/event_61.dat syn_events/event_62.dat syn_events/event_63.dat syn_events/event_64.dat syn_events/event_65.dat syn_events/event_66.dat syn_events/event_67.dat syn_events/event_68.dat syn_events/event_69.dat syn_events/event_70.dat syn_events/event_71.dat syn_events/event_72.dat syn_events/event_73.dat syn_events/event_74.dat syn_events/event_75.dat syn_events/event_76.dat syn_events/event_77.dat syn_events/event_78.dat syn_events/event_79.dat syn_events/event_80.dat syn_events/event_81.dat syn_events/event_82.dat syn_events/event_83.dat syn_events/event_84.da
[…]

### 2023-07-13 paper/paper.tex

@@
 \begin{abstract}
-The astrophysical mergers produce gravitational waves (GWs), which carry insightful information of their sources such as Binary Black holes (BBHs) mergers.
+% Fine for thesis, drop for paper
+Compact binary  mergers produce gravitational waves (GWs), which carry  information about their sources.
 The LIGO-VIRGO-KAGRA (LVK) detects GWs signals, and we use those measurements to infer BBHs population properties such as mass, spin, and eccentricity distribution in the universe.
-Although mass and spin distribution gives information of the source formation, evolution, and environment but eccentricity is the unique and measurable signature of strong events.
-Hence, the current parameter estimation (PE) methods of GWs signals do not include the eccentricity due to the lack of waveform with eccentricity.
-Therefore, for our analysis, we generated four synthetic populations of non-spinning, non-precessing, lower mass $(10 M_\odot-50 M_\odot)$, eccentric binary black holes (EBBHs) using a modified power law model, each consisting of $100$ events with different eccentricity distribution $\sigma_\epsilon = (0.05,0.1,0.15,0.2)$.
-Furthermore, to compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we estimate how our sources would be characterized by parameter inferences which omitted the effects of eccentricity.
+Binary component masses and spins  are well-known to provide insight into the source formation, evolution, and environment, but similar outcomes are found from strongly-interacting and gently-interacting formation channels.
+ Unlike these two, eccentricity is a unique and measurable signature of strongly-interacting events.
+Unfortunately, most available parameter estimation (PE) results for GWs signals do not include eccentricity: models for binary merger with eccentricity have minimal availability and parameter coverage.
+ %
+To assess how well eccentricity can be identified, we generated four synthetic populations of non-spinning, non-precessing, lower mass $(10 M_\odot-50 M_\odot)$, eccentric binary black holes (EBBHs) using a modified power law model, each consisting of $100$ events with different eccentricity distribution $\sigma_\epsilon = (0.05,0.1,0.15,0.2)$.
+Furthermore, to compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we estimate how our sources would be characterized by parameter inferences that omitted the effects of eccentricity.
 We apply the Markov Chain Monte Carlo (MCMC) method to constrain the model parameters: event rate $\mathcal{R}$, $\alpha$, $m_{min}$, $m_{max}$, and $\sigma_\epsilon$.
-We are also able to recover the injected eccentricity for each population \textbf{(need to add more specfic information after runs)}.
+We can also recover the injected eccentricity for each population \textbf{(need to add more specific information after runs)}.
 All the tools we used in this analysis are publicly available in a GitHub repository, and one can use them to infer the synthetic population of EBBHs.
@@
 \section{Introduction}
-Black holes are intriguing and compact objects in the universe \cite{Frolic_BH_Book_2011}.
-They get exciting and enlightening when they coalesce and emit gravitational waves (GWs) \cite{Frolic_BH_Book_2011, Indrajit_GW_intro_1999}.
-These waves carry insightful information about the BBH system, such as mass, spin, period, eccentricity, distance, location, formation, and evolution.
-Primarily, gravitational waves may be the only way to detect those chaotic mergers in space because there was no physical evidence before LIGO's \cite{LIGO_2015} first detection of a BBH merger named GW150914.
-Afterward, LIGO-VIRGO-KAGRA (LVK) \cite{LIGO_2015, VIRGO_2012, Virgo-2015,kagra-2013} are detecting gravitational waves from various compact binaries such as binary black holes (BBHs), binary neutron stars (BNS), and binary made of a neutron star and black hole.
-The researchers use those GWs signal to infer the intrinsic and extrinsic properties of associated binary such as mass, spin, and eccentricity.
 
-Mass and spin are widely used to understand a binary's evolution and formation channel, ignoring the significant parameter eccentricity.
-Because the existence of eccentricity makes things difficult to comprehend and challenges the current understanding of the formation, environment, and evolution of such compact objects, various scenarios may develop eccentricity in BBHs orbits, such as stellar scattering, dynamic interactions in dense environments, or interaction of the third object with a binary.
+Black holes exist throughout the universe \cite{Frolic_BH_Book_2011}.
+Sometimes pairs of black holes can coalesce and emit gravitational waves (GWs) \cite{Frolic_BH_Book_2011,Indrajit_GW_intro_1999}.
+These waves carry information about the BBH system, such as mass, spin, period, eccentricity, distance, and location.
+Indirectly, these properties may provide clues into how BBHs formed.
+Gravitational waves are the primary way to detect those chaotic mergers in space, starting with LIGO's \cite{LIGO_2015} first detection of a BBH merger (GW150914).
+Afterward, LIGO-VIRGO-KAGRA (LVK) \cite{LIGO_2015, VIRGO_2012, Virgo-2015,kagra-2013} continued to regularly detect gravitational waves from various compact binaries such as binary black holes (BBHs), binary neutron stars (BNS), and binaries composed of a neutron star and black hole \cite{gwtc-1-2019,gwtc-2-2020,gwtc-2.1-2021,gwtc-3-2021}.
+%From the GWs signal to infer the intrinsic and extrinsic properties of associated binary such as mass, spin, and eccentricity.
+
+Mass and spin are widely used to understand a binary's properties, evolution and formation channel.  
+%ignoring the significant parameter eccentricity.
+Because the existence of eccentricity makes things difficult to comprehend and challenges the current understanding of the formation, environment, and evolution of such compact objects,
+However, many of these same formation scenarios also can introduce eccentricity in BBHs orbits, such as stellar scattering, dynamic interactions in dense environments, or interaction of the third object with a binary.
 Therefore, a more insightful method to understand different populations and determine how a binary may have formed is to constrain the eccentricity along with mass and spin \cite{Rod-2018, zevin-samsing-2019, samsing-2018, Rodriguez-2018, Antonini-2014}.
 
-Low energy and slow-forming binaries usually start with lesser eccentricities, and eventually, gravitational wave radiation circularizes their orbits \cite{Peters-1964} before entering the LVK frequency band.
-However, binaries formed in more violent, energetic environments, such as globular clusters, can have large eccentricities.
-These binaries retain measurable eccentricities after entering the LVK sensitivity band, potentially making orbital eccentricity an important parameter in determining formation channels \cite{Rod-2018,zevin-samsing-2019,samsing-2018, Rodriguez-2018,Antonini-2014}.
-We may also have stellar mass higher eccentric mergers at the lower frequency searches \cite{sesana-2016,chen-2017}, which can be measurable by detectors like Laser Interferometer Space Antenna (LISA) \cite{LISA-2017}.
-These observations would allow for long-term tracking of BBHs orbital properties, which can be used to infer the formation mechanism \cite{Breivik-2016}.
+Low energy and slow-forming binaries usually start with modest eccentricities but at very low frequencies; over many decades of frequency evolution, gravitational wave radiation circularizes their orbits \cite{Peters-1964} before entering the LVK frequency band.
+However, binaries formed in more violent, energetic environments, such as globular clusters, can be formed closer to the LVK frequency band and thus retain more of their large natal eccentricities when observationally accessible.
+Previous studies have demonstrated that identifying  orbital eccentricity could differentiate between formation channels \cite{Rod-2018,zevin-samsing-2019,samsing-2018, Rodriguez-2018,Antonini-2014}.
 
-Although eccentricity is a signature that can be measured with GWs and is unique to extreme events, all the confirmed detection till now shows a nearly circular orbit in the LVK frequency band.
-The first potential reason for getting circular orbits is that eccentric effects are more evident in low-mass events, and current searches are more efficient for higher-mass mergers.
-Secondly, it may result from selection biases in the waveforms because LVK detectors only use circular waveforms for parameter estimation (PE), which better present binaries evolved in an isolated environment.
-Studies show that we must consider the multiple formation channels to understand the population better.
-Because a single channel does not contribute more than $70\%$ of the observation sample of BBHs \cite{zevin-2021}.
-We also have growing shreds of evidence that suggests BBHs like GW190521 may exhibit unambiguous eccentricity or faint signatures \cite{Gamba_2022_GW190521_dynamical,yumeng-2023,Isobel-2022}, which leads towards the dynamical formation of the source.
-So, it is critical for astrophysical implications to assess eccentricity distribution.
+Although eccentricity is a signature that can be measured with GWs and is unique to extreme events, all the confirmed detections so far may be consistent with a nearly circular orbit in the LVK frequency band \cite{Isobel-2022,hector-jake-2022}.
+However, some investigations suggest that  BBHs like GW190521 may exhibit some indications of eccentricity \cite{Gamba_2022_GW190521_dynamical,yumeng-2023}.
+Unfortunately, these inferences rely on incompletely-surveyed waveforms, produced either by direct simulation or (more customarily) by phenomenological approximations tuned to those simulations.
+At present, available phenomenological models which allow for eccentricity only cover part of the possible parameters: nonprecessing binaries, for example.
+While researchers are actively producing the eccentric waveform models analytically \cite{Huerta-2014} and numerically \cite{gold-2016,hinder-2010, Healy-2022, Alessandro-2022, Campanelli-2009}, at present parameter inference capabilities are limited.
 
-The sensitivity of LVK detectors is increasing with time, leading to more detection in each observing run.
-We will have hundreds of events in future runs, and with the upcoming O4 run, we expect detection every other day \cite{detection_rate_2016,detection_rate_2015}.
-Interestingly, the new detections lead us to the reasonable disagreement on the masses, spin, event rate, and formation scenarios \cite{LSC-BBH-2016, LSC-GW150914-2016}, which pushes researchers to develop new models \cite{Mandel-2016,marchant-2016}, and also provide more information about the observable GWs \cite{Barausse-2018, Abbot-2016-schotastic,dvorkin-2016}.
-Additionally, researchers are actively producing the eccentric waveform models analytically \cite{Huerta-2014} and numerically \cite{gold-2016,hinder-2010, Healy-2022, Alessandro-2022, Campanelli-2009}.
-Those waveforms in the bank will give us real PE with eccentricity shortly.
-Therefore, identifying and understanding the populations of EBBHs with a growing number of detection will give us more insightful information about the formation and evolution of massive stars from birth to death over cosmic time.
+Despite severe limitations on single-event parameter inference, a few proof-of-concept studies have investigated how to identify the presence of an eccentric subpopulation from observations of many massive binary black holes
+\cite{lower-2018-ecc-pop,Fang-2019,wu-2020}.
+The techniques used in these investigations borrow from extensive studies on reconstructing the population properties of quasicircular binaries over the whole mass spectrum
+\cite{Dan_2019, Michael-2015, Mandel_2017_Errors, samsing-hamers-2019,  belczynski-2016, Colm-2017, Akinobu-2017, Michael-zevin-2017, farr-2017-nature, Richard-2017-natal-kicks, Dan-Richard-2018,Abbot-2019-pop}.
+Because eccentricity is so poorly constrained by short GW observations of massive BH binaries, these studies find eccentricity can only be resolved with great difficulty, even with many observations.
 
-Recent works investigate how measurements of BBHs mass and spin distributions can elucidate the population properties of BBHs \cite{Dan_2019, Michael-2015, Mandel_2017_Errors, samsing-hamers-2019,  belczynski-2016, Colm-2017, Akinobu-2017, Michael-zevin-2017, farr-2017-nature, Richard-2017-natal-kicks, Dan-Richard-2018,Abbot-2019-pop}.
-Researchers also study the population of BBHs with eccentricity \cite{lower-2018-ecc-pop,Fang-2019}.
-However, most research is inclined towards the high mass range or real events \cite{wu-2020}.
-Therefore, this paper focused on the non-spinning, non-precessing, lower mass $(10 M_\odot - 50 M_\odot)$, and lesser eccentric $(0-0.2)$ merger.
-In addition, we will use mass ratio $q=m_1/m_2$ with condition $m_1>m_2$ and total mass $M=m_1+m_2=100M_\odot$.
-We present the correlation of event rate, mass, and eccentricity distribution because studies show that binary parameters and event rate are strongly coupled.
+ %
+By contrast, for low-mass binaries, previous studies have  demonstrated that eccentricities can be particularly well-constrained for low-mass objects, owing to their long modulated inspiral   \cite{favata-scaling-2022}.
+We use this parameter inference investigation as our prototype for the synthetic inferences about synthetic GW sources used in our proof-of-concept study.
+
+In this study, we focused on the non-spinning, non-precessing, lower mass $(10 M_\odot - 50 M_\odot)$, and lesser eccentric $(0 - 0.2)$ merger.
+In addition, we will use mass ratio $q = m_1 / m_2 $ with condition $m_1>m_2$ and total mass $M = m_1 + m_2 = 100 M_\odot$.
 We also infer how well we can recover the mass, eccentricity, and event rate using the $100$ eccentric events.
 To extend our analysis, we compared constrained parameters using the EBBBhs and NEBBHs, which show a considerable difference.
+
+%We present the correlation of event rate, mass, and eccentricity distribution because studies show that binary parameters and event rate are strongly coupled.
 
@@
 
-%\begin{itemize}
-%    \item We have N discrete detection.
-%    \item each detection has data $d_1,d_2,d_3,...,d_N$
-%    \item each data point $d_N$ has properties denoted by $\lambda_1,\lambda_2,...,\lambda_i$
-%    \item  $ \lambda_i$ are parameters such as mass, spin, eccentricity, and location.
-%    \item Each parameter has its uncertainty.
-%    \item This uncertainty is described by the probability of the data given the parameter value. We call it the likelihood function of one single event. $\mathcal{L}(\lambda) = p(d|\lambda)$
-%    \item the waveform gives the parameter value.
-%    \item waveform is a mathematical function.
-%    \item Once you have a likelihood function, you can use prior( usually it's uniform: which gives equal probability to each event) to find the posterior.
-$p(\lambda|d) \propto p(d|\lambda) p(\lambda)$
-
-%\end{itemize}
-
-\subsection{Population Inference}
+\subsection{Bayesian Inference}
 
@@
 %
-where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or also known as evidence.
+where $p(\Lambda|d_1,d_2,d_3,...,d_N)$ is posterior, $p(\Lambda)$ is prior, and $p(d_1,d_2,d_3,...,d_N)$ is normalization constant or also known as evidence.
 
@@
 \ref{fig:pop3d0.05_0.2}.
-The complete description of each EBBHs population is provided in Table \ref{tab:pop_prop}.
 
@@
 
-\begin{table*}[]
-    \centering
-    \begin{tabular}{c|cccccc}
-        \hline \hline
-        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{2_{max}} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
-        First  & 12.19 & 49.89 & 10.17 & 46.67 & 0.0004 & 0.0494\\ \hline
-        Second  & 15.42 & 49.79 & 10.46 & 47.59 & 0.0003 & 0.0999\\ \hline
-        Third  & 16.17 & 49.83 & 10.69 & 47.52 & 0.0002 & 0.148\\ \hline
-        Forth  & 12.52 & 49.89 & 10.08 & 49.51 & 0.0012 & 0.1984\\ \hline
-    \end{tabular}
-    \caption{Population Properties of EBBHs}
-    \label{tab:pop_prop}
-\end{table*}
+%\begin{table*}[]
+%    \centering
+%    \begin{tabular}{c|cccccc}
+%        \hline \hline
+%        Populations & $m_{1_{min}} [M_\odot] $ & $m_{1_{max}} [M_\odot]$ & $m_{2_{min}} [M_\odot]$ & $m_{2_{max}} [M_\odot]$ & $\epsilon_{min}$ & $\epsilon_{max}$\\ \hline
+%        First  & 12.19 & 49.89 & 10.17 & 46.67 & 0.0004 & 0.0494\\ \hline
+%        Second  & 15.42 & 49.79 & 10.46 & 47.59 & 0.0003 & 0.0999\\ \hline
+%        Third  & 16.17 & 49.83 & 10.69 & 47.52 & 0.0002 & 0.148\\ \hline
+%        Forth  & 12.52 & 49.89 & 10.08 & 49.51 & 0.0012 & 0.1984\\ \hline
+%    \end{tabular}
+%    \caption{Population Properties of EBBHs}
+%    \label{tab:pop_prop}
+%\end{table*}
 
@@
 We used the following relation for conversion, and it will provide the masses based on the condition $m_1\geq m_2$.
-
-\begin{equation}
-    m_1 = \frac{1}{2} M_c \eta^{-3/5} (1+\sqrt{\eta_v}),
-\end{equation}
-
-\begin{equation}
+%
+\begin{align}
+    m_1 = \frac{1}{2} M_c \eta^{-3/5} (1+\sqrt{\eta_v}), \\
     m_2 = \frac{1}{2} M_c \eta^{-3/5} (1-\sqrt{\eta_v}), 
-\end{equation}
+\end{align}
 where $\eta_v = 1-4\eta$, we kept the samples with non-negative values and ignored the negative samples to avoid the square root issues.
@@
 To compare the synthetic EBBHs population with the non-eccentric BBHs (NEBBHS), we estimate how our sources would be characterized by parameter inferences which omitted the effects of eccentricity.
-Following \cite{2021_scaling_paper}, the following effective chirp mass is well-constrained by observations dominated by the inspiral of a slightly eccentric binary:
+Following \cite{favata-scaling-2022}, the following effective chirp mass is well-constrained by observations dominated by the inspiral of a slightly eccentric binary:
 \begin{align}
@@
 
-%To get the non-eccentric population, we apply the Eq.
-\ref{eq:scaling} on the population given in the table.
-\ref{tab:pop_prop}.
-After applying the scaling equation, we get the new populations with scaled masses, giving us the new minimum and maximum values for the primary and secondary masses given in table \ref{tab:popscl_prop}.
-
 The significant effect after scaling is the mass shift, which can be observed in Fig.
@@
 \label{sec:pop_inference}
-%\renewcommand{\arraystretch}{2}
-\begin{table*}[]
+We have use the prior's given in the table \ref{tab:prior}
+
+\begin{table*}
     \centering
@@
         \hline \hline
-       Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
-      First& $1.42^{+0.58}_{-2.64}$ & $-0.88^{+3.17}_{-1.89}$ & $8.36^{+7.49}_{-3.51}$ & $58.44^{+10.21}_{-24.31}$ & $0.30^{+0.54}_{-0.25}$ \\ \hline
-      Second & $1.36^{+0.63}_{-2.65}$ & $-0.93^{+3.29}_{-1.92}$ & $8.71^{+7.36}_{-3.81}$ & $57.84^{+10.81}_{-25.53}$ & $0.33^{+0.51}_{-0.25}$ \\ \hline
-      Third & $1.39^{+0.58}_{-2.69}$ & $-0.69^{+2.98}_{-1.95}$ & $9.21^{+6.31}_{-3.99}$ & $57.85^{+10.73}_{-27.04}$ & $0.33^{+0.51}_{-0.23}$  \\ \hline
-      Forth & $1.27^{+0.72}_{-2.68}$ & $-0.90^{+3.30}_{-1.98}$ & $8.63^{+7.85}_{-3.72}$ & $57.76^{+11.03}_{-25.82}$ & $0.37^{+0.47}_{-0.23}$  \\ \hline
+       Quantity & $\log_{10}(\frac{\mathcal{R}}{\rm Gpc^{-3}yr^{-1}})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
+      Synthetic population & 2 & -1 & 10 & 50 & 0.05-0.2 \\ \hline
+      Prior range & $[-3,3]$ & $[-5,5]$ & $[1-20]$ & $[30-80]$ & $[0-0.5]$ \\ \hline
+      Prior distribution & Log-uniform & Uniform & Uniform & Uniform & Uniform  \\ \hline \hline
     \end{tabular}
-    \caption{Population Inference for EBBHs}
-    \label{tab:inference_EBBHS}
-\end{table*}
-
-\begin{table*}[]
-    \centering
-    \begin{tabular}{c|cccc}
-        \hline \hline
-        Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ \\ \hline
-      First & $1.85^{+0.14}_{-1.23}$ & $-1.37^{+1.43}_{-0.94}$ & $7.54^{+3.68}_{-3.04}$ & $63.29^{+8.00}_{-13.08}$  \\ \hline
-      Second & $1.83^{+0.14}_{-1.38}$ & $-1.12^{+1.47}_{-0.91}$ & $7.69^{+3.88}_{-3.15}$ & $61.91^{+8.59}_{-13.87}$  \\ \hline
-      Third & $1.84^{+0.15}_{-1.55}$ & $-0.38^{+1.03}_{-0.96}$ & $8.07^{+3.44}_{-3.27}$ & $61.70^{+8.49}_{-24.26}$   \\ \hline
-      Forth & $1.86^{+0.16}_{-1.85}$ & $-0.31^{+1.17}_{-1.14}$ & $6.99^{+8.77}_{-2.58}$ & $61.15^{+8.72}_{-16.84}$  \\ \hline
-    \end{tabular}
-    \caption{Population Inference for NEBBHs}
-    \label{tab:inference_NEBBHS}
+    \caption{Parameter used to generate synthetic population and prior used for the inference}
+    \label{tab:prior}
 \end{table*}
@@
 \end{figure*}
-Same population.
+
 \begin{figure*}
@@
 
-Need to write it later.
+%\renewcommand{\arraystretch}{2}
+\begin{table*}
+    \centering
+    \begin{tabular}{c|ccccc}
+        \hline \hline
+       Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ & $\sigma_\epsilon$ \\ \hline
+      First& $1.42^{+0.58}_{-2.64}$ & $-0.88^{+3.17}_{-1.89}$ & $8.36^{+7.49}_{-3.51}$ & $58.44^{+10.21}_{-24.31}$ & $0.30^{+0.54}_{-0.25}$ \\ \hline
+      Second & $1.36^{+0.63}_{-2.65}$ & $-0.93^{+3.29}_{-1.92}$ & $8.71^{+7.36}_{-3.81}$ & $57.84^{+10.81}_{-25.53}$ & $0.33^{+0.51}_{-0.25}$ \\ \hline
+      Third & $1.39^{+0.58}_{-2.69}$ & $-0.69^{+2.98}_{-1.95}$ & $9.21^{+6.31}_{-3.99}$ & $57.85^{+10.73}_{-27.04}$ & $0.33^{+0.51}_{-0.23}$  \\ \hline
+      Forth & $1.27^{+0.72}_{-2.68}$ & $-0.90^{+3.30}_{-1.98}$ & $8.63^{+7.85}_{-3.72}$ & $57.76^{+11.03}_{-25.82}$ & $0.37^{+0.47}_{-0.23}$  \\ \hline \hline
+    \end{tabular}
+    \caption{Population Inference for EBBHs}
+    \label{tab:inference_EBBHS}
+\end{table*}
 
-% \begin{align} \label{eq:strain_mode}
-% h(t,\vartheta,\phi;\bm{\lambda}) = 
-% \sum_{\ell=2}^{\infty} \sum_{m=-\ell}^{\ell} \frac{D_{\rm ref}}{D} h^{\ell m}(t;\bm{\lambda}) \Y{-2}_{\ell m} \left(\vartheta, \phi \right) \, ,
-% \end{align}
-
-% \begin{widetext}
-% \begin{align}
-% \ln {\cal L}(\bm{\lambda}, \theta) 
-% &= (D_{\rm ref}/D) \text{Re} \sum_k \sum_{\ell m}(F_k \Y{-2}_{\ell m})^* Q_{k,lm}(\bm{\lambda},t_k)\nonumber \\
-% &   -\frac{(D_{\rm ref}/D)^2}{4}\sum_k \sum_{\ell m \ell' m'}
-% \left[
-% {
-% |F_k|^2 [\Y{-2}_{\ell m}]^*\Y{-2}_{\ell'm'} U_{k,\ell m,\ell' m'}(\bm{\lambda})
-% }
-% % \right. \nonumber \\ & \left.
-%  {
-% +  \text{Re} \left( F_k^2 \Y{-2}_{\ell m} \Y{-2}_{\ell'm'} V_{k,\ell m,\ell'm'} \right)
-% }
-% \right]
-% \label{eq:def:lnL:Decomposed}
-% \end{align}
-% \end{widetext}
-
-% \begin{eqnarray}
-% {\cal L}_{\rm margT} \equiv  \int {\cal L} \frac{dt}{T}
-% \label{eq:lnL:tmarg}
-% \end{eqnarray}
-
-% \begin{table*}
-% \begin{tabular}{lrr|ccccc|rr}
-% Version & srate & modes & $\tau_{start}$ & $\tau_{setup}$ & $\tau_{ad}$ & $\tau_{it,like}$ &$\tau_{it,rest}$ &
-% $\frac{T_{ILE}}{N_{eval}}$ & GPU \\  %\hline 
-%   &   Hz & m & sec & sec & & $\mu$sec & $\mu$sec  &sec  & use  \%\\ \hline 
-% % ~/parse_report.sh profile_nogpu_pcdev13.log | more
-% CPU & 16384 & $\pm 2 $ & 20 & 2.4 &&540 & 20 &  690  \\ 
-%        & 4096 & $\pm 2 $ &   20  &&&& 20 \\ \hline
-% % ./parse_report.sh 20190130-profile_nogpu_HM_pcdev13.log  | more
-% % /parse_report.sh ./profile_nogpu_HM_lowres_pcdev13.log 
-% %    setup time: PrecomputeLikelihoodTerms, includes waveform generation. 
-% %   evaluation: FactoredLogLikelihodTimeMarginalized Divide by actual number of calls, since not a block!
-% %    
-% CPU & 16384 & $\pm 2,\pm 1 $ & 20 & 1.5 && 680 & 20 &  1060  \\ 
-%        & 4096 & $ \pm 2, \pm 1 $ &   20 &&&& 20  \\ \hline
-
-%GPU (a) & 16384 & $\pm 2 $  & 20  & & && & 270 \\
-%            & 4096 &$\pm 2 $  &  20 &  & & & & 45 \\ \hline
-
-%GPU (b) & 16384 & $\pm 2$ & 20  & 1.8 & 1 & 0.85& 20 &28 & 15\\
-%        & 4096 & $\pm 2$  & 20 & $1.2 $ &  1  & 0.75 & 20  & 25\\ \hline
-
-% GPU (b) & 16384 & $\pm 2, \pm 1$ & 20 & 1 && 4.2 & 20  & 38  \\
-%        & 4096 & $\pm 2, \pm 1$ & 20 & 1&& 2.5  & 20 & 35 & \\ \hline
-
-% GPU (c) & 16384 & $\pm 2 $  & 20  &6  & & 18&  58&160 &  \\
-%             & 4096 &$\pm 2 $  &  20 & 3.7 & & 11  & 58  & 140 & $\simeq 50$ \\
-% \end{tabular}
-% Compute node at LIGO-WA
-% \caption{\label{tab:CostBreakdown}\textbf{Profiling performance: Binary black holes}: Evaluation costs for the
-%   marginalized likelihood on default
-%   hardware, for a two-mode system $(l,m)=\pm 2$ analyzing $T=8\unit{s}$ of data with a massive binary black hole
-%   $m_1=35 M_\odot,M_2=30 M_\odot$.  The last column indicates peak GPU utilization.
-% }
-%\end{table*}
-
-% \subsection{Binary black hole analysis}
-% \label{sec:sub:BBHFull}
-
-% \begin{figure*}
-
-% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_m2.png}
-% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_m1_cum.png}
-% % python plot_mean_variance.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency_meanVar.dat  --output bbh_zerospin_lnL_meanVar
-% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_meanVar.png}
-% % python plot_convergence.py --convergence-file 20190203-bbh-zerospin-batch_gpu_lowlatency.dat
-% \includegraphics[width=0.45\textwidth]{figures/bbh_zerospin_lnL_converge.png}
-% \caption{\label{fig:BBH:MultiIterate}\textbf{Convergence of BBH analysis: Zero spin}: Results for marginal posterior distributions
-%   of our fiducial synthetic binary black hole.  Solid contours show credible intervals; solid one-dimensional distributions
-%   show marginal CDFs and PDFs for the corresponding variable; and colored points indicate the location $\bm{\lambda}$ and
-%   value of the underlying marginalized likelihood evaluations.   
-% %The corresponding dotted curves show an analysis using   only the $m=\pm 2$ modes \editremark{perform both} 
-% \emph{Left panel } Posterior distribution
-%   over  $\mc$ and
-%   $\delta=(m_1-m_2)/M$.    \emph{Right panel}: Marginal 1d CDFs of $\mc$, showing convergence.
-% \emph{Bottom left}: Mean and variance of  \AddedResponse{the array $\ln{\cal L}_{\rm marg}(\bm{\lambda}_j)$  for
-% $j=1,2,\ldots N_{\rm eval}$ indexing all candidate sets of intrinsic parameters $\bm{\lambda}_j$ performed in that iteration},  showing that after the
-% first iteration the
-% candidate points are consistent with the posterior (i.e., no proposed point has very low $\ln {\cal L}_{\rm marg}$).
-% \emph{Bottom right panel}: The estimated evidence $Z = \int d\bm{\lambda} {\cal L}_{\rm marg}$ versus iteration number.  As systematic fitting error dominates our
-% error budget, Monte Carlo error is not shown.
-% }
-% \end{figure*}
+\begin{table*}[]
+    \centering
+    \begin{tabular}{c|cccc}
+        \hline \hline
+        Inference & $log_{10}(\frac{\mathcal{R}}{Gpc^{-3}yr^-1})$ & $\alpha$ & $m_{min} [M_\odot] $ & $m_{max} [M_\odot]$ \\ \hline
+      First & $1.85^{+0.14}_{-1.23}$ & $-1.37^{+1.43}_{-0.94}$ & $7.54^{+3.68}_{-3.04}$ & $63.29^{+8.00}_{-13.08}$  \\ \hline
+      Second & $1.83^{+0.14}_{-1.38}$ & $-1.12^{+1.47}_{-0.91}$ & $7.69^{+3.88}_{-3.15}$ & $61.91^{+8.59}_{-13.87}$  \\ \hline
+      Third & $1.84^{+0.15}_{-1.55}$ & $-0.38^{+1.03}_{-0.96}$ & $8.07^{+3.44}_{-3.27}$ & $61.70^{+8.49}_{-24.26}$   \\ \hline
+      Forth & $1.86^{+0.16}_{-1.85}$ & $-0.31^{+1.17}_{-1.14}$ & $6.99^{+8.77}_{-2.58}$ & $61.15^{+8.72}_{-16.84}$  \\ \hline \hline
+    \end{tabular}
+    \caption{Population Inference for NEBBHs}
+    \label{tab:inference_NEBBHS}
+\end{table*}
 
@@
 
+The first potential reason for getting circular orbits is that eccentric effects are more evident in low-mass events, and current searches are more efficient for higher-mass mergers.
+Secondly, it may result from selection biases in the waveforms because LVK detectors only use circular waveforms for parameter estimation (PE), which better present binaries evolved in an isolated environment.
+
+Studies show that we must consider the multiple formation channels to understand the population better.
+Because a single channel does not contribute more than $70\%$ of the observation sample of BBHs \cite{zevin-2021} \textbf{previous statement speculative/unsupported - remove}.
+
+which leads towards the dynamical formation of the source. So, it is critical for astrophysical implications to assess eccentricity distribution
+
+We may also have stellar mass higher eccentric mergers at the lower frequency searches \cite{sesana-2016,chen-2017}, which can be measurable by detectors like Laser Interferometer Space Antenna (LISA) \cite{LISA-2017}.
+These observations would allow for long-term tracking of BBHs orbital properties, which can be used to infer the formation mechanism \cite{Breivik-2016}.
+
+The sensitivity of LVK detectors is increasing with time, leading to more detection in each observing run.
+We will have hundreds of events in future runs, and with the upcoming O4 run, we expect detection every other day \cite{detection_rate_2016,detection_rate_2015}.
+Interestingly, the new detections lead us to the reasonable disagreement on the masses, spin, event rate, and formation scenarios \cite{LSC-BBH-2016, LSC-GW150914-2016}, which pushes researchers to develop new models \cite{Mandel-2016,marchant-2016}, and also provide more information about the observable GWs \cite{Barausse-2018, Abbot-2016-schotastic,dvorkin-2016}.
+Therefore, identifying and understanding the populations of EBBHs with a growing number of detection will give us more insightful information about the formation and evolution of massive stars from birth to death over cosmic time.
+
 \begin{acknowledgements}
@@
 
-\appendix
-In Progress,
+%\appendix
+%In Progress,
 

### 2023-07-16 population-genrator/plotting.py
moving populations plots from jupyter to python file

@@ -8,7 +8,8 @@ Created on Thu Jun 15 21:33:41 2023
 import numpy as np
 import matplotlib.pyplot as plt
 import glob
-
+from mpl_toolkits.mplot3d import Axes3D
+import mplcursors
 # Get a list of all files matching the pattern
 file_list = glob.glob("./rPE_events/event_*.dat")
 
@@ -28,8 +29,8 @@ plt.title("Scatter Plot")
 plt.xlabel("$m_1 [M_\odot]$")
 plt.ylabel("$m_2 [M_\odot]$")
 #plt.legend()  # Display legend
-plt.show()
 plt.savefig("scatter_plot.png")
+plt.figure()
 
 
 # Plotting the whole population with median and std of each event
@@ -39,6 +40,51 @@ plt.rc('lines', **lines)
 plt.xlabel("$m_1 [M_\odot]$")
 plt.ylabel("$m_2 [M_\odot]$")
 plt.plot(combined_data[:,0], combined_data[:,1], 'ro', markersize=10)
-plt.show()
 plt.savefig("mean_masses.png")
+plt.figure()
 
+# making population plots
+# Read the data from the text file
+data = np.loadtxt("weighted_population.dat")
+# Extract the columns
+m1 = data[:, 0]
+m2 = data[:, 1]
+ecc = data[:, 2]
+
+# Create a 3D plot
+fig = plt.figure()
+ax = fig.add_subplot(111, projection='3d')
+
+# Plot the data points as dots with colors
+sc = ax.scatter(m1, m2, ecc, c=ecc, cmap='plasma', marker='o')
+
+# Set labels for the axes
+ax.set_xlabel('$m_1[M_\odot]$')
+ax.set_ylabel('$m_2[M_\odot]$')
+ax.set_zlabel('$\epsilon$')
+
+# Set a title for the plot
+#ax.set_title('Eccentric Synthetic Population')
+
+# Add interactivity with mplcursors
+cursors = mplcursors.cursor(sc, hover=True)
+cursors.connect("add", lambda sel: sel.annotation.set_text(f"({sel.target[0]:.2f}, {sel.target[1]:.2f}, {sel.target[2]:.2f})"))
+
+# Add a colorbar
+cbar = fig.colorbar(sc)
+cbar.set_label('$\epsilon$')
+plt.savefig('pop3d_0.05.png', bbox_inches='tight')
[…]

### 2023-07-16 new_powerlaw_mcmc.py
Seprate chain plots

@@ -589,10 +589,24 @@ def _main(raw_args=None):
             ax.set_xlim(0, len(samples))
             ax.set_ylabel(labels[i])
             plt.savefig("chain.png")
+#            plt.figure()
          axes[-1].set_xlabel("step number");
 
-        # plt.plot(posterior_pos[:,:,0].T, '-', color='k')
-        # plt.savefig('chain1.png')
+         plt.figure()
+         plt.plot(posterior_pos[:,:,0], '-', color='k')
+         plt.savefig('rate.png')
+         plt.figure()
+         plt.plot(posterior_pos[:,:,1], '-', color='k')
+         plt.savefig('alpha.png')
+         plt.figure()
+         plt.plot(posterior_pos[:,:,2], '-', color='k')
+         plt.savefig('min.png')
+         plt.figure()
+         plt.plot(posterior_pos[:,:,3], '-', color='k')
+         plt.savefig('max.png')
+         plt.figure()
+         plt.plot(posterior_pos[:,:,4], '-', color='k')
+         plt.savefig('sigecc.png')
       # irint(nsamples,n_walkers,ndim)
 # Functions which pre-compute quantities that are used at multiple steps
 # in the MCMC, to reduce run time. These specifically compute the rate
